### 2024-11-12 paper.tex
Update paper for JCGT review

\documentclass[final]{jcgt}

\usepackage{graphicx}
\usepackage[labelsep=quad,indention=10pt]{subfig}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage{xspace}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}

\usepackage{tikz}
\usetikzlibrary{fadings}
\usetikzlibrary{patterns}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes}
\input{extensionAlgis.sty}
%plot part 
\usepackage{pgfplots} 
\pgfplotsset{compat=1.18}
\usepackage{pgfplotstable}
\usetikzlibrary{intersections}
\usepgfplotslibrary{fillbetween}
\usetikzlibrary{patterns}

\usepackage{multirow}

\setciteauthor{David Algis, Bérenger Bramas, Emmanuelle Darles, Lilian Aveneau}
\setcitetitle{Arc Blanc: a real time ocean simulation framework}

% Mark submissions with the date of submission using the following line:
\submitted{\today}

% Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
\accepted{submitted}{accepted}{published}{Editor Name}{vol}{issue}{1}{1}{year}
\seturl{http://jcgt.org/published/vol/issue/num/}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\Framework{The Arc Blanc framework\xspace}
\def\framework{the Arc Blanc framework\xspace}
\def\etal{\emph{et al.}\xspace}
\def\ie{\emph{i.e.}\xspace}
\def\cf{\emph{cf.}\xspace}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Arc Blanc: a real time ocean simulation framework}

\author{%
    David Algis\\Université de Poitiers%\\Univ. Limoges\\CNRS, XLIM, France\\Studio Nyx
    \and Berenger Bramas\\INRIA Nancy
    \and Emmanuelle Darles\\Université de Poitiers%\\ Univ. Limoges\\ CNRS, XLIM, France
    \and Lilian Aveneau\\Université de Poitiers%\\ Univ. Limoges\\ CNRS, XLIM, France
}

\teaser{
    \includegraphics[width=4in]{Figures/experiment1-contrast.png}
    \caption{Example of rendering of \framework.}
    \label{fig:teaser}
}

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{firstpagestyle}

\maketitle

\begin{abstract}
	\small
	The oceans cover the vast majority of the Earth.
	Therefore, their simulation has many scientific, industrial and military interests, including computer graphics domain.
	By fully exploiting the multi-threading power of GPU and CPU, current state-of-the-art tools can achieve real-time ocean simulation, even if it is sometimes needed to reduce the physical realism for large scenes.
	Although most of the building blocks for implementing an ocean simulator are described in the literature, a clear explanation of how they interconnect is lacking.
	Hence, this paper proposes to bring all these components together, detailing all their interactions, in a comprehensive and fully described real-time framework that simulates the free ocean surface and the coupling between solids and fluid.
	This article also presents several improvements to enhance the physical realism of our model. 
	The two main ones are: 
	calculating the real-time velocity of ocean fluids at any depth; 
	computing the input of the fluid to solid algorithm.
\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Context}
\label{subsec:intro-context}
Simulating the dynamics of an ocean and its surrounding elements remains an important objective in scientific computing.
As all the oceans encompass a substantial portion of the Earth's surface, understanding their behavior is of paramount importance in scientific, industrial, entertainment and military fields.

The computational power of modern GPUs and CPUs, coupled with the latest high-speed memory, has greatly simplified the processing of fluid dynamics simulations.
They make it possible to perform real-time simulations on large discrete grids surrounded by solids, albeit with some physical approximations, thanks to their massively parallel capabilities.

While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (\cf Figure \ref{fig:overviewFramework}).
This new framework incorporates three different methods from the literature:
\begin{itemize}
	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
	\item The simulation of the alteration of a free surface by a solid, inspired by Coords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
\end{itemize}
To bring together these previous works, this new framework proposes two new contributions:
\begin{itemize}
	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
	\item An algorithm to compute the input of the Coords and Staadt method more realistically from the free surface.
\end{itemize}

\begin{figure}
	\centering
	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
	\input{Tikz/methodsInteractions1.tikz}
	\caption{Overview of \framework components: free surface, solid to fluid and fluid to solid coupling.}
	\label{fig:overviewFramework}
\end{figure}

This article is organized as follows.
Section~\ref{sec:relatedwork} presents previous works related to ocean simulation, including its interaction with solid.
Sections \ref{sec:oceanWaves} to \ref{sec:solidToFluidAction} detail \framework, including the ocean surface simulation in Section~\ref{sec:oceanWaves}, the action of the fluid to solids in Section~\ref{sec:fluidToSolidAction} and the action of a solid to the fluid in Section~\ref{sec:solidToFluidAction}.
Section~\ref{sec:results_and_discussions} discusses some results obtained with \framework.
At last, Section~\ref{sec:conclusions} gives general conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Related works}
\label{sec:relatedwork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This section presents the main works related to the simulation of ocean, including the free surface, the fluid to solid and solid to fluid coupling.

\subsubsection{Ocean free surface simulation}
\label{subsubsec:oceanFreeSurfaceSim}
J. Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} presented what could be considered a seminal work in real-time ocean simulation. His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solutions to an approximation of the Navier-Stokes equation on a 2D height field. This method is commonly known as the "Tessendorf method".

The realism of this method was significantly improved by Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, who proposed an empirically based directional spectrum.
In addition, he proposed to add a parameter called "swell", which allows the transition from empirical directional propagation to parallel waves.

Another notable improvement of Tessendorf's method is the reduction of artifact effects through the use of multiple displacement layers.
This technique, introduced by Dupuy and Bruneton (2012) \cite{dupuyRealtimeAnimationRendering2012}, involves stacking multiple displacement layers, each one representing different scales of wave motion.
By combining these layers, the artifacts and repetitive patterns that can often occur in simulated water surfaces are effectively reduced or eliminated.
Finally, Tessendorf (2017) \cite{tessendorfGilliganPrototypeFramework2017} summarized these works in a prototype framework for large-scale water simulation.

\Framework aggregates all these methods in order to simulate the ocean free surface.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Fluid to solid coupling}
\label{subsubsec:fluidToSolidCoupling}
There are several ways to describe the motion of a solid in a given fluid.
Many particle-based methods have been developed in recent years, as presented by Koshier \etal (2022) \cite{koschierSurveySPHMethods2022}.
Akinci \etal (2012) \cite{akinciVersatileRigidfluidCoupling2012} have proposed to calculate contact forces between fluid particles and boundary ones in combination with Predictive-Corrective Incompressible Smoothed Particles Hydrodynamics (PCISPH) \cite{solenthalerPredictivecorrectiveIncompressibleSPH2009}.
Ihmsen \etal (2014) \cite{ihmsenImplicitIncompressibleSPH2014} have applied the same principle to Implicit Incompressible SPH (IISPH), and Bender and Koschier (2017)  \cite{benderDivergenceFreeSPHIncompressible2017} to Divergence Free SPH (DFSPH).
These works have been improved and stabilized by Gissler  \etal \cite{gisslerInterlinkedSPHPressure2019}.
These methods look promising because of their physical accuracy.
Nevertheless, their performance is linked to the number of particles used for the simulation, which makes them impossible to apply in real time for large scale environments.

Another approach related to the dynamics of a ship in water is the calculation of the so-called \emph{Response Amplitude Operator} (RAO), which is expressed using strip theory by Salvesen  \etal (1970) \cite{salvesenShipMotionsSea1970}. Fonseca and Guedes Soares (1998) \cite{fonsecaTimeDomainAnalysisLargeAmplitude1998} improved the method to support more nonlinear effects and large amplitude motions. Varela and Guedes Soares (2011) \cite{varelaInteractiveSimulationShip2011} have used a table of pre-calculated RAO, which are later interpolated to achieve an interactive frame rate.
However, to our knowledge, this type of method is either too expensive for real-time simulation or not general enough because it requires some physical pre-computation based on CFD or experimental data.

Finally, another attempt to achieve real-time performance is to approximate the forces induced by the fluid on the solid on a mesh.
Basically this is a less accurate method, but it's more efficient and general as it avoids significant pre-calculation.
Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} in his thesis proposed a set of three forces: buoyancy, drag and lift.
Kellomaki (2014) \cite{kellomakiRigidBodyInteraction2014} focused only on buoyancy and drag forces.
Casas-Yrurzum  \etal \cite{casas-yrurzumRealtimePhysicsSimulation2012a} simulates the motion of speedboats by decomposing the solid as a set of small cubes.
Kerner (2016) \cite{kernerWaterInteractionModel2016} wrote a full article on boat physics for video games, adding more forces to simulate slamming and linear damping.
\Framework relies on these forces approximation, aggregating these four last papers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Solid to fluid coupling}
\label{subsubsec:solidToFluidCoupling}
The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the FDM to simulate waves on the free surface.
This method was improved by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
Canabal  \etal (2016) \cite{canabalDispersionKernelsWater2016} improved the physical accuracy of the convolutional method by introducing a dispersion kernel.
Jeschke and Woktan (2017) \cite{jeschkeWaterWavePackets2017} took the work of C. Yuskel a bit further by introducing the notion of a wave packet, which carries information about an entire wave train.
One year later, Jeschke  \etal (2018) \cite{jeschkeWaterSurfaceWavelets2018} proposed to use a wavelet transformation that discretizes the wave amplitudes as a function of space, frequency and direction.
Finally, Schreck  \etal (2019) \cite{schreckFundamentalSolutionsWater2019} based their work on the method of fundamental solutions to generate ambient waves on a large scale domain interacting with complex boundaries.

For solid-fluid coupling, \framework based its calculations on the finite-difference method (FDM) described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
%-------------------------------------------------------------------------
\section{Ocean waves}
\label{sec:oceanWaves}
This section summarizes Tessendorf's method and the various improvements it has received.
It also presents our first contributions: calculation of the fluid velocity at any depth.

Let us introduce some notations as follows:
\begin{itemize}
	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water. \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{\cf appendix \ref{sec:annexe} for more details on this assumption.}.
	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x z\right]^T$ and at time $t$.
	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The vector $\phi(\mathbf x,y,t)\in\mathbb R^3$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
\end{itemize}

Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{\cf appendix \ref{sec:annexe-NS}.}:
\begin{equation}
	\label{equa:systemTessendorf}
	\left\lbrace
	\begin{array}{ll}
		\frac{\partial\phi}{\partial t} = -gh(\mathbf x,t)            & \text{for }y=0,            \\
		\Delta\phi=0                                                  & \text{for }-H\leq y\leq 0, \\
		\frac{\partial h}{\partial t}=\frac{\partial\phi}{\partial y} & \text{for }y=0,            \\
		\frac{\partial\phi}{\partial y}=0                             & \text{for }y=-H.
	\end{array}
	\right.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(t, \mathbf{k}) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being an integer constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(t, \mathbf{k}) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, \cf appendix \ref{sec:annexe} for more details on this assumption.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraph \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{JONSWAP frequency spectrum}
\label{subsubsec:jonswap}

\Framework uses the frequency spectrum developed by Hasselmann \etal (1973) \cite{k.hasselmannMeasurementsWindwaveGrowth1973} in the Joint North Sea Observation Project (JONSWAP).
This spectrum was developed for deep sea water, and hence is in line with our deep water hypothesis.
It is expressed as follows:
\begin{equation}
	S(\omega)=\frac{\alpha g^2}{\omega^5} \exp\left(-\frac{5}{4}\left(\frac{\omega_p}{\omega}\right)^4\right)\gamma^r
\end{equation}
where:
\begin{itemize}
	\item The constant $\alpha=0.076\left(\frac{U^2}{F g}\right)^{0.22}$, $U$ being the wind speed in $m.s^{-1}$ and $F$ the fetch\footnote{The fetch is the length of the area over which the wind is acting on the water, that is to say the distance from a lee shore.} in $m$.
	\item The constant $\omega_p=22\left(\frac{g^2}{U F}\right)$.
	\item The constant $\gamma=3.3$.
	\item The exponent term $r$ is defined as follows:
	      \begin{equation}
		      r=\exp\left(-\frac{\left(\omega-\omega_p\right)^2}{2 \sigma^2 \omega_p^2}\right)
	      \end{equation}
	\item The term $\sigma$ is as follows:
	      \[
		      \sigma=\begin{cases}
			      0.07 & \omega \leq \omega_p, \\
			      0.09 & \omega>\omega_p.
		      \end{cases}
	      \]
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Custom Donelan-Banner spectrum}
\label{subsubsec:donelanbanner}

In \framework, the directional spectrum relies on the work of Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, that mixes together the Donelan-Banner spectrum to control the direction of the waves and a "swell parameter".
Let us recall that, if the wind direction is given by $\theta_0$, the wind direction bias $\theta$ can be defined as follows:
\begin{equation}
	\theta = atan2(k_z, k_x) - \theta_0
\end{equation}
Moreover, the Donelan-Banner $D_{DB}$ spectrum is given as follows:
\begin{equation}
	D_{DB}\left(\omega,\theta\right)=\frac{1}{2}Q_{DB}\left(\omega\right)\beta_s\left(\operatorname{sech}\left(\beta_s\theta\right)\right)^2.
	\label{equa:ddb}
\end{equation}

This expression depends on the ratio $r_{\omega}=\omega/\omega_p$ of $\omega$ and $\omega_p$, the latter being the same as in JONSWAP spectrum.
Hence, in \framework $\beta_s$ is defined as a mix of Donelan \etal (1985) \cite{donelanDirectionalSpectraWindgenerated1985} for $r_{\omega}<1.6$, and Banner (1990) \cite{bannerEquilibriumSpectraWind1990} for $r_{\omega}\geq1.6$.
Moreover, the case $r_{\omega}<0.56$ is removed, as suggested by C. Horvath.
Altogether, this leads to the following expression of $\beta_s$:
\begin{equation}
	\begin{aligned}
		 & \beta_s= \begin{cases}%
			            2.61\left(r_{\omega}\right)^{1.3}  & \text{for } r_{\omega}<0.95          \\
			            2.28\left(r_{\omega}\right)^{-1.3} & \text{for } 0.95 \leq r_{\omega}<1.6 \\
			            10^\epsilon                        & \text{for }  r_{\omega}\geq1.6       %
		            \end{cases}
	\end{aligned}
\end{equation}
where
\begin{equation}
	\epsilon=0.8393 \exp \left(-0.567\ln\left(r_{\omega}^2\right)\right)-0.4.
\end{equation}
Any directional spectrum $D(\omega,\theta)$ must satisfy the normalization condition:
\begin{equation}
	\label{equa:normalization}
	\int_{-\pi}^\pi D(\omega,\theta)d\theta = 1
\end{equation}
Hence, $Q_{DB}(\omega)$ is a normalization factor to make $D_{DB}(\omega, \theta)$ respect the condition \ref{equa:normalization}:
\begin{equation}
	\begin{aligned}
		Q_{DB}(\omega) & = \frac{1}{\displaystyle\int_{-\pi}^\pi \frac{1}{2}\beta_s\left(\operatorname{sech}\left(\beta_s \theta\right)\right)^2 d\theta} = \frac{1}{\tanh\left(\beta_s\pi\right)}
	\end{aligned}
\end{equation}

\Framework simulates swell, corresponding to more elongated waves that have traveled out of their generating area and that mixes with wind-generated local waves.
As in Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, it relies on the swell parameter $\xi$ related to the spectrum $D_{\xi}$, defined as follows:
\begin{equation}
	D_{\xi}(\omega, \theta)=|\cos (\theta / 2)|^{2 s_{\xi}}
\end{equation}
where the spreading swell function $s_\xi$ is defined by:
\begin{equation}
	s_\xi = 16 \tanh(1 / r_\omega) \xi^2
\end{equation}
Note that, with $\xi=0$, the spectrum $D_{0}(\omega, \theta)=\frac{1}{2\pi}$ is constant while for $\xi=1$ the waves become much more elongated.

The product between the Donelan-Banner and $\xi$ spectra is denoted by $D_{DB\xi}$, and is expressed as follows:
\begin{equation}
	D_{DB\xi}(\omega, \theta) =Q_{DB\xi}(\omega) D_{DB}(\omega, \theta) D_{\xi}(\omega, \theta)
\end{equation}
where the normalization factor is:
\begin{equation}
	Q_{DB\xi}(\omega)=\frac{1}{\displaystyle\int_{-\pi}^\pi D_{DB}(\omega, \theta) D_{\xi}(\omega, \theta) \mathrm{d} \theta}.
\end{equation}
This normalization factor $Q_{DB\xi}(\omega)$ is computationally expensive.

It can be noticed that it depends only on the ratio $r_{\omega}$.
Hence, \framework uses a simple method to approximate it, through interpolations relying on multiple Lagrangian polynomials as follows:
\begin{equation}
	\widetilde{Q_{DB\xi}}(r_{\omega})=
	\begin{cases}
		7.1467551 r_{\omega}^2 - 13.4662001 r_{\omega} + 7.75651088    & \text{for }r_{\omega} < 0.94,        \\
		- 0.69906109 r_{\omega}^2 + 0.77975933 r_{\omega} + 0.10169164 & \text{for }0.94 \leq r_{\omega} < 5, \\
		- 2.1860997 r_{\omega}^2 +0.0269209 r_{\omega} +0.00016283     & \text{for } 5 \leq r_{\omega} < 100, \\
		1.2038847 r_{\omega} + 0.0008147                               & \text{else.}
	\end{cases}
\end{equation}

Computed using $10000$ random samples, the normalization integral (Equation \ref{equa:normalization}) of such an approximation is $1.0699$ in average.

Note that we applied a first normalization factor $Q_{DB}$ to the simple Donelan Banner spectrum \ref{equa:ddb} before calculating this second normalization factor $\widetilde{Q_{DB\xi}}(r_{omega})$. In fact, if we simply approximate the full-spectrum normalization factor before applying it, we'll be faced with a normalization problem, because for small values of $r_{omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.

Finally, \framework uses a last parameter $\delta\in[0,1]$ that represents how much the waves orientation is constant ($\delta = 0$) or not ($\delta = 1$) (\cf Figure \ref{fig:dispersionEffect}).
Its meaning is as follows:
\begin{itemize}
	\item $\delta = 0$ indicates a \emph{neutral} directional spectrum, that has no consequence on $\tilde{h}_0$;
	\item $\delta=1$ indicates a spectrum equals to $D_{DB\xi}(\omega, \theta)$.
\end{itemize}
This gives the final expression of the custom Donelan-Banner spectrum used in \framework:
\begin{equation}
	D(\omega, \theta)=\left(1-\delta\right)\frac{1}{2\pi}+\delta D_{DB\xi}(\omega,\theta).
\end{equation}

\begin{figure}[htbp]
	\centering

	\subfloat[Using $\delta = 0$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/seaDispersion0SeaState5-contrast.png}
		\label{fig:dispersion0}
	}
	\hspace{1cm}
	\subfloat[Using $\delta = 1$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/seaDispersion1SeaState5-contrast.png}
		\label{fig:dispersion1}
	}
	\caption{Effect of the dispersion parameter on sea state $5$ on the Beaufort scale.}
	\label{fig:dispersionEffect}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water displacement and derivative}

\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
As proposed in Sections 4.5 and 4.6 of Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999}, it is calculated as follows:
\begin{equation}\label{equa:hori-displacement}
	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(t, \mathbf{k})\exp\left(i\mathbf{k}\cdot\mathbf x\right).
\end{equation}

Furthermore, to obtain additional details on the sea surface (\emph{e.g.} foam positions, normal vectors...) the derivative of the horizontal and vertical displacements must be computed, respectively of $\mathbf{D}(\mathbf x,t)$ and $h(\mathbf x,t)$.
Taking advantage of the symmetry $\frac{\partial D_z}{\partial x}(\mathbf x,t) = \frac{\partial D_x}{\partial z}(\mathbf x,t)$, this relies on five derivatives only:

\begin{align*}
	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right).
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cascades}
\label{subsec:cascades}
While the expression of the water height in Equation \ref{equa:heightWater} is convenient, it may result in an ocean surface that appears overly repetitive.
A naive solution to handle this problem consists in using a huge resolution $N$, but it has a major impact on performance.
\Framework uses the more efficient solution proposed by Dupuy and Bruneton (2012) \cite{dupuyRealtimeAnimationRendering2012}.
It consists in dividing the spectrum into $c$ layers or \emph{cascades} of different length.
Each of these layers represents a different wavelength in Equation~\ref{equa:heightWater}.
The length of the $i$-th cascade is denoted by $L_i$.
These lengths $L_i$ are parameters in the expression of the wave vector component described in Equation~\ref{equa:heightWater}.
Moreover, the different cascades are cut off at arbitrary value to avoid overlap between them, as illustrated in Figure~\ref{fig:cascades}.

\Framework uses by default $3$ cascades with the following settings:
\begin{itemize}
	\item $L_0 = 4$ operating for waves length $k\in [0,\frac{12\pi}{16}]$
	\item $L_1 = 16$ operating for waves length $k\in [\frac{12\pi}{16},\frac{12\pi}{256}]$
	\item $L_2 = 256$ operating for waves length $k\in [\frac{12\pi}{256},+\infty]$.
\end{itemize}

\begin{figure}
	\centering
	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
	\input{Tikz/cascades.tikz}
	\caption{Example of a spectrum split into $3$ different cascades: each cascade operates on a different interval of wave vector length.}
	\label{fig:cascades}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{\cf appendix \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

To our knowledge, this expression of the velocity has never been used in a real time ocean simulator before.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Waves' height implementation details}
This section gives some technical details of the implementation of an ocean with Tessendorf's method in \framework.

First, it should be noticed that the expressions of the ocean surface's height (Equation~\ref{equa:heightWater}), the displacements (Equation~\ref{equa:hori-displacement}) and their five derivatives, and the velocity (Equation~\ref{equa:velocity}) have "a form of an inverse discrete Fourier transform".
More precisely, the \textbf{frequency space} coefficients appearing before the term $\exp(i\mathbf{k}\cdot\mathbf x)$ are stored in a square matrix of size $N\times N$, for each expression.
Then, the full sum is computed through an \textbf{inverse fast Fourier transform} (IFFT) applied on this coefficients' matrix.
In \framework, this trick allows computing these eight expressions without any significant impact on performance, even for huge domains.

Second, as \framework simulates a tile of water of size $N\times N$ using $c$ cascades, it stores the height, the displacement and so on in an 1D array of size $c$ containing a vector of 2D arrays of size $N\times N$.

Third, many terms are independent on the time $t$: $\tilde{h}_0(\mathbf{k})$, $\tilde{h}_0^*(-\mathbf{k})$ and the wave data ($\mathbf{k}$ and $\omega(k)$).
\Framework computes them for each cascade wave vector at each spectrum change only, but not at each iteration.
Then it does a combination with time-dependent terms at each iteration, giving an expression for the coefficients in frequency space for the eight IFFT.

Last but not least, one might think that applying IFFT should be enough to get the expected result.
Nevertheless, using the standard IFFT algorithm (such as the one of Cooley and Tukey (1965) \cite{cooleyAlgorithmMachineCalculation1965}) leads to some incorrect geometry changes.
Unlike usual IFFT, the used range is $\left[-N/2, N/2\right]$ and so the wave vectors have negative components (\emph{e.g.} the height).
To fit with the theory of Fourier transforms, in \framework each value at index $i,j$ of the 2D matrices is multiplied by $(-1)^{i+j}$.

In \framework, the number of IFFT applications is divided by two using Theorem \ref{theorem:ifft} on Hermitian matrices\footnote{A matrix $X=(x_{n,m})$ is Hermitian if it is equal to its own conjugate transpose: $X = X^{\dagger}$ or $x_{n,m}=x^*_{m,n}$.} (\cf annex \ref{sec:annexe-2IFFT} for its demonstration).
\begin{theorem}
	\label{theorem:ifft}
	Let: $X=(x_{n,m}), Y=(y_{n,m}) \in M_n(\mathbb{C})$ be two Hermitian matrices.
	Then the following relation is valid:
	\begin{equation}
		\mathcal{F}^{-1}\left(X+iY\right) =
		\Re\left(\mathcal{F}^{-1}\left(X\right)\right)
		+ i\Re\left(\mathcal{F}^{-1}\left(Y\right)\right)
	\end{equation}
\end{theorem}
Hence, in \framework the coefficients in frequency space of two Hermitian matrices $X$ and $Y$ are combined in a single matrix $H=X+iY$.
Using Theorem \ref{theorem:ifft}, one complex IFFT suffices to obtain the real parts of two IFFT $\mathcal{F}^{-1}\left(X\right)$ and $\mathcal{F}^{-1}\left(Y\right)$.
This treatment allows reducing IFFT computation times by a factor close to 2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Waves' velocity implementation details}
Contrary to the displacement and its derivative which are height maps and 2D functions, the velocity must be known in 3D space.
Hence, its calculation is more difficult and requires more computation time.
Computing the velocity for all the points of a 3D grid, or even only the points where it must be evaluated for fluid to solid coupling (Section~\ref{subsubsec:fluidToSolidCoupling}) is not possible in real-time, as it involves a large to huge number of IFFT.
Unfortunately, the dependency of the attenuation (Equation \ref{equa:attenuationDeep}) on both the wave vectors and $y$ makes the depth dependency of the velocity fundamentally tied to the wave vector, preventing from any simplifications to reduce the needed calculations.

To reduce the complexity of this problem, \framework opts for an interpolation scheme, involving a reasonable number of IFFTs that is proportional to the size of the discretization.
% More precisely, in \framework the velocity is obtained using around 10 IFFTs only.

More concretely, the velocity must be computed at any horizontal position in $I=[y_{min}, y_{max}]$.
In \framework, $N$ growing depths $y_i$ are chosen in this interval $I$.
For each of these depths $y_i$, the $i$-th velocity $\mathbf{v_i}(\mathbf x,y_i,t)$ is computed using one IFFT, leading to $N$ IFFTs for all the sampled depths $y_i$.
Then, the velocity $\mathbf{v}(\mathbf x,y,t)$ at any depth $y\in I$ is computed using an ad hoc interpolation from the different $\mathbf{v_i}$.

A first way to do this interpolation mechanism relies on a linear scheme.
Nevertheless, it produces a very inaccurate result.
Hence, \framework uses an exponential scheme made on logarithmic sampling of the interval $I$.
They are discussed in the next two sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Exponential interpolation}
The vertical dependency of velocity links to the exponential attenuation function from Equation \ref{equa:attenuationDeep}.
Hence, instead of a linear interpolation \framework uses an exponential interpolation $f_e(x)$ that should resemble the following expression:
\begin{equation}
	f_e(x)=\alpha \exp(\beta x).
\end{equation}
Knowing $f_e(a)=f(a)$ and $f_e(b)=f(b)$, it is straightforward to obtain the two values $\alpha$ and $\beta$ as follows:
\begin{equation}
	\begin{cases}
		\displaystyle\alpha = \frac{f(a)}{\exp(\beta a)} \\
		\displaystyle\beta = \frac{\ln(|f(b)|) - \ln(|f(a)|)}{b - a}
	\end{cases}
\end{equation}
This exponential interpolation function $f_e(x)$ is applied to extract the magnitude and rotation angle of the velocity $\mathbf v(\mathbf x,y,t)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Logarithmic distribution}
As the velocity decays exponentially with depth, the ones close to the ocean surface play a more important role in accuracy than the one at the bottom of the ocean.
Hence, a regular sampling along the depth seems inappropriate, albeit the exponential interpolation.

\Framework uses the following logarithmic distribution:
\begin{equation}
	d(y_i) =
	\begin{cases}
		\beta \ln(\alpha|y_i|^2+1) ~~\text{if}~y_i>0     \\
		-\beta \ln(\alpha|y_i|^2+1) ~~\text{if}~y_i\leq0 \\
	\end{cases}
\end{equation}
where $\alpha=0.0001$ and $\beta = -y_{min}/\left(2 \ln(\alpha |y_{min}|^2 + 1)\right)$.

This distribution is designed based on the following considerations.
First, while the velocity decays exponentially, it relies on a logarithmic similar to $y_i\mapsto \ln(|y_i|+1)$.
Since this function transforms the negative $y_i$ into a positive one, the result is multiplied by the sign of $y_i$.
Also, this function decreases slowly (actually with $\ln(250+1)=5.5$), hence the logarithmic distribution uses the square of $|y_1|$.
Lastly, the parameter $\alpha$ controls the "squeezing" in the center, while the parameter $\beta$ controls the deepest value of $d$.
Notice that $\beta$ is chosen as the solution of $d(y_{min})=\frac{y_{min}}{2}$ but not at $y_{min}$: it is the solution giving the best accuracy after multiple tries.

Figure \ref{fig:comparisonGrid} compares the logarithmic discretization to the uniform discretization. As expected, the behavior with the former leads to fewer points close to $y_{min}$ and more points close to $y_{max}$.
Note also the higher density of points close to $y=0$.

\begin{figure}[ht]
	\centering
	\subfloat{\includegraphics[width=0.7\textwidth]{Figures/gridInterpolationDegree8.png}\label{fig:gridInterpolationDegee8}}
	\hfill
	\subfloat{\includegraphics[width=0.3\textwidth]{Figures/gridInterpolationDegree8Legend.png}\label{fig:gridInterpolationDegree8Legend}}
	\caption{This figure shows a comparison between a \textcolor{bleuF}{uniform discretization in blue} and a \textcolor{redF}{logarithmic discretization in red}. The dotted parts represent the last interpolation toward a null velocity at $y_{min}$.}
	\label{fig:comparisonGrid}
\end{figure}

Moreover, using the logarithmic distribution leads to $d(y_{min})=\frac{y_{min}}{2}$.
Assuming that the velocity at $y_{min}$ is null, then for any depth between $y_{min}$ and $\frac{y_{min}}{2}$ the velocity can be computed between the one at $\frac{y_{min}}{2}$ and $0$.

Figure \ref{fig:interpolationAccuracyComparison} compares the accuracy using linear mechanism and the exponential interpolation with the logarithmic discretization.
Clearly the latter gives much better accuracy, as expected.

\begin{figure}[ht]
	\centering
	\subfloat{\includegraphics[width=0.5\textwidth]{Figures/interpolationAccuracyComparison.png}\label{subfig:interpolationAccuracyComparison}}
	\hfill
	\subfloat{\includegraphics[width=0.5\textwidth]{Figures/interpolationAccuracyComparisonZoom.png}\label{subfig:interpolationAccuracyComparisonZoom}}
	\caption{The left figure represents the accuracy of different type of interpolation in function of the size of the discretization $N$. The right one does a zoom on small velocities between $0$ and $0.1$.
		They use 10000 positions from a continuous uniform distribution in the parallelepiped $[-1000,1000]\times[-125,4.5]\times[-1000, 1000]$ and with a wind speed of 20 $m.s^{-1}$.
		The interpolated velocity $\tilde{\mathbf{v}}$ and the analytical velocity $\mathbf{v}$ are evaluated.
		The accuracy is defined by the mean of all positions of $\big|||\tilde{\mathbf{v}}||-||\mathbf{v}||\big|$.
		It can be noticed that considering a uniform discretization, then exponential interpolation has a better accuracy than linear one.
		Logarithmic discretization is much better than uniform discretization, whatever the used interpolation.
		At least, exponential interpolation gives slightly better result than linear interpolation, even for low $N$.}
	\label{fig:interpolationAccuracyComparison}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Choosing the interpolation degree}

Knowing the interpolation method, the interpolation degree (number of points in the table) must be decided.
Obviously, using more points results in better accuracy, but also more computation time.
Then, a trade-off between performance and accuracy must be determined.
With this aim, the \textit{multi-objective optimization} method is used.

Briefly, this method reduces a multi-criteria optimization problem into a one dimensional one.
It defines a one dimensional function as a linear sum of different objective functions; here, it concerns the performance and accuracy for a given interpolation degree.

Hence, this optimization function $J$ is defined as follows:
\begin{equation}
	J(d) = \alpha P(d) + \beta A(d)
\end{equation}
where $d$ is the interpolation degree, $P(d)$ measures the performance of the velocity interpolation, $A(d)$ measures the accuracy of the velocity interpolation, and $\alpha$ and $\beta$ are the weight arbitrary assigned to the performances and the accuracy functions.

To make $P(d)$ and $A(d)$ having the same order of magnitude, the units are the \textbf{decisecond} for performance and the \textbf{meter per seconds} for accuracy (\cf figure \ref{fig:interpolationCriteria}).
Moreover, the performance being the more important for real-time process, the following arbitrary choice are made for the coefficients of $J$:
\begin{equation}
	\begin{cases}
		\alpha = 1 \\
		\beta = 2
	\end{cases}
\end{equation}

Figure \ref{fig:multiObjectiveFunction} shows the function $J$ using these parameters. It appears that its minimum is obtained with $d=8$.
It is the value used in \framework.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/interpolationMeanPerformancesAndAccuracy.png}
	\caption{These figures represent the different objective for choosing the interpolation degree. In blue, the mean accuracy of velocity interpolation in meters per seconds, we denote this function $A(d)$. In red, the mean performances of velocity interpolation in deciseconds, we denote this function $P(d)$. We estimate this function with $10000$ sample.}
	\label{fig:interpolationCriteria}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/multiObjectiveAccuracy1Duration2.png}
	\caption{Multi-objective function $P$ to optimize, with two objectives: performance and accuracy.
		Its minimum is reached by the interpolation degree of $8$. For more details on each objective, \cf Figures \ref{fig:interpolationCriteria} which represent them individually.}
	\label{fig:multiObjectiveFunction}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Coupling solid and fluid}
\label{sec:couplingSolidAndFluid}
The physical simulation based on the Tessendorf's method is very fast, and the spectrum approach leaves a lot of control to the user.
However, it doesn't simulate the interaction between the fluid and a solid, which is needed for adding some boats in the simulation.
To overcome this flaw, \framework uses two methods: a first one for fluid-to-solid action, and a second one for solid-to-fluid action.

These two methods are designed with the following constraints in mind:
\begin{enumerate}
	\item \label{enum:simpleParam} Rely on simple parameterization for non-physicists, to be easily usable by artists;
	\item \label{enum:noPrecal} Not requiring too many pre-calculations, to avoid an overly complex framework;
	\item \label{enum:perf} Have excellent performance (a few milliseconds at most), so that \framework reaches real time globally, including all simulation and rendering stages.
\end{enumerate}

%-------------------------------------------------------------------------
\section{Fluid-to-solid action}
\label{sec:fluidToSolidAction}
The action of the fluid on a given solid can be approximated by a \emph{set of forces} acting continuously on the \emph{contact surface} between the fluid and the solid.
In \framework, the contact surface is represented using the triangulation of the solid's mesh, leading to a discrete approximation of the set of forces.
Moreover, to remain real-time \framework uses forces computed from simple geometric and physical parameters only.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geometrical Parameters}
\label{subsec:geomParam}
In \framework, the forces are computed for each boat or body, independently of the others.
The calculation of the forces needs the following geometrical data from a body and the water surface:
the lower hull (part of the hull in the water), the waterline, and the upper hull (part of the hull above the water).

At first, \framework checks the vertical position of body's each vertex relative to the water surface.
Then, it assigns a status to each body's face: submerged, partially submerged or not submerged.
The triangles that are partially submerged are divided as done in Kerner (2016) \cite{kernerWaterInteractionModel2016}, producing the waterline as the intersection edges between the mesh and the water surface.
This waterline is used later to compute the mask in Section~\ref{subsec:mask}.
Then the body surface is decomposed into a set $\left\lbrace T_i^s\right\rbrace$ of submerged triangles, and $\left\lbrace T_i^{\neg s}\right\rbrace$ of not submerged ones.
\Framework computes the area of each submerged or not submerged triangle at this step too.

The buoyancy calculation presented in Section \ref{subsubsec:buoyancy} requires the full submerged volume $v_w$ of the mesh.
\Framework uses the prism approximation proposed by Bajo \etal (2020) \cite{bajoRealisticBuoyancyModel2020a}.
The principles of their method are as follows: from each submerged triangle $T_i^s$ of the body, a prism is build up to the water surface; hence, each prism volume can be calculated, and so the full submerged volume $v_w$ by adding all of them.

At last, from all the submerged faces $T_i^s$ \framework computes the center of immersion $c_i$ used for the buoyancy calculation as presented in Section~\ref{subsec:forces}.

%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Forces}
\label{subsec:forces}
\Framework uses three forces as in Kellomaki (2014) \cite{kellomakiRigidBodyInteraction2014}:
\begin{enumerate}
	\item Buoyancy $\mathbf{F}_b$;
	\item Water drag $\mathbf{F}_w$;
	\item Air drag $\mathbf{F}_a$.
\end{enumerate}
These forces are computed for each triangle of the lower or upper hull, and so after the different geometric parameters being calculated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Buoyancy}
\label{subsubsec:buoyancy}
The buoyancy is computed once for the whole boat from the following analytical formula:
\begin{equation}
	\mathbf{F_b} = - v_w \rho_w \mathbf g,
\end{equation}
where $v_w$ is the submerged volume in $m^3$ (cf. Section~\ref{subsec:geomParam}) and $\rho_w$ is the water density.
The water density is obtained at any depth by linear interpolation using the measure of the International Towing Tank Conference (ITTC) \cite{ittcFreshWaterSeawater2011}.
The buoyancy is applied to the center of immersion $c_i$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Water drag}
\label{subsubsec:waterDrag}
\Framework uses a drag force depending on the relative speed, as in many previous works \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014,kernerWaterInteractionModel2016}.
The water drag force is calculated for each triangle $T_i^s$ of the submerged body as follows:
\begin{equation}
	\mathbf{F_w}= -\frac{1}{2} C_d^w \rho_w A_i^{\perp} \left\Vert\mathbf{v_{i_{rel}}^w}\right\Vert\mathbf{v_{i_{rel}}^w},
\end{equation}
where:
\begin{itemize}
	\item $\mathbf{v_{i_{rel}}^w}=\mathbf{v_i}-\mathbf{v_i^w}$ is the relative velocity in $m.s^{-1}$ of the triangle $T_i^s$, equals to the difference between its velocity $\mathbf{v_i}$ and the velocity $\mathbf{v_{i}^w}$ of the water at its center;
	\item $A_i^{\perp}$ is the projected area in $m^2$ of the triangle $T_i^s$ with respect to the relative velocity unit direction (\ie the vector $\mathbf{v_{i_{rel}}^w}/\left\Vert\mathbf{v_{i_{rel}}^w}\right\Vert$);
	\item $\rho_w$ is still the density of water in $kg.m^{-3}$;
	\item $C_d^w\in\mathbb{R}^3$ is the drag coefficient of the body in the water, and is given by the user.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Air drag}
\label{subsubsec:airDrag}
The force corresponding to the air drag is designed similarly to the water drag one, but using the not submerged part $\left\lbrace T_i^{\neg s}\right\rbrace$ of the body.
For each triangle $T_i^{\neg s}$, it corresponds to the following expression:
\begin{equation}
	\mathbf{F_a}= -\frac{1}{2} C_d^a\rho_a A_i^{\perp}\left\Vert\mathbf{v_{i_{rel}}^a}\right\Vert \mathbf{v_{i_{rel}}^a}
\end{equation}
where:
\begin{itemize}
	\item $\mathbf{v_{i_{rel}}^a}=\mathbf{v_i}-\mathbf{v_{i}^a}$ is the relative velocity in $m.s^{-1}$ of the triangle $T_i^{\neg s}$, equals to the difference between its velocity $\mathbf{v_i}$ and the velocity $\mathbf{v_{i}^a}$ of the air at its center;
	\item $A_i^{\perp}$ is the projected area in $m^2$ of the triangle $T_i^{\neg s}$ with respect to the relative velocity unit direction (\ie the vector $\mathbf{v_{i_{rel}}^a}/\left\Vert\mathbf{v_{i_{rel}}^a}\right\Vert$);
	\item $\rho_a$ is the density of air\footnote{According to the international standard atmosphere, the air density at sea level for $20$°C is $1.204~kg.m^{-3}$ cf. \url{https://www.digitaldutch.com/atmoscalc/table.htm}} in $kg.m^{-3}$;
	\item $\rho_a$ is the density of air in $kg.m^{-3}$;
	\item $C_d^a\in\mathbb{R}^3$ is the drag coefficient of the body in the air, and is given by the user.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fluid-to-solid algorithm}
\label{subsec:algoForces}
At each iteration, the fluid-to-solid algorithm proceeds as follows:
\begin{enumerate}
	\item Computing the vertices' position of the body relative to the water height (Section \ref{subsec:geomParam});
	\item Computing the sets of submerged and not submerged triangles, by splitting the triangles having vertices in above and below the water surface (Section \ref{subsec:geomParam});
	\item Computing the buoyancy force for each body (Section \ref{subsubsec:buoyancy});
	\item Computing the two drag forces for each triangle, and summing them for each body (Sections \ref{subsubsec:waterDrag} and \ref{subsubsec:airDrag});
	\item Applying the forces in the physical integration scheme.
\end{enumerate}

A summary of the forces that are applied on solid can be found on figure \ref{fig:forcesComputation}.

\begin{figure}[!h]
	\centering
	\input{Tikz/forcesComputation.tikz}
	\caption{Summary of forces calculation. Using the same notation as Section~\ref{subsec:forces} it follows:
		the sum of the buoyancy forces \textcolor{jauneC}{$F_b$} in \textcolor{jauneC}{yellow}, applied to the center of immersion $c_i$;
		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to $c_i$ too;
		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the submerged part;
		and finally the gravity force \textcolor{greenF}{$F_g$} in \textcolor{greenF}{green}, applied to the center of gravity.
		Different velocities are displayed above the ship:
		the wind velocity \textcolor{redF}{$v_a$} in \textcolor{redF}{red};
		the water velocity \textcolor{bleuF}{$v_w$} in \textcolor{bleuF}{blue};
		and the object velocity \textcolor{noirC}{$v$} in \textcolor{noirC}{black}.}
	\label{fig:forcesComputation}
\end{figure}

%-------------------------------------------------------------------------
\section{Solid-to-fluid action}
\label{sec:solidToFluidAction}
\Framework relies on the method developed by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to simulate the action of the solid on the fluid.
The simulation of this action is 3-dimensional by nature.
However, it has to be simplified to satisfy the third requirement of Section~\ref{sec:couplingSolidAndFluid}, concerning computation time performance.
As for fluid simulation in Section~\ref{sec:oceanWaves}, \framework assumes that the interaction from a solid to the fluid can be simulated as a 2D height field.
Furthermore, it considers a single frequency for the waves generated by a solid $M$.
These specific waves are called interactive waves.
These approximations may seem too strong, but as mentioned above, they are necessary to satisfy the requirements defined in Section~\ref{sec:couplingSolidAndFluid}.

Hence, \framework assumes that the interactive waves are solutions of the following classical 2-dimensional wave equation:
\begin{equation}
	\label{eq:2dwave}
	\nabla h(\mathbf x,t)-\frac{1}{c^2}\frac{\partial^2 h(\mathbf x,t)}{\partial t^2}=0
\end{equation}
where $h(\mathbf x,t)$ is the height of the water surface at horizontal position $\mathbf x$ and time $t$, $\nabla=\frac{\partial^2}{\partial x^2}+\frac{\partial^2}{\partial z^2}$ is the Laplacian in 2D and $c$ is the wave velocity expressed in $m.s^{-1}$.
\Framework uses the following Dirichlet boundary condition:
\begin{equation}
	h(\mathbf x,t) = 0 \text{~~~on the boundary.}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Finite-difference method}
\label{subsec:finiteDifferenceMethod}
Following Gomez (2000) \cite{gomezInteractiveSimulationWater2000},
\framework  uses the FDM with an explicit scheme to numerically solve the equation (\ref{eq:2dwave}).
It uses a square simulation zone $Z$ of dimension $L\times L$ around a given mesh $M$.
This zone is regularly discretized with a $dx$ step.
Time is also regularly discretized with a time step $dt$.

In the FDM simulation, \framework denotes by $h_{i,j}^n$ the water height $h(i\times dx,j\times dx,n \times dt)$ for discrete location $(i,j)\times dx$ and time $n\times dt$.
This leads to the following explicit scheme:
\begin{equation}
	\label{eq:fdm_scheme}
	h_{i,j}^{n+1} = a\left(h_{i+1,j}^n+h_{i-1,j}^n+h_{i,j+1}^n+h_{i,j-1}^n-4h_{i,j}^n\right) + 2h_{i,j}^n - h_{i,j}^{n-1}
\end{equation}
where
\begin{equation}
	a=\frac{c^2dt^2}{dx^2}.
\end{equation}
The Dirichlet boundary condition gives $h_{i,j}^n=0$ for $i$ or $j$ being on the boundary of $Z$.
Since a new wave is generated at each time step, a damping effect is added to make them disappear gradually.
Hence, the right part of Equation (\ref{eq:fdm_scheme}) is multiplied by the \textbf{damping factor} $d(t)=d^n$.
This damping factor is linearly related to the velocity of the mesh $v_M(t)$ to ensure numerical stability as follows:
\begin{equation}
	d(t) =
	\begin{cases}
		\left(1-\frac{||v_M(t)||}{v_{max}}\right)d_0 + \frac{||v_M(t)||}{v_{max}}d_{max} & \text{for~~}||v_M(t)||\leq v_{max}, \\
		d_{max}                                                                          & \text{else}
	\end{cases}
\end{equation}
where $d_{max}$ is the maximum damping value obtained when the body's velocity is above a given threshold $v_{max}$,
and $d_0$ is the damping for a null velocity.
For a plausible appearance, \framework uses the following values by default:
\begin{equation}
	\begin{cases}
		d_0     & = 0.98,       \\
		d_{max} & = 0.999,      \\
		v_{max} & = 5 m.s^{-1}.
	\end{cases}
\end{equation}
Nevertheless, the user can adjust these values in a way that a body $M$ causes more or less interaction waves.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Grid translation}
\label{subsec:gridTranslation}
The waves generated by a given body $M$ are defined in its local coordinates system.
Hence, when $M$ moves, the waves have to move accordingly as done by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
A first solution consists of applying the translation to the grid at times $t$ and $t-dt$ and then of using the integration scheme of Equation (\ref{eq:fdm_scheme}).
Nevertheless, these time-consuming shifts can be avoided by using the following domain translation:
\begin{equation}
	\begin{cases}
		k = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{dx}\right\rfloor    \\
		l = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{dx}\right\rfloor    \\
		o = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{dx}\right\rfloor \\
		p = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{dx}\right\rfloor
	\end{cases}
\end{equation}
where $\mathbf p(t)=\left[p_x(t)\ p_z(t)\right]^T$ is the position of the mesh $M$ at time $t$, $k$ and $l$ (respectively $o$ and $p$) are the shifted indices defined by the translation of $M$ at times $t+dt$ from times $t$ (respectively from times $t-dt$).
Then, the integration scheme can be rewritten as follows:
\begin{equation}
	\label{eq:fdm_scheme_corrected}
	h_{i,j}^{n+1} = d^n\left(a\left(h_{k+1,l}^n+h_{k-1,l}^n+h_{k,l+1}^n+h_{k,l-1}^n- 4 h_{k,l}^n\right) + 2 h_{k,l}^n - h_{o,p}^{n-1}\right)
\end{equation}

Since $M$ moves, the shift indexes $k$, $l$, $o,$ and $p$ can be outside the boundary of the simulation zone $Z$.
To solve this problem, the size $m$ of the boundary of $Z$ is increased such that $m>1$ (\cf Figure \ref{fig:examGrid}).
\Framework takes $m=16$ which is large enough for large translations.

% Pattern Info
\begin{figure}[!h]
	\centering
	\input{Tikz/mdfGrid.tikz}
	\caption{An example of zone $Z(t)$ at a time $t$ using boundary size $m=5$.
		In \textcolor{bleuF}{dotted blue} the mesh ${M}$ at time $t$.
		In \textcolor{redF}{red} the same mesh at time $t+dt$.}
	\label{fig:examGrid}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stability}
\label{subsec:stability}
The stability of the scheme defined by Equation (\ref{eq:fdm_scheme_corrected}) is related to the CFL condition (for Courant–Friedrichs–Lewy), here given as follows:
\begin{equation}
	\frac{c^2dt^2}{dx^2}\leq 0.5
\end{equation}
The parameter $dt$ is considered fixed throughout the simulation.
One contribution of this article is to adapt the simulation's parameters $c$ and $dx$ to the CFL, as presented below.

Preliminary tests have shown that $dx$ must be small for a low velocity $\mathbf{v_M}(t)$ of $M$.
Moreover, $dx$ must be large enough for $Z$ to encompass $M$ and the waves generated by $M$.
The size of $M$ is defined from its bounding box $(b_x, b_y, b_z)$ as $S(M)=\max(b_x, b_z)$.
\Framework lets the user define the bounds $dx_{min}$ and $dx_{max}$ so that $Z$ is at least twice as large as $M$.
Then $dx$ is defined as follows:
\begin{equation}
	dx =
	\begin{cases}
		0.999~dt                    & ~~ \text{if}~||\mathbf{v_M}(t)||<1    \\
		||\mathbf{v_M}(t)||0.999~dt & ~~ \text{if}~||\mathbf{v_M}(t)||\geq1
	\end{cases}
\end{equation}
Finally $dx$ is clamped between $dx_{min}$ and $dx_{max}$.
The value of $c$ is deduced from $dx$ and $dt$ to validate the CFL as follows:
\begin{equation}
	c = \sqrt{0.49} \frac{dx}{dt}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Wave generation}
\label{subsec:mask}
While the FDM allows to simulate the propagation of the interactive waves, one more input is required for the wave generation: an alteration of the free surface of the height map.
\Framework uses a new way to compute this input, which provides a satisfying result no matter the form of the body generating the waves.
The input of the FDM is called \textbf{mask}.

The idea behind the mask is to distort the grid of the FDM by reproducing the intersection surface between $M$ and the free surface.
Here, free surface refers to the ambient waves of Tessendorf's method \ref{sec:oceanWaves} plus the interactive waves generated by all solids except $M$.
It does \textbf{not} include the free surface generated by $M$, or the simulation would face a divergence issue.

This intersection has already been computed as part of the geometric parameters in Section~\ref{subsec:geomParam}.
It is given as a list of 3D points forming the intersection polygons of the hull with the free surface.
These points are moved from the world coordinates system to the one of $Z$, the surrounding grid of $M$.

Hence, for a body $M$ the wave generation relies on two different steps:
First, a mask is defined from this list of 3D points, corresponding to the vertices of $Z$ contained into one intersected polygons.
Second, the height $h$ of each point in the mask is determined using an ad hoc function.
These two steps are detailed in the following two sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Calculate the mask}
The FDM needs the height of the vertices below the body.
This is computed from the mask, composed of the vertices of $Z$ contained into one intersection polygon at least.
This is done using the \textbf{Point in Polygon} strategy described in Haines (1994) \cite{hainesPointPolygonStrategies1994} and later in Cormen (2001) \cite{CormenIntroductionAlgorithms2001}.

The principle of this algorithm is as follows:
for each vertex $v$ of the grid $Z$, a ray starting at $v$ and with direction $z$ is generated, and its number of intersection with all the intersection polygons is computed.
$v$ is in the polygon if this number of intersections is odd; otherwise it is not.

To reduce the computation times, this strategy is applied to the vertices $v$ included into the bounding box of the intersections polygons.
Since the number of vertices $v$ of $Z$ is at least twice as the size of $M$, this simple technic reduces efficiently the number of ray to polygon calculations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Evaluate the height of vertices in the mask}
\label{subsubsec:heightMask}
Knowing all the vertices in the mask, their height must be given as input of the FDM.
\Framework uses a simple model that tries to reproduce parts of the idea of Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
By construction of the grid $Z$, $M$ has is bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.
Moreover, the boat forms a $V$ shape, with its center a bit below the extremity of the side.

To simplify, $h(\mathbf x)$ is defined as the sum of two orthogonal functions $f(x)$ and $g(z)$,
where $f(x)$ represent the side effect, and $g(z)$ the front/back effect.

%%%%%%%%%%%%%%
The function $f$ models the side effect using a shape in $V$.
It is defined as follows:
\begin{equation}
	f(x) = \frac{\left|x-c_x\right|}{b_x}.
\end{equation}
The $V$-shape is obtained from the absolute value of the abscissa $x$ shifted to the center $c_x$ of $M$.
The amplitude of the height is controlled by the inverse of the bounding box size $b_x$ of $M$ on the $x$ axis.

%%%%%%%%%%%%%%
The purpose of $g$ is to shift up the front of the interactive wave and shift down its back.
The idea is that a moving body displaces the water from its submerged hull, producing a high interactive wave in front of the body.
This effect is modeled using a quite simple affine function $g(z)=a\cdot z+b$.
The slope coefficient $a$ is defined considering the maximum height $h_b$ on the back of the mask and the height $h_f$ on its front.

While $h_b$ is defined by the user, the value of $w_f$ is estimated as follows:
\begin{equation}
	h_f = V_{M}h_{M}\times i_f\times\frac{v_w}{v_{M}},
\end{equation}
where $V_M$ is the velocity of the body $M$, $h_M$ is the full height of $M$, $i_f$ is a user parameter allowing more user control, $v_w$ is the submerged volume of $M$ (\cf Section~\ref{sec:fluidToSolidAction}), and $v_{M}$ is the full volume of $M$.
Hence, for more realism $h_f$ is made proportional to the velocity of $M$:
the faster $M$ moves, the bigger the waves generated by $M$ at its front are.

The values of $a$ and $b$ are straightforward to obtain by considering the height at the back and the front, so using the relations $g(z_{min})=h_b$ and $g(z_{max})=h_f$,
where $z_{min}$ is the $z$-coordinates of the back and $z_{max}$ the $z$-coordinate of the front.
From the first relation, it comes $b=h_b-a\cdot z_{min}$.
From the second, it comes $a=\left(h_f-h_b\right)/b_z$, where $b_z=z_{max}-z_{min}$ is the bounding box extend in $z$ axis.
This allows to simplify the term $b$ as $b=\left(h_bz_{max}-h_fz_{min}\right)/b_z$.

The functions $f(x)$ and $g(z)$ are then used to obtain the height of the mask as follows,
but using the user factor $b_w$ to allow more user control:
\begin{equation}
	h(\mathbf x)=b_w\left(f(x)+g(z)\right).
\end{equation}

Note that $h$ is only well-defined if the bow of $M$ is oriented on the positive $z$ axis.
Therefore, a 2d rotation is first applied to the vertices $\mathbf x$ before applying $h$.

Figure \ref{fig:mask} shows some views of the application of this interactive waves process for a simple rectangular boat.

\begin{figure}[htbp]
	\centering

	\subfloat[A theoretical view of the water height function $h(\mathbf x,t)$ that will be applied to a simple rectangle.]
	{
		\includegraphics[width=0.25\textwidth]{Figures/Mask3D_theoric_view.png}
		\label{subfig:maskTheoreticalView}
	}
	\hspace{1cm}
	\subfloat[The mask of a simple boat encoded in a single channel texture.]
	{
		\includegraphics[width=0.25\textwidth]{Figures/MaskHeightMap.png}
		\label{subfig:maskHeightMap}
	}
	\hspace{1cm}
	\subfloat[The mask of a simple boat in a 3D scene with a very low sea state.]
	{
		\includegraphics[width=0.25\textwidth]{Figures/Mask3D-contrast.png}
		\label{subfig:mask3D}
	}
	\caption{Three different views of the mask.}
	\label{fig:mask}
\end{figure}

%-------------------------------------------------------------------------
\section{Results and discussions}
\label{sec:results_and_discussions}

\begin{figure}
	\centering
	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt 
	\input{Tikz/methodsInteractions2.tikz}
	\caption{Interactions between the ocean waves simulation (Section~\ref{sec:oceanWaves}), the fluid-to-solid method (Section~\ref{sec:fluidToSolidAction}) and the solid-to-fluid method (Section~\ref{sec:solidToFluidAction}).
		The arrow represents an interaction between two methods.
		The arrow direction indicates that the method from which the arrow originates impacts the method to which the arrow is pointing to.}
	\label{fig:interactionMethods}
\end{figure}

Sections \ref{sec:oceanWaves}, \ref{sec:fluidToSolidAction} and \ref{sec:solidToFluidAction} describe three methods to handle the main parts of an ocean simulation: the free surface, and the coupling between a solid and a fluid.
Each method has been chosen to find the best ratio between high performance and physics accuracy.

Notice that each method interacts naturally with the two others, as shown in Figure~\ref{fig:interactionMethods}.
Indeed, Tessendorf's method generates the free surface.
As depicted by arrow $(0)$, the free surface method impacts the force calculation of the fluid-to-solid method.
The force calculation of the fluid-to-solid method impacts the mask calculation (arrow $(1)$), which is the input of the solid-to-fluid method.
The solid-to-fluid method modifies the free surface around the solid and is an input of the ocean waves simulation (arrow $(2)$).
All these interactions loops through the simulation over time. 
As a result, the three components of \framework have an impact on the two others (arrows $(3)$, $(4)$ and $(5)$) at the next time-step.
For example, since the solid-to-fluid method impacts the free surface (arrow $(2)$), and the free surface impacts the force calculation used in the fluid-to-solid (arrow $(0)$), then indirectly the solid-to-fluid method has an indirect impact on the fluid-to-solid action (arrow $(3)$).
%The free surface is used to evaluate the forces applied to the bodies, \ie to compute the position of each triangle relatively to the water.
%The position of these triangles define the parameters for the mask calculation, \ie the input parameter of the FDM that modifies the free surface that will be used for forces calculation, and so on.

Two experiments with different solids have been conducted to demonstrate the efficiency of \framework.
These experiments ran on a \href{https://www.intel.fr/content/www/fr/fr/products/sku/199316/intel-core-i710700-processor-16m-cache-up-to-4-80-ghz/specifications.html}{Intel Core i7-10700} and a \href{https://www.nvidia.com/fr-fr/geforce/graphics-cards/compare/?section=compare-20}{Nvidia GeForce RTX 2070 Super}.
The ocean was simulated with three cascades using a resolution of $256\times265$.
Moreover, the resolution of FDM grids is set to $512\times 512$.

\begin{figure}[htbp]
	\centering

	\subfloat[\emph{One-solid} using sea state $1$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/experiment1-1-contrast.png}
		\label{fig:experiment1-1}
	}
	\hspace{1cm}
	\subfloat[\emph{One-solid} using sea state $6$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/experiment1-2-contrast.png}
		\label{fig:experiment1-2}
	}
    \vspace{1cm}
 	\subfloat[\emph{Ten-solids} using sea state $1$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/experiment2-contrast.png}
		\label{fig:experiment2-1}
	}
	\hspace{1cm}
	\subfloat[\emph{Ten-solids} using sea state $4$.]
	{
		\includegraphics[width=0.4\textwidth]{Figures/experiment2-3-contrast.png}
		\label{fig:experiment2-2}
	}
	\caption{The two experiments with different sea state: \emph{one-solid} on top row, and \emph{ten-solids} on bottom one.}
	\label{fig:experiments}
\end{figure}

The first experiment called \emph{one-solid} simulates a single solid corresponding to a motorboat.
This motorboat is running in the scene.
Its model is composed of $165$ triangles for the simulation. 
Two captures of \emph{one-solid} can be seen in the top row of \figurename~\ref{fig:experiments}.
The sea-state is 1 on the Beaufort scale on \figurename~\ref{fig:experiment1-1}, leading to a very calm sea.
Hence, the interaction waves appear clearly behind the motorboat. 
On the contrary, the sea-state is 6 for \figurename~\ref{fig:experiment1-2}, leading to big waves and almost completely hiding interaction waves.

The second experiment called \emph{ten-solids} simulates ten solids.
It extends the first experiment as follows:
\begin{itemize}
	\item $1$ sailing boat with a mesh made of $270$ triangles.
	\item $2$ yachts with a mesh made of $146$ triangles.
	\item $6$ zodiacs with a mesh made of $568$ triangles.
\end{itemize} 
This makes a total of $4135$ triangles. 
Each of these ten solids uses its own grid for the fluid to solid coupling FDM. 
Two captures of the second experiment can be seen in the bottom row of \figurename~\ref{fig:experiments}.
The sea state is 1 for \figurename~\ref{fig:experiment2-1} where again small interaction waves can be seen for the motorboat (the others are not propelled). 
On \figurename~\ref{fig:experiment2-2} the sea state is set to 4, leading to relatively big waves.
Using 10 boats and so FDM grids does not impact too much the computation times, as seen below.

\tablename~\ref{fig:performances} shows a summary of GPU computation times for each method and with these two experiments. 
CPU computation times are not considered, as they are negligible compared to GPU ones. 
First, it can be observed that real-time performance is largely achieved, even with \emph{ten-solids}. 
Besides, as most of the computations are performed on the GPU, it can be observed that computation times do not grow linearly with the number of solids. 
Indeed, \framework processes each solid and each FDM grid simulation in parallel.

\begin{figure}[!h]
	\centering
	\begin{tabular}{|p{0.28\columnwidth}|p{0.24\columnwidth}|>{\centering\arraybackslash}p{0.14\columnwidth}|>{\centering\arraybackslash}p{0.14\columnwidth}|}
		\hline
        & & XP 1 & XP 2 \\
        & & (\emph{one-solid}) & (\emph{ten-solids}) \\
		\hline
		Tessendorf's method & Height (\S~\ref{subsec:waterHeight}) & 0.376 & 0.376\\
        \cline{2-4}
		(\S~\ref{sec:oceanWaves}) & Velocity (\S~\ref{subsec:velWater}) & 1.175 & 1.175\\
		\cline{2-4}
		  & Total & 1.551 & 1.551\\
		\hline
		Fluid-to-solid coupling & Geometry (\S~\ref{subsec:geomParam}) & 1.125 & 4.588\\
		\cline{2-4}
        (\S~\ref{sec:fluidToSolidAction}) & Forces (\S~\ref{subsec:forces}) & 0.406 & 3.677\\
		\cline{2-4}
        & Total & 1.531 & 8.265\\
		\hline
		Solid-to-fluid coupling & FDM (\S~\ref{subsec:finiteDifferenceMethod}) & 0.100                    & 0.831                    \\
		\cline{2-4}
        (\S~\ref{sec:solidToFluidAction}) & Mask (\S~\ref{subsec:mask}) & 0.238 & 1.656\\
		\cline{2-4}
        & Total & 0.338 & 2.487\\
		\hline
		\rowcolor{bleuC}\textbf{Total} & & \textbf{3.414} & \textbf{12.297}\\
		\hline
	\end{tabular}
	\caption{%
        GPU computation times in milliseconds for the main parts of \framework. 
        Only GPU execution times is shown, as CPU ones are negligible.}
	\label{fig:performances}
\end{figure}

%-------------------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This article presents three methods that together produce a realistic albeit approximate simulation of the ocean, and the way used to combine them together. % indiquer la nouveauté du papier ici...
These methods are simple to implement and fully parallelizable on GPU.
\Framework is a demonstration that these methods achieve real-time performance on a desktop PC (\cf section \ref{sec:results_and_discussions}).

This article gives an original expression for the ocean velocity associated with the Tessendorf method, as well as the key details of its implementation. This velocity is an important input for the realistic calculation of fluid-induced forces on solids.
This also improves the stability and physical coherence of the FDM for wave generation.

The two solid/fluid coupling methods are highly efficient, but they present many physical approximations.
Hence, to increase the realism of \framework, future works will include the hybridization of a smoothed particle hydrodynamics (SPH) method for solid/fluid coupling and Tessendorf methods on a larger scale. % pas certain de comprendre ta dernière phrase :-p

%-------------------------------------------------------------------------
\subsection*{Acknowledgements}
%Add the ANRT supports here, including the contract number...
The first author David Algis has been supported by the \href{https://www.anrt.asso.fr/fr}{Association nationale de recherche et de technologie (ANRT)}\footnote{Contract number 2021/1157} and by the \href{https://www.magelis.org/}{Pôle Image Magélis}. We would like to thank the colleague of David Algis at Studio Nyx, particularly Jeremy Bois and David Deckeur. Thanks to Naval Group for initiating this project. Finally, we would also like to thank Yannick Privat at the \href{https://irma.math.unistra.fr/}{Institut de recherche mathématique avancée (IRMA)} which help us on the velocity calculation.

% All entries in the bibliography \textbf{must} have URLs unless no URL is possible (e.g., book, very old thesis).

%-------------------------------------------------------------------------
\small
\bibliographystyle{jcgt}
\bibliography{bibliography}

% \section*{Index of Supplemental Materials}
% When supplemental materials such as video, data sets, and source code are provided with an article, briefly describe them by directory or filename here.

%-------------------------------------------------------------------------
\section*{Author Contact Information}

\hspace{-2mm}
\begin{tabular}{p{0.7\textwidth}p{0.5\textwidth}}
	David Algis\newline
	Bât. H1 - SP2MI\newline
	TSA 41123\newline
	86073 Poitiers Cedex 9\newline
	FRANCE\newline
	\href{mailto:david.algis@univ-poitiers.fr}{david.algis@univ-poitiers.fr}
\end{tabular}

%-------------------------------------------------------------------------
%\newpage
\appendix

\section*{Appendices}
%-------------------------------------------------------------------------
\section{Water velocity}
\label{sec:annexe}
The water height is deduced from system of equations described in Equations (\ref{equa:systemTessendorf}), and its solution according to Tessendorf is given by Equation (\ref{equa:heightWater}) in Section \ref{sec:oceanWaves}.
This appendix uses these equations to obtain an expression of the velocity potential and the water velocity.
To simplify the derivation, only one wave vector $\mathbf{k}$ is considered in Equation (\ref{equa:heightWater}) as Equations (\ref{equa:systemTessendorf}) are all linear.

First, the velocity potential for $y=0$ is deduced from first equation in \ref{equa:systemTessendorf}, and Equations (\ref{equa:heightWater}) and \ref{equa:heightWaterBis} as follows:
\begin{align}
	\phi(\mathbf{x},0,t)
	 & = \int \frac{\partial \phi}{\partial t}(\mathbf{x},0,t)dt \nonumber                                                                                                             \\
	 & =\int-gh(\mathbf{x},t)dt \nonumber                                                                                                                                              \\
	 & = -g\exp(i\mathbf{k}.\mathbf{\mathbf{x}})\int\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)dt \nonumber                     \\
	 & = \frac{ig}{\omega(k)}\exp(i\mathbf{k}.\mathbf{\mathbf{x}})\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\label{equa:phiy0}
\end{align}
where $\mathbf{x}=\left[\begin{array}{cc}x&z\end{array}\right]^T$.
It is assuming that potential velocity at any depth $y$ can be written as follows:
\begin{equation}
	\phi(\mathbf{x},y,t) = \tilde{\phi}(y, t,\mathbf{k})\exp(i \mathbf{k}\cdot \mathbf{x}).
\end{equation}
Notice that this expression is convenient as it can be computed with an inverse Fast Fourier Transform.
The function $\tilde{\phi}(y, t,\mathbf{k})$ is defined as follows:
\begin{equation}
	\tilde{\phi}(y, t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
\end{equation}
where \textit{attenuation function} $E(\mathbf{k},y)$ has to be determined from Equations \ref{equa:systemTessendorf}.
A first condition comes from Equation \ref{equa:phiy0}:
\begin{equation}
	E(\mathbf{k}, 0)=1 \label{equa:initialE}
\end{equation}
Then, the second equation from system \ref{equa:systemTessendorf} can be imposed on the attenuation function.
First, the left term is developed as follows:
\begin{align}
	\Delta \phi\nonumber= & \left(-k_x^2 E(\mathbf{k},y) -k_z^2E(\mathbf{k},y) + E''(\mathbf{k},y)\right)\nonumber                                                                                      \\
	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
	\exp\left( i \mathbf{k}\cdot \mathbf{x}\right)\nonumber                                                                                                                                             \\
	=                     & \left(E''(\mathbf{k},y)-k^2E(\mathbf{k},y)\right) \nonumber                                                                                                                 \\
	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) \exp\left(i \mathbf{k}\cdot \mathbf{x}\right)
\end{align}
Hence, condition $\Delta \phi=0$ is equivalent to find a solution of the following second order ordinary differential linear equation:
\begin{equation}
	E''(\mathbf{k}, y)-k^2E(\mathbf{k}, y) = 0.
\end{equation}
It can be deduced that for $-H<y\leq 0$, $E(\mathbf{k}, y)$ is as follows:
\begin{equation}
	E(\mathbf{k}, y)=C_1\cosh(ky+C_2)
\end{equation}
where $C_1,C_2\in\mathbb{R}$ are constant.
The derivative regarding height $y$ of $\phi(\mathbf{x},-H,t)$ can be rewritten as follows:
\begin{align*}
	\frac{\partial\phi(\mathbf{x},y,t)}{\partial y}= &
	\frac{\partial E(\mathbf{k},y)}{\partial y}\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)               \\
	=                                                & kC_1\sinh(ky+C_2)                                                                                                                                                      \\
	                                                 & \times\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)
\end{align*}
Hence, at depth $y=-H$ the fourth equation of system \ref{equa:systemTessendorf} leads to:
\begin{align}
	\sinh(-kH+C_2) & = 0 \nonumber \\
	C_2            & = kH.
\end{align}
The constant $C_1$ is obtained using Equation~\ref{equa:initialE} as follows:
\begin{align}
	E(\mathbf{k},0) & =1\nonumber           \\
	C_1\cosh(C_2)   & =1\nonumber           \\
	C_1             & =\frac{1}{\cosh(kH)}.
\end{align}
To resume, the full expression of the attenuation function for $-H\leq y\leq 0$ is as follows:
\begin{equation}
	E(\mathbf{k}, y)=\frac{\cosh\left(ky+kH\right)}{\cosh(kH)}.
\end{equation}
However, this expression cannot be used for large values of $k$ and sea height $H$, where the hyperbolic cosine becomes bigger than the limit of real numbers even in double precision.
Hence, $E$ quickly reach the indeterminate form $\frac{\infty}{\infty}$.
To fix this problem, the following physical assumption is made: it is assumed \textbf{deep water}, \ie that the sea depth $H$ tends towards $\infty$.
First, $E$ is rewritten as follows:
\begin{align*}
	E(\mathbf{k},y) & = \frac{\cosh(ky+kH)}{\cosh(kH)}\nonumber                            \\
	                & = \frac{\cosh(ky)\cosh(kH) + \sinh(kY)\sinh(kH)}{\cosh(kH)}\nonumber \\
	                & = \cosh(ky) + \sinh(ky)\tanh(kH)
\end{align*}
Then, considering that $\underset{H \rightarrow \infty}{\lim}\tanh(kH)=1$, the following approximation is obtained:
\begin{align}\label{equa:attenuationDeep1}
	E(\mathbf{k},y) \approx & \underset{H \rightarrow \infty}{\lim} E(\mathbf{k},y)\nonumber               \\
	\approx                 & \underset{H \rightarrow \infty}{\lim}\cosh(ky) + \sinh(ky)\tanh(kH)\nonumber \\
	\approx                 & \exp(ky).
\end{align}

This expression is valid for height $y\leq0$, but not usable above for $y>0$.
Several possibilities exist to overcome this problem:
\begin{enumerate}
	\item The most naive one assumes $E(\mathbf k,y>0)=1$.
	\item A more realistic solution is based on a linear extrapolation using the derivative at $y=0$, as follows:
	      \begin{equation}
		      \label{equa:attenuationExtrapolation}
		      E(\mathbf{k}, y)=
		      \begin{cases}
			      E(\mathbf k,0)+y\frac{\partial E}{\partial y}(\mathbf k,0) & \text{if~}y>0, \\
			      \exp(ky)                                                   & \text{else.}
		      \end{cases}
	      \end{equation}
\end{enumerate}

More sophisticated solutions are summarized in the Molin's book (2002) \cite{molinHydrodynamiqueStructuresOffshore2002}.
Nevertheless, they are too complex to compute as they make the attenuation function depending on $h(\mathbf x,t)$.
For this reason, \framework uses Equation~\ref{equa:attenuationExtrapolation}, with the following simplified form:
\begin{equation}
	\label{equa:attenuationFull}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if~}y>0, \\
		\exp(ky) & \text{else.}
	\end{cases}
\end{equation}

Note that while the third equation of system \ref{equa:systemTessendorf} is not used here, it can help to deduce the \emph{dispersion relation}.

Equation~\ref{equa:attenuationDeep1} and the definition of potential velocity $\nabla\phi=\mathbf{v}$ lead to the velocity $\mathbf{v}$. First, function $\tilde{\phi}$ is rewritten as follows:
\begin{equation}
	\tilde{\phi}(y,t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
\end{equation}
Then the gradient of $\phi$ is as follows:
\begin{equation}
	\nabla \phi = \begin{pmatrix}
		v_x \\
		v_y \\
		v_z
	\end{pmatrix} =\tilde{\mathbf{v}}(y,\mathbf{k},t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
\end{equation}
where $\tilde{\mathbf{v}}(\mathbf{k},t,y)$ is the vector equal to:
\begin{equation}
	\tilde{\mathbf{v}}(y,\mathbf{k},t) = \exp(ky)\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
	\begin{pmatrix}
		\frac{-k_xg}{\omega(k)} \\
		i\omega(k)              \\
		\frac{-k_zg}{\omega(k)}
	\end{pmatrix}.
\end{equation}
Finally, reintroducing a sum on a set of waves vectors $\mathbf{k}$ leads to:
\begin{equation}
	\nabla \phi =
	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k},t, y)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Navier-Stokes equations approximation}\label{sec:annexe-NS}
This appendix proposes a derivation of Tessendorf's equations \ref{equa:systemTessendorf} from Navier-Stokes equations. As stated in Teman (2001) \cite{temamNavierStokesEquations2001} and assuming conservation of mass, Navier-Stokes equations are given by:
\begin{equation}
	\begin{cases}
		\frac{\partial}{\partial t}\left(\rho \mathbf{u}\right) + \mathbf{u} \cdot \nabla (\rho \mathbf{u}) = - \nabla p + \rho\mathbf{g} +  \nabla^2 (\rho\nu \mathbf{u}) \\
		\frac{\partial}{\partial t}\left(\rho \mathbf{u}\right) + \nabla\cdot (\rho\mathbf{u}) = 0
	\end{cases}
	\label{equa:ns0}
\end{equation}
where $\mathbf{u} :\mathbb{R}^3\mapsto \mathbb{R}^3$ is the flow velocity, $\rho$ is the time-dependent density of the fluid, $p: \mathbb{R}^3\mapsto \mathbb{R}$ is the pressure of the fluid, $\nu$ is the dynamic viscosity of the fluid.
The first equation is called \emph{Cauchy momentum equation} and the second \emph{continuity equation}.

Boundary conditions are necessary for simulations. In the case of solid, the boundary equation is as follows:
\begin{equation}
	\mathbf{u}\cdot\mathbf{n} = \mathbf{u}_{solid} \cdot \mathbf{n}
\end{equation}
where, $\mathbf{u}_{solid}$ is the velocity of the solid and $\mathbf{n}$ is the normal to its surface.

Some assumptions are made to simplify these equations, and to derive Tessendorf equations.
They are discussed below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Incompressibility}
As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530m.s^{-1}$\footnote{\cf \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.
More formally:
\begin{equation}
	\rho(t)=\rho.
\end{equation}
Consequently, density can be factorized in each term of equations \ref{equa:ns0}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Inviscid}
A fluid can be considered inviscid if $Re>>1$ (\cf Section~2.3.1 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}) where $Re=VL/\nu$ is the \emph{Reynolds number} of the fluid defined by the order of magnitude of its velocity $V$, its characteristic length $L$ and its dynamic viscosity $\nu$.
For Tessendorf derivation, $V$ is the velocity of waves, $L$ its wave length and for seawater at $20$°C its is stated that $\nu=1.0508.10^{-6}$ (\cf the measure of the ITTC \cite{ittcFreshWaterSeawater2011}).
Therefore, the inequality is true for product $VL >> 10^{-5}$.
For these values the ocean can be considered as flat.
Hence, the viscosity term $\nu$ of Navier-Stokes equations \ref{equa:ns0} can be neglected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Irrotational}
If a fluid is irrotational at time $t_0$, it must remain irrotational unless there are changes in the boundary conditions.
Clearly, flat sea are irrotational fluid, corresponding to $\nabla\times\mathbf{u} = \mathbf{0}$.
Therefore, it can assume that for calm waves the fluid remains irrotational.
Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{\cf \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.

These three assumptions lead to the Bernoulli equation (\cf Section~5.3.2 of Hulin \etal (2001), which is the following simplified version of the Cauchy momentum equation:
\begin{equation}
	\frac{\partial \phi}{\partial t} + \frac{1}{2}\|\nabla\phi\|^2 + \frac{p}{\rho} + gy = 0
	\label{equa:bernoulli}
\end{equation}

Moreover, the continuity equation is now:
\begin{equation}
	\Delta \phi = 0 \label{equa:continuityS}
\end{equation}

Nevertheless, these non-linear equations are still too complex to be solved at large scaled and in real-time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Interface condition}
It is necessary to define a boundary condition at the interface between the ocean fluid and the air fluid.
To simplify, it is assume that the air can be represented as a region with constant pressure\footnote{As shown by the \href{https://en.wikipedia.org/wiki/Barometric_formula}{barometric formula} the atmospheric pressure has a very low evolution in function of altitude.}.
Since air density is a hundred of times lighter than water, its effects on the free surface are assumed to be negligible.
As with an incompressible fluid, only the differences in pressure matter.
Therefore, the air pressure is defined as an arbitrary constant, leading to:
\begin{equation}
	p = 0\text{~~~~~~~~~~~at free surface.}
\end{equation}

Hence, this allows to remove the pressure term in Bernoulli equation \ref{equa:bernoulli} at free surface.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Height field}
At this point, equations are still too complex to be used in real-time, especially in terms of dimension:
it seems inconceivable to solve them at large scale in 3d.
Therefore, to reduce their dimensions it is assumed that the surface geometry is described by a height field, as follows:
\begin{equation}
	y = h(\mathbf x, t).
\end{equation}

This hypothesis excludes a lot of physical phenomena like breaking waves, but it helps to reach large scale domain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Free surface follows the velocity of water}
Next step assumes that the free surface follows the velocity of the water.
Hence, for all $\mathbf x$ on the horizontal axes at time $t + dt$ the height field is as follows:
\begin{equation}
	h(\mathbf{x^{\prime}}, t + dt) = h(\mathbf x, t) + {v_y}dt\label{equa:freeSurfaceFollowWater0}
\end{equation}
where:
\begin{equation}
	\mathbf{x^{\prime}} = \mathbf x + dt\left[\begin{array}{c}v_x\\ v_y\end{array}\right]
	\label{equa:freeSurfaceFollowWater1}
\end{equation}

The Taylor expansion of $h(\mathbf{x^{\prime}}, t + dt)$ leads to:
\begin{equation}
	h(\mathbf{x^{\prime}}, t + dt) = h(\mathbf{x}, t + dt) + v_xd_t\frac{\partial h(\mathbf{x}, t)}{\partial x} + v_zd_t\frac{\partial h(\mathbf{x}, t)}{\partial z} + \ldots
\end{equation}

Neglecting the nonlinear terms in the Taylor expansion and reintegrating the development of $h(\mathbf{x^{\prime}}, t + dt)$ into Equation \ref{equa:freeSurfaceFollowWater0} and then dividing by $dt$ leads to the following result:
\begin{equation}
	\begin{aligned}
		\frac{h(\mathbf x, t + dt) - h(\mathbf x, t)}{dt} + v_x\frac{\partial h(\mathbf x, t)}{\partial x} + v_z\frac{\partial h(\mathbf x, t)}{\partial z} & = \frac{h(\mathbf x, t)}{dt} + v_y \\
		\frac{\partial h}{\partial t} + v_x\frac{\partial h}{\partial x} + v_z\frac{\partial h}{\partial z}                                                 & = v_y
	\end{aligned}
\end{equation}

This leads to the following \emph{advection equation}:
\begin{equation}
	\frac{D}{Dt}(h(\mathbf x) - y) = 0
	\label{equa:advection}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Remove quadratic term}
The quadratic term in the Bernoulli equation \ref{equa:bernoulli} makes solving the equation too complex.
Therefore, the velocity is assumed to be sufficiently small to neglect the quadratic term relative to the others.
It leads to the "final" Bernoulli equation expressed as follows:
\begin{equation}
	\frac{\partial \phi}{\partial t} = -gh(\mathbf x,t) \label{equa:bernoulliS}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Smooth free surface}
Similarly, the nonlinear terms in the advection equation \ref{equa:advection} make the equation's resolution too complex.
So the free surface height is assumed to be sufficiently smooth to neglect these terms relative to others.
As a result, advection equation \ref{equa:advection} becomes:
\begin{equation}
	\frac{\partial h}{\partial t} = \frac{\partial \phi}{\partial y} \label{equa:advectionS}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Flat ocean bottom}
Now, it is assumed that the ocean depth is constant, and equal to $H$.
Then, the ocean floor is stationary and its normal to the ocean floor is $(0,1,0)$.
Hence, boundary conditions at $y=-H$ are defined as:
\begin{equation}
	\begin{aligned}
		\mathbf{u}\cdot\mathbf{n}        & = 0 \\
		\frac{\partial \phi}{\partial y} & = 0
	\end{aligned}
	\label{equa:bottomCondition}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Infinite ocean}
To avoid boundary conditions on $\mathbf x$, it is assumed that the ocean is horizontally infinite and periodic, allowing a finite simulation area.
Therefore, the simulation area size is defined squared as $L \times L$.
It leads to an ocean tile that can be infinitely duplicated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Small free surface}
To simplify the location at which all equations are solved, it is assumed that the height of the free surface is sufficiently low. Hence, the equations can be solved at $\mathbf x=\mathbf0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Conclusion}
Finally, putting all together leads to the Tessendorf solution.
More precisely, Tessendorf equations system is made from Equations (\ref{equa:bernoulliS}), (\ref{equa:continuityS}), (\ref{equa:advectionS}), (\ref{equa:bottomCondition}).

%%%%%%%%%%%%%%%%
\section{Two Fourier transforms in one}
\label{sec:annexe-2IFFT}
This section proposes a proof of Theorem \ref{theorem:ifft}.

\begin{proof}
	This demonstration relies on the following formula that links the real part of the IFFT and the sum of the coefficients:
	\begin{equation}
		\label{equa:realFormula}
		\mathcal{F}^{-1}\left(\frac{1}{2}\left(A+A^{\dagger}\right)\right) =
		\Re\left(\mathcal{F}^{-1}\left(A\right)\right)
	\end{equation}
	where $A=(a_{n,m})$ is any Hermitian complex matrix of $M_n(\mathbb{C})$, and $A^{\dagger}$ the transpose of its conjugate.
	Using the definition and linearity of the discrete Fourier transform, it follows for one term $(k,l)$ of the IFFT:
	\begin{align}
		\mathcal{F}^{-1}\left(A+A^{\dagger}\right)_{k,l} & =
		\mathcal{F}^{-1}\left(A\right)_{k,l} + \mathcal{F}^{-1}\left(A^{\dagger}\right)_{k,l}\nonumber                                                                                                                  \\
		                                                 & =\sum_{n,m}a_{n,m}\exp\left(-i\frac{2\pi}{N}\left(kn+lm\right)\right) + \sum_{m,n}a_{m,n}^{\ast}\exp\left(-i\frac{2\pi}{N}\left(km+ln\right)\right)\nonumber \\
		                                                 & =\sum_{n,m}(a_{n,m}+a_{n,m}^{\ast})\exp\left(-i\frac{2\pi}{N}\left(kn+lm\right)\right)\nonumber                                                              \\
		                                                 & = 2\Re\left(\mathcal{F}^{-1}\left(A\right)_{k,l}\right)
	\end{align}

	Since $A$ is Hermitian by hypothesis, Equation (\ref{equa:realFormula}) becomes:
	\begin{equation}
		\mathcal{F}^{-1}\left(A\right) = \Re\left(\mathcal{F}^{-1}\left(A\right)\right)
	\end{equation}

	Applying this last equation to $X$ and $iY$ leads to:
	\begin{equation}
		\begin{aligned}
			\mathcal{F}^{-1}\left(X+iY\right) & =
			\mathcal{F}^{-1}\left(X\right) + i\mathcal{F}^{-1}\left(Y\right)                                                                       \\
			                                  & = \Re\left(\mathcal{F}^{-1}\left(X\right)\right) + i\Re\left(\mathcal{F}^{-1}\left(Y\right)\right)
		\end{aligned}
	\end{equation}

\end{proof}
%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%
\afterdoc
%%%%%%%%%%%%%%%
\end{document}

### 2024-11-12 paper.tex
Fixed author name

@@
 	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
-	\item The simulation of the alteration of a free surface by a solid, inspired by Coords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
+	\item The simulation of the alteration of a free surface by a solid, inspired by Cords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
 \end{itemize}
@@
 	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
-	\item An algorithm to compute the input of the Coords and Staadt method more realistically from the free surface.
+	\item An algorithm to compute the input of the Cords and Staadt method more realistically from the free surface.
 \end{itemize}
@@
 Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the FDM to simulate waves on the free surface.
-This method was improved by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
+This method was improved by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
 Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
@@
 \label{sec:solidToFluidAction}
-\Framework relies on the method developed by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to simulate the action of the solid on the fluid.
+\Framework relies on the method developed by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to simulate the action of the solid on the fluid.
 The simulation of this action is 3-dimensional by nature.
@@
 The waves generated by a given body $M$ are defined in its local coordinates system.
-Hence, when $M$ moves, the waves have to move accordingly as done by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
+Hence, when $M$ moves, the waves have to move accordingly as done by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
 A first solution consists of applying the translation to the grid at times $t$ and $t-dt$ and then of using the integration scheme of Equation (\ref{eq:fdm_scheme}).
@@
 Knowing all the vertices in the mask, their height must be given as input of the FDM.
-\Framework uses a simple model that tries to reproduce parts of the idea of Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
+\Framework uses a simple model that tries to reproduce parts of the idea of Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
 By construction of the grid $Z$, $M$ has is bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.

### 2024-11-12 paper.tex
Fixed abbreviation that was used before definition

@@
 The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
-Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the FDM to simulate waves on the free surface.
+Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the finite-difference method (FDM) to simulate waves on the free surface.
 This method was improved by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
@@
 
-For solid-fluid coupling, \framework based its calculations on the finite-difference method (FDM) described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
+For solid-fluid coupling, \framework based its calculations on the (FDM) described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
 %-------------------------------------------------------------------------

### 2024-11-12 paper.tex
Fixed mask equations

@@
 	\label{eq:2dwave}
-	\nabla h(\mathbf x,t)-\frac{1}{c^2}\frac{\partial^2 h(\mathbf x,t)}{\partial t^2}=0
-\end{equation}
-where $h(\mathbf x,t)$ is the height of the water surface at horizontal position $\mathbf x$ and time $t$, $\nabla=\frac{\partial^2}{\partial x^2}+\frac{\partial^2}{\partial z^2}$ is the Laplacian in 2D and $c$ is the wave velocity expressed in $m.s^{-1}$.
+	\Delta h(\mathbf x,t)-\frac{1}{c^2}\frac{\partial^2 h(\mathbf x,t)}{\partial t^2}=0
+\end{equation}
+where $h(\mathbf x,t)$ is the height of the water surface at horizontal position $\mathbf x$ and time $t$, $\Delta=\frac{\partial^2}{\partial x^2}+\frac{\partial^2}{\partial z^2}$ is the Laplacian in 2D and $c$ is the wave velocity expressed in $m.s^{-1}$.
 \Framework uses the following Dirichlet boundary condition:
@@
 	\begin{cases}
-		\left(1-\frac{||v_M(t)||}{v_{max}}\right)d_0 + \frac{||v_M(t)||}{v_{max}}d_{max} & \text{for~~}||v_M(t)||\leq v_{max}, \\
+		\left(1-\frac{||\mathbf{v_M}(t)||}{v_{max}}\right)d_0 + \frac{||v_M(t)||}{v_{max}}d_{max} & \text{for~~}||v_M(t)||\leq v_{max}, \\
 		d_{max}                                                                          & \text{else}
@@
 \begin{equation}
-	h_f = V_{M}h_{M}\times i_f\times\frac{v_w}{v_{M}},
-\end{equation}
-where $V_M$ is the velocity of the body $M$, $h_M$ is the full height of $M$, $i_f$ is a user parameter allowing more user control, $v_w$ is the submerged volume of $M$ (\cf Section~\ref{sec:fluidToSolidAction}), and $v_{M}$ is the full volume of $M$.
+	h_f = \mathbf{v_M}h_{M}\times i_f\times\frac{v_w}{v_{M}},
+\end{equation}
+where $\mathbf{v_M}$ is the velocity of the body $M$, $h_M$ is the full height of $M$, $i_f$ is a user parameter allowing more user control, $v_w$ is the submerged volume of $M$ (\cf Section~\ref{sec:fluidToSolidAction}), and $v_{M}$ is the full volume of $M$.
 Hence, for more realism $h_f$ is made proportional to the velocity of $M$:

### 2024-11-12 paper.tex
Use correct quotation mark

@@
 His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solutions to an approximation of the Navier-Stokes equation on a 2D height field.
-This method is commonly known as the "Tessendorf method".
+This method is commonly known as the “Tessendorf method”.
 
 The realism of this method was significantly improved by Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, who proposed an empirically based directional spectrum.
-In addition, he proposed to add a parameter called "swell", which allows the transition from empirical directional propagation to parallel waves.
+In addition, he proposed to add a parameter called “swell”, which allows the transition from empirical directional propagation to parallel waves.
 
@@
 
-In \framework, the directional spectrum relies on the work of Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, that mixes together the Donelan-Banner spectrum to control the direction of the waves and a "swell parameter".
+In \framework, the directional spectrum relies on the work of Horvath (2015) \cite{horvathEmpiricalDirectionalWave2015}, that mixes together the Donelan-Banner spectrum to control the direction of the waves and a “swell parameter”.
 Let us recall that, if the wind direction is given by $\theta_0$, the wind direction bias $\theta$ can be defined as follows:
@@
 
-First, it should be noticed that the expressions of the ocean surface's height (Equation~\ref{equa:heightWater}), the displacements (Equation~\ref{equa:hori-displacement}) and their five derivatives, and the velocity (Equation~\ref{equa:velocity}) have "a form of an inverse discrete Fourier transform".
+First, it should be noticed that the expressions of the ocean surface's height (Equation~\ref{equa:heightWater}), the displacements (Equation~\ref{equa:hori-displacement}) and their five derivatives, and the velocity (Equation~\ref{equa:velocity}) have “a form of an inverse discrete Fourier transform”.
 More precisely, the \textbf{frequency space} coefficients appearing before the term $\exp(i\mathbf{k}\cdot\mathbf x)$ are stored in a square matrix of size $N\times N$, for each expression.
@@
 Also, this function decreases slowly (actually with $\ln(250+1)=5.5$), hence the logarithmic distribution uses the square of $|y_1|$.
-Lastly, the parameter $\alpha$ controls the "squeezing" in the center, while the parameter $\beta$ controls the deepest value of $d$.
+Lastly, the parameter $\alpha$ controls the “squeezing” in the center, while the parameter $\beta$ controls the deepest value of $d$.
 Notice that $\beta$ is chosen as the solution of $d(y_{min})=\frac{y_{min}}{2}$ but not at $y_{min}$: it is the solution giving the best accuracy after multiple tries.
@@
 Therefore, the velocity is assumed to be sufficiently small to neglect the quadratic term relative to the others.
-It leads to the "final" Bernoulli equation expressed as follows:
+It leads to the “final” Bernoulli equation expressed as follows:
 \begin{equation}

### 2024-11-12 paper.tex
Remove cf usage which makes no sense see https://scientistseessquirrel.wordpress.com/2016/06/13/friends-dont-let-friends-use-cf/

@@
 \def\ie{\emph{i.e.}\xspace}
-\def\cf{\emph{cf.}\xspace}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
-This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (\cf Figure \ref{fig:overviewFramework}).
+This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (see Figure \ref{fig:overviewFramework}).
 This new framework incorporates three different methods from the literature:
@@
 	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water.
-\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{\cf appendix \ref{sec:annexe} for more details on this assumption.}.
+\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{see appendix \ref{sec:annexe} for more details on this assumption.}.
 	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x z\right]^T$ and at time $t$.
@@
 
-Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{\cf appendix \ref{sec:annexe-NS}.}:
+Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{see appendix \ref{sec:annexe-NS}.}:
 \begin{equation}
@@
 \end{equation}
-where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, \cf appendix \ref{sec:annexe} for more details on this assumption.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
+where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see appendix \ref{sec:annexe} for more details on this assumption.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
 Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
@@
 
-Finally, \framework uses a last parameter $\delta\in[0,1]$ that represents how much the waves orientation is constant ($\delta = 0$) or not ($\delta = 1$) (\cf Figure \ref{fig:dispersionEffect}).
+Finally, \framework uses a last parameter $\delta\in[0,1]$ that represents how much the waves orientation is constant ($\delta = 0$) or not ($\delta = 1$) (see Figure \ref{fig:dispersionEffect}).
 Its meaning is as follows:
@@
 \label{subsec:velWater}
-Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{\cf appendix \ref{sec:annexe} for more details on the origin of this expression}:
+Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{see appendix \ref{sec:annexe} for more details on the origin of this expression}:
 \begin{equation}\label{equa:velocity}
@@
 
-In \framework, the number of IFFT applications is divided by two using Theorem \ref{theorem:ifft} on Hermitian matrices\footnote{A matrix $X=(x_{n,m})$ is Hermitian if it is equal to its own conjugate transpose: $X = X^{\dagger}$ or $x_{n,m}=x^*_{m,n}$.} (\cf annex \ref{sec:annexe-2IFFT} for its demonstration).
+In \framework, the number of IFFT applications is divided by two using Theorem \ref{theorem:ifft} on Hermitian matrices\footnote{A matrix $X=(x_{n,m})$ is Hermitian if it is equal to its own conjugate transpose: $X = X^{\dagger}$ or $x_{n,m}=x^*_{m,n}$.} (see annex \ref{sec:annexe-2IFFT} for its demonstration).
 \begin{theorem}
@@
 
-To make $P(d)$ and $A(d)$ having the same order of magnitude, the units are the \textbf{decisecond} for performance and the \textbf{meter per seconds} for accuracy (\cf figure \ref{fig:interpolationCriteria}).
+To make $P(d)$ and $A(d)$ having the same order of magnitude, the units are the \textbf{decisecond} for performance and the \textbf{meter per seconds} for accuracy (see figure \ref{fig:interpolationCriteria}).
 Moreover, the performance being the more important for real-time process, the following arbitrary choice are made for the coefficients of $J$:
@@
 	\caption{Multi-objective function $P$ to optimize, with two objectives: performance and accuracy.
-		Its minimum is reached by the interpolation degree of $8$. For more details on each objective, \cf Figures \ref{fig:interpolationCriteria} which represent them individually.}
+		Its minimum is reached by the interpolation degree of $8$. For more details on each objective, see Figures \ref{fig:interpolationCriteria} which represent them individually.}
 	\label{fig:multiObjectiveFunction}
@@
 Since $M$ moves, the shift indexes $k$, $l$, $o,$ and $p$ can be outside the boundary of the simulation zone $Z$.
-To solve this problem, the size $m$ of the boundary of $Z$ is increased such that $m>1$ (\cf Figure \ref{fig:examGrid}).
+To solve this problem, the size $m$ of the boundary of $Z$ is increased such that $m>1$ (see Figure \ref{fig:examGrid}).
 \Framework takes $m=16$ which is large enough for large translations.
@@
 \end{equation}
-where $\mathbf{v_M}$ is the velocity of the body $M$, $h_M$ is the full height of $M$, $i_f$ is a user parameter allowing more user control, $v_w$ is the submerged volume of $M$ (\cf Section~\ref{sec:fluidToSolidAction}), and $v_{M}$ is the full volume of $M$.
+where $\mathbf{v_M}$ is the velocity of the body $M$, $h_M$ is the full height of $M$, $i_f$ is a user parameter allowing more user control, $v_w$ is the submerged volume of $M$ (see Section~\ref{sec:fluidToSolidAction}), and $v_{M}$ is the full volume of $M$.
 Hence, for more realism $h_f$ is made proportional to the velocity of $M$:
@@
 These methods are simple to implement and fully parallelizable on GPU.
-\Framework is a demonstration that these methods achieve real-time performance on a desktop PC (\cf section \ref{sec:results_and_discussions}).
+\Framework is a demonstration that these methods achieve real-time performance on a desktop PC (see section \ref{sec:results_and_discussions}).
 
@@
 \paragraph{Incompressibility}
-As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530m.s^{-1}$\footnote{\cf \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
+As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530m.s^{-1}$\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
 Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.
@@
 \paragraph{Inviscid}
-A fluid can be considered inviscid if $Re>>1$ (\cf Section~2.3.1 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}) where $Re=VL/\nu$ is the \emph{Reynolds number} of the fluid defined by the order of magnitude of its velocity $V$, its characteristic length $L$ and its dynamic viscosity $\nu$.
-For Tessendorf derivation, $V$ is the velocity of waves, $L$ its wave length and for seawater at $20$°C its is stated that $\nu=1.0508.10^{-6}$ (\cf the measure of the ITTC \cite{ittcFreshWaterSeawater2011}).
+A fluid can be considered inviscid if $Re>>1$ (see Section~2.3.1 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}) where $Re=VL/\nu$ is the \emph{Reynolds number} of the fluid defined by the order of magnitude of its velocity $V$, its characteristic length $L$ and its dynamic viscosity $\nu$.
+For Tessendorf derivation, $V$ is the velocity of waves, $L$ its wave length and for seawater at $20$°C its is stated that $\nu=1.0508.10^{-6}$ (see the measure of the ITTC \cite{ittcFreshWaterSeawater2011}).
 Therefore, the inequality is true for product $VL >> 10^{-5}$.
@@
 Therefore, it can assume that for calm waves the fluid remains irrotational.
-Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{\cf \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.
-
-These three assumptions lead to the Bernoulli equation (\cf Section~5.3.2 of Hulin \etal (2001), which is the following simplified version of the Cauchy momentum equation:
+Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{see \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.
+
+These three assumptions lead to the Bernoulli equation (see Section~5.3.2 of Hulin \etal (2001), which is the following simplified version of the Cauchy momentum equation:
 \begin{equation}

### 2024-11-12 paper.tex
Remove abbreviation that wasn't defined

@@
 Varela and Guedes Soares (2011) \cite{varelaInteractiveSimulationShip2011} have used a table of pre-calculated RAO, which are later interpolated to achieve an interactive frame rate.
-However, to our knowledge, this type of method is either too expensive for real-time simulation or not general enough because it requires some physical pre-computation based on CFD or experimental data.
+However, to our knowledge, this type of method is either too expensive for real-time simulation or not general enough because it requires some physical pre-computation based on computational fluid dynamics or experimental data.
 

### 2024-11-12 paper.tex
Remove typos

@@
 
-For solid-fluid coupling, \framework based its calculations on the (FDM) described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
+For solid-fluid coupling, \framework based its calculations on the FDM described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
 %-------------------------------------------------------------------------

### 2024-11-12 paper.tex
Add missing space in vector

@@
 \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{see appendix \ref{sec:annexe} for more details on this assumption.}.
-	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x z\right]^T$ and at time $t$.
+	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~z\right]^T$ and at time $t$.
 	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.

### 2024-11-12 paper.tex
Fixed potential flow that was defined as a vector but is a scalar

--- 2024-11-12 paper.tex
@@
 \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{see appendix \ref{sec:annexe} for more details on this assumption.}.
-	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x z\right]^T$ and at time $t$.
+	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~tz\right]^T$ and at time $t$.
 	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
-	\item The vector $\phi(\mathbf x,y,t)\in\mathbb R^3$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
+	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
 	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.

### 2024-11-12 paper.tex
Fix typo
Fixed typos for greek variable

@@
 
-The product between the Donelan-Banner and $\xi$ spectra is denoted by $D_{DB\xi}$, and is expressed as follows:
+The product between the Donelan-Banner and $D_{\xi}$ spectra is denoted by $D_{DB\xi}$, and is expressed as follows:
 \begin{equation}
@@
 
-Note that we applied a first normalization factor $Q_{DB}$ to the simple Donelan Banner spectrum \ref{equa:ddb} before calculating this second normalization factor $\widetilde{Q_{DB\xi}}(r_{omega})$.
-In fact, if we simply approximate the full-spectrum normalization factor before applying it, we'll be faced with a normalization problem, because for small values of $r_{omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.
+Note that we applied a first normalization factor $Q_{DB}$ to the simple Donelan Banner spectrum (Equation \ref{equa:ddb}) before calculating this second normalization factor $\widetilde{Q_{DB\xi}}(r_{\omega})$.
+In fact, if we simply approximate the full-spectrum normalization factor before applying it, we'll be faced with a normalization problem, because for small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.
 

### 2024-11-12 paper.tex
Make notation more clear in interpolation degree calculation

@@
 More concretely, the velocity must be computed at any horizontal position in $I=[y_{min}, y_{max}]$.
-In \framework, $N$ growing depths $y_i$ are chosen in this interval $I$.
-For each of these depths $y_i$, the $i$-th velocity $\mathbf{v_i}(\mathbf x,y_i,t)$ is computed using one IFFT, leading to $N$ IFFTs for all the sampled depths $y_i$.
+In \framework, $d$ growing depths $y_i$ are chosen in this interval $I$.
+For each of these depths $y_i$, the $i$-th velocity $\mathbf{v_i}(\mathbf x,y_i,t)$ is computed using one IFFT, leading to $d$ IFFTs for all the sampled depths $y_i$.
 Then, the velocity $\mathbf{v}(\mathbf x,y,t)$ at any depth $y\in I$ is computed using an ad hoc interpolation from the different $\mathbf{v_i}$.
@@
 \begin{equation}
-	d(y_i) =
+	l_d(y_i) =
 	\begin{cases}
@@
 Also, this function decreases slowly (actually with $\ln(250+1)=5.5$), hence the logarithmic distribution uses the square of $|y_1|$.
-Lastly, the parameter $\alpha$ controls the “squeezing” in the center, while the parameter $\beta$ controls the deepest value of $d$.
-Notice that $\beta$ is chosen as the solution of $d(y_{min})=\frac{y_{min}}{2}$ but not at $y_{min}$: it is the solution giving the best accuracy after multiple tries.
+Lastly, the parameter $\alpha$ controls the “squeezing” in the center, while the parameter $\beta$ controls the deepest value of $l_d$.
+Notice that $\beta$ is chosen as the solution of $l_d(y_{min})=\frac{y_{min}}{2}$ but not at $y_{min}$: it is the solution giving the best accuracy after multiple tries.
 
@@
 
-Moreover, using the logarithmic distribution leads to $d(y_{min})=\frac{y_{min}}{2}$.
+Moreover, using the logarithmic distribution leads to $l_d(y_{min})=\frac{y_{min}}{2}$.
 Assuming that the velocity at $y_{min}$ is null, then for any depth between $y_{min}$ and $\frac{y_{min}}{2}$ the velocity can be computed between the one at $\frac{y_{min}}{2}$ and $0$.

### 2024-11-12 paper.tex
Remove informal expression

@@
 
-Knowing the interpolation method, the interpolation degree (number of points in the table) must be decided.
-Obviously, using more points results in better accuracy, but also more computation time.
+Knowing the interpolation method, the interpolation degree $d$ (number of points in the table) must be decided.
+Note, that using more points results in better accuracy, but also more computation time.
 Then, a trade-off between performance and accuracy must be determined.

### 2024-11-12 paper.tex
Fixed typo missing upper case for figure reference
Remove informal expression and makes abbreviation more uniform. Fixed equation error

@@
 
-A summary of the forces that are applied on solid can be found on figure \ref{fig:forcesComputation}.
+A summary of the forces that are applied on solid can be found on Figure \ref{fig:forcesComputation}.
 
@@
 
-Hence, \framework assumes that the interactive waves are solutions of the following classical 2-dimensional wave equation:
+Hence, \framework assumes that the interactive waves are solutions of the following 2D wave equation:
 \begin{equation}
@@
 	\begin{cases}
-		\left(1-\frac{||\mathbf{v_M}(t)||}{v_{max}}\right)d_0 + \frac{||v_M(t)||}{v_{max}}d_{max} & \text{for~~}||v_M(t)||\leq v_{max}, \\
+		\left(1-\frac{||\mathbf{v_M}(t)||}{v_{max}}\right)d_0 + \frac{||\mathbf{v_M}(t)||}{v_{max}}d_{max} & \text{for~~}||\mathbf{v_M}(t)||\leq v_{max}, \\
 		d_{max}                                                                          & \text{else}

### 2024-11-12 paper.tex
Clarify section reference

@@
 The idea behind the mask is to distort the grid of the FDM by reproducing the intersection surface between $M$ and the free surface.
-Here, free surface refers to the ambient waves of Tessendorf's method \ref{sec:oceanWaves} plus the interactive waves generated by all solids except $M$.
+Here, free surface refers to the ambient waves of Tessendorf's method (see Section \ref{sec:oceanWaves}) plus the interactive waves generated by all solids except $M$.
 It does \textbf{not} include the free surface generated by $M$, or the simulation would face a divergence issue.

### 2024-11-12 paper.tex
Harmonize the dimension abbreviation

@@
 Note that $h$ is only well-defined if the bow of $M$ is oriented on the positive $z$ axis.
-Therefore, a 2d rotation is first applied to the vertices $\mathbf x$ before applying $h$.
+Therefore, a 2D rotation is first applied to the vertices $\mathbf x$ before applying $h$.
 
@@
 At this point, equations are still too complex to be used in real-time, especially in terms of dimension:
-it seems inconceivable to solve them at large scale in 3d.
+it seems inconceivable to solve them at large scale in 3D.
 Therefore, to reduce their dimensions it is assumed that the surface geometry is described by a height field, as follows:

### 2024-11-12 paper.tex
Fixed mask calculations
Add missing space in sound equation

@@
 
-While $h_b$ is defined by the user, the value of $w_f$ is estimated as follows:
-\begin{equation}
-	h_f = \mathbf{v_M}h_{M}\times i_f\times\frac{v_w}{v_{M}},
+While $h_b$ is defined by the user, the value of $h_f$ is estimated as follows:
+\begin{equation}
+	h_f = ||\mathbf{v_M}||h_{M}\times i_f\times\frac{v_w}{v_{M}},
 \end{equation}
@@
 \paragraph{Incompressibility}
-As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530m.s^{-1}$\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
+As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530~m.s^{-1}$\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
 Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.

### 2024-11-12 paper.tex
Change footnote position for more clarity

@@
 \paragraph{Incompressibility}
-As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal to $c_{sound} = 1530~m.s^{-1}$\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile}.}.
+As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile} to $c_{sound} = 1530~m.s^{-1}$.}.
 Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.

### 2024-11-14 paper.tex
presents previous works -> presents the previous works...
Motivate using Cords

@@
 This article is organized as follows.
-Section~\ref{sec:relatedwork} presents previous works related to ocean simulation, including its interaction with solid.
+Section~\ref{sec:relatedwork} presents the previous works related to ocean simulation, including its interaction with solid.
 Sections \ref{sec:oceanWaves} to \ref{sec:solidToFluidAction} detail \framework, including the ocean surface simulation in Section~\ref{sec:oceanWaves}, the action of the fluid to solids in Section~\ref{sec:fluidToSolidAction} and the action of a solid to the fluid in Section~\ref{sec:solidToFluidAction}.
@@
 Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} presented what could be considered a seminal work in real-time ocean simulation.
-His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solutions to an approximation of the Navier-Stokes equation on a 2D height field.
+His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solution to an approximation of the Navier-Stokes equations on a 2D height field.
 This method is commonly known as the “Tessendorf method”.
@@
 
-For solid-fluid coupling, \framework based its calculations on the FDM described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
+For solid-fluid coupling, \framework based its calculations on the FDM described in Gomez (2000) \cite{gomezInteractiveSimulationWater2000} and its improvement by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}, because its ability to be real-time while providing quite realistic behavior.
+
 %-------------------------------------------------------------------------

### 2024-11-14 paper.tex
Fix plural
Add capital letter for footnote...

@@
 This section summarizes Tessendorf's method and the various improvements it has received.
-It also presents our first contributions: calculation of the fluid velocity at any depth.
+It also presents our first contribution: calculation of the fluid velocity at any depth.
 
@@
 	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water.
-\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{see appendix \ref{sec:annexe} for more details on this assumption.}.
+\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See appendix \ref{sec:annexe} for more details on this assumption.}.
 	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~tz\right]^T$ and at time $t$.
@@
 
-Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{see appendix \ref{sec:annexe-NS}.}:
+Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See appendix \ref{sec:annexe-NS}.}:
 \begin{equation}
@@
 \label{subsec:velWater}
-Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{see appendix \ref{sec:annexe} for more details on the origin of this expression}:
+Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See appendix \ref{sec:annexe} for more details on the origin of this expression}:
 \begin{equation}\label{equa:velocity}
@@
 \paragraph{Incompressibility}
-As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile} to $c_{sound} = 1530~m.s^{-1}$.}.
+As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal\footnote{See \url{https://en.wikipedia.org/wiki/Sound_speed_profile} to $c_{sound} = 1530~m.s^{-1}$.}.
 Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.
@@
 Therefore, it can assume that for calm waves the fluid remains irrotational.
-Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{see \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.
+Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{See \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.
 

### 2024-11-14 paper.tex
Fix typo in vector position (tz -> z)

@@
 \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See appendix \ref{sec:annexe} for more details on this assumption.}.
-	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~tz\right]^T$ and at time $t$.
+	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x~z\right]^T$ and at time $t$.
 	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.

### 2024-11-14 paper.tex
remove redundancy

@@
 \end{equation}
-where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being an integer constant given by user, and $\tilde{h}$ is given by:
+where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
 \begin{equation}

### 2024-11-14 paper.tex
fix plural

@@
 \end{equation}
-where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraph \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.
+where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.
 

### 2024-11-14 paper.tex
Fix acronym (jonsWAp)

@@
 
-\Framework uses the frequency spectrum developed by Hasselmann \etal (1973) \cite{k.hasselmannMeasurementsWindwaveGrowth1973} in the Joint North Sea Observation Project (JONSWAP).
+\Framework uses the frequency spectrum developed by Hasselmann \etal (1973) \cite{k.hasselmannMeasurementsWindwaveGrowth1973} in the Joint North Sea Wave Project (JONSWAP).
 This spectrum was developed for deep sea water, and hence is in line with our deep water hypothesis.

### 2024-11-14 paper.tex
Precise the meaning

@@
 	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water.
-\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See appendix \ref{sec:annexe} for more details on this assumption.}.
+\Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See Appendix \ref{sec:annexe} for more details on this assumption.}.
 	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x~z\right]^T$ and at time $t$.
@@
 
-Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See appendix \ref{sec:annexe-NS}.}:
+Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See Appendix  \ref{sec:annexe-NS}.}:
 \begin{equation}
@@
 \end{equation}
-where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see appendix \ref{sec:annexe} for more details on this assumption.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
-Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
+where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
+Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
 In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
@@
 \label{subsec:velWater}
-Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See appendix \ref{sec:annexe} for more details on the origin of this expression}:
+Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See Appendix   \ref{sec:annexe} for more details on the origin of this expression}:
 \begin{equation}\label{equa:velocity}

### 2024-11-14 paper.tex
fix redundancy
fix elusion

@@
 \end{equation}
-Hence, $Q_{DB}(\omega)$ is a normalization factor to make $D_{DB}(\omega, \theta)$ respect the condition \ref{equa:normalization}:
+Hence, $Q_{DB}(\omega)$ is a normalization factor to make $D_{DB}(\omega, \theta)$ respect condition \ref{equa:normalization}:
 \begin{equation}
@@
 Note that we applied a first normalization factor $Q_{DB}$ to the simple Donelan Banner spectrum (Equation \ref{equa:ddb}) before calculating this second normalization factor $\widetilde{Q_{DB\xi}}(r_{\omega})$.
-In fact, if we simply approximate the full-spectrum normalization factor before applying it, we'll be faced with a normalization problem, because for small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.
+In fact, if we simply approximate the full-spectrum normalization factor before applying it, we will be faced with a normalization problem, because for small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.
 

### 2024-11-14 paper.tex
Fix Donelan-Banner Hulk normalization factor explanation

@@
 
-Note that we applied a first normalization factor $Q_{DB}$ to the simple Donelan Banner spectrum (Equation \ref{equa:ddb}) before calculating this second normalization factor $\widetilde{Q_{DB\xi}}(r_{\omega})$.
-In fact, if we simply approximate the full-spectrum normalization factor before applying it, we will be faced with a normalization problem, because for small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically difficult to normalize.
+Note that a first normalization factor $Q_{DB}$ is applied to the Donelan-Banner spectrum (Equation \ref{equa:ddb}) before calculating the second normalization factor $\widetilde{Q_{DB\xi}}(r_{\omega})$.
+In fact, a normalization problem will occur if the full-spectrum normalization factor is simply approximated before its application, because for small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically challenging to normalize.
 

### 2024-11-14 paper.tex
Fix images size and plural

@@
 
-Finally, \framework uses a last parameter $\delta\in[0,1]$ that represents how much the waves orientation is constant ($\delta = 0$) or not ($\delta = 1$) (see Figure \ref{fig:dispersionEffect}).
+Finally, \framework uses a last parameter $\delta\in[0,1]$ that represents how much the waves orientation is constant ($\delta = 0$) or not ($\delta = 1$) (see Figure~\ref{fig:dispersionEffect}).
 Its meaning is as follows:
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/seaDispersion0SeaState5-contrast.png}
+		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion0SeaState5-contrast.png}
 		\label{fig:dispersion0}
 	}
-	\hspace{1cm}
+	\hfill
 	\subfloat[Using $\delta = 1$.]
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/seaDispersion1SeaState5-contrast.png}
+		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion1SeaState5-contrast.png}
 		\label{fig:dispersion1}
@@
 \Framework uses the more efficient solution proposed by Dupuy and Bruneton (2012) \cite{dupuyRealtimeAnimationRendering2012}.
-It consists in dividing the spectrum into $c$ layers or \emph{cascades} of different length.
+It consists in dividing the spectrum into $c$ layers or \emph{cascades} of different lengths.
 Each of these layers represents a different wavelength in Equation~\ref{equa:heightWater}.

### 2024-11-14 paper.tex
change term

@@
 
-In \framework, the number of IFFT applications is divided by two using Theorem \ref{theorem:ifft} on Hermitian matrices\footnote{A matrix $X=(x_{n,m})$ is Hermitian if it is equal to its own conjugate transpose: $X = X^{\dagger}$ or $x_{n,m}=x^*_{m,n}$.} (see annex \ref{sec:annexe-2IFFT} for its demonstration).
+In \framework, the number of IFFT applications is divided by two using Theorem \ref{theorem:ifft} on Hermitian matrices\footnote{A matrix $X=(x_{n,m})$ is Hermitian if it is equal to its own conjugate transpose: $X = X^{\dagger}$ or $x_{n,m}=x^*_{m,n}$.} (see annex \ref{sec:annexe-2IFFT} for its proof).
 \begin{theorem}

### 2025-01-10 paper.tex
Add option for inverse forward search with tex
Remove an incoherent name for caption fig 3
Fixed typo on non submerged part

@@
+% !TEX options=-synctex=-1 %for inverse forward search 
 \documentclass[final]{jcgt}
@@
 	}
-	\caption{Effect of the dispersion parameter on sea state $5$ on the Beaufort scale.}
+	\caption{Effect of the parameter $\delta$ on sea state $5$ on the Beaufort scale.}
 	\label{fig:dispersionEffect}
@@
 		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to $c_i$ too;
-		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the submerged part;
+		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the unsubmerged part;
 		and finally the gravity force \textcolor{greenF}{$F_g$} in \textcolor{greenF}{green}, applied to the center of gravity.

### 2025-01-10 paper.tex
Clarify fig 11

@@
 \begin{equation}
+	\label{equa:mask_full}
 	h(\mathbf x)=b_w\left(f(x)+g(z)\right).
@@
 
-Figure \ref{fig:mask} shows some views of the application of this interactive waves process for a simple rectangular boat.
+Figure \ref{fig:mask} shows some views of the application of this interactive waves process.
 
@@
 
-	\subfloat[A theoretical view of the water height function $h(\mathbf x,t)$ that will be applied to a simple rectangle.]
+	\subfloat[A theoretical view of the height of the mask, with input a simple rectangle.]
 	{
@@
 	\hspace{1cm}
-	\subfloat[The mask of a simple boat encoded in a single channel texture.]
+	\subfloat[The height of the mask of a simple boat encoded in a single channel texture.]
 	{
@@
 	\hspace{1cm}
-	\subfloat[The mask of a simple boat in a 3D scene with a very low sea state.]
+	\subfloat[The height of the mask of a simple boat in a 3D scene with a very low sea state.
+Boat rendering and wave propagation has been disabled to only display the initial step of the height of the mask.]
 	{
@@
 	}
-	\caption{Three different views of the mask.}
+	\caption{Three different views of the height of the mask defined by Equation \ref{equa:mask_full}.}
 	\label{fig:mask}

### 2025-01-10 paper.tex
Fixed sign typos in appendix A
Reordered appendix A for velocity calculation to improve coherency

@@
 \begin{equation}
-	\tilde{\phi}(y, t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
+	\tilde{\phi}(y, t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
 \end{equation}
@@
 	\Delta \phi\nonumber= & \left(-k_x^2 E(\mathbf{k},y) -k_z^2E(\mathbf{k},y) + E''(\mathbf{k},y)\right)\nonumber                                                                                      \\
-	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
+	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
 	\exp\left( i \mathbf{k}\cdot \mathbf{x}\right)\nonumber                                                                                                                                             \\
 	=                     & \left(E''(\mathbf{k},y)-k^2E(\mathbf{k},y)\right) \nonumber                                                                                                                 \\
-	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) \exp\left(i \mathbf{k}\cdot \mathbf{x}\right)
+	                      & \times \frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) \exp\left(i \mathbf{k}\cdot \mathbf{x}\right)
 \end{align}
@@
 	\frac{\partial\phi(\mathbf{x},y,t)}{\partial y}= &
-	\frac{\partial E(\mathbf{k},y)}{\partial y}\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)               \\
+	\frac{\partial E(\mathbf{k},y)}{\partial y}\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)               \\
 	=                                                & kC_1\sinh(ky+C_2)                                                                                                                                                      \\
-	                                                 & \times\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)
+	                                                 & \times\frac{ig}{\omega(k)}\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)\exp\left(i\mathbf{k}\cdot\mathbf x\right)
 \end{align*}
@@
 
-This expression is valid for height $y\leq0$, but not usable above for $y>0$.
+Note that while the third equation of system \ref{equa:systemTessendorf} is not used here, it can help to deduce the \emph{dispersion relation}.
+
+Equation~\ref{equa:attenuationDeep1} and the definition of potential velocity $\nabla\phi=\mathbf{v}$ lead to the velocity $\mathbf{v}$. First, function $\tilde{\phi}$ is rewritten as follows:
+\begin{equation}
+	\tilde{\phi}(y,t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
+\end{equation}
+Then the gradient of $\phi$ is as follows:
+\begin{equation}
+	\nabla \phi = \begin{pmatrix}
+		v_x \\
+		v_y \\
+		v_z
+	\end{pmatrix} =\tilde{\mathbf{v}}(y,\mathbf{k},t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
+\end{equation}
+where $\tilde{\mathbf{v}}(\mathbf{k},t,y)$ is the vector equal to:
+\begin{equation}
+	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
+	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
+	\left[
+		\begin{array}{c}
+			\displaystyle\frac{-k_xg}{\omega(k)} \\
+			\displaystyle i\omega(k)             \\
+			\displaystyle\frac{-k_zg}{\omega(k)}
+		\end{array}
+		\right],
+\end{equation}
+
+Finally, reintroducing a sum on a set of waves vectors $\mathbf{k}$ leads to:
+\begin{equation}
+	\nabla \phi =
+	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k},t, y)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right).
+\end{equation}
+
+Note that the expression of $E\left(\mathbf{k},y\right)$ (see Equation \ref{equa:attenuationDeep1}) is valid for height $y\leq0$, but not usable above for $y>0$.
 Several possibilities exist to overcome this problem:
@@
 	\end{cases}
-\end{equation}
-
-Note that while the third equation of system \ref{equa:systemTessendorf} is not used here, it can help to deduce the \emph{dispersion relation}.
-
-Equation~\ref{equa:attenuationDeep1} and the definition of potential velocity $\nabla\phi=\mathbf{v}$ lead to the velocity $\mathbf{v}$. First, function $\tilde{\phi}$ is rewritten as follows:
-\begin{equation}
-	\tilde{\phi}(y,t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
-\end{equation}
-Then the gradient of $\phi$ is as follows:
-\begin{equation}
-	\nabla \phi = \begin{pmatrix}
-		v_x \\
-		v_y \\
-		v_z
-	\end{pmatrix} =\tilde{\mathbf{v}}(y,\mathbf{k},t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
-\end{equation}
-where $\tilde{\mathbf{v}}(\mathbf{k},t,y)$ is the vector equal to:
-\begin{equation}
-	\tilde{\mathbf{v}}(y,\mathbf{k},t) = \exp(ky)\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
-	\begin{pmatrix}
-		\frac{-k_xg}{\omega(k)} \\
-		i\omega(k)              \\
-		\frac{-k_zg}{\omega(k)}
-	\end{pmatrix}.
-\end{equation}
-Finally, reintroducing a sum on a set of waves vectors $\mathbf{k}$ leads to:
-\begin{equation}
-	\nabla \phi =
-	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k},t, y)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
 \end{equation}

### 2025-01-10 paper.tex
Fixed case in equation 56
Make the order of parameters more consistent k,x < y < t

@@
 	\label{equa:heightWater}
-	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(t, \mathbf{k}) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
+	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(\mathbf{k}, t) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
 \end{equation}
@@
 	\label{equa:heightWaterBis}
-	\tilde{h}(t, \mathbf{k}) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
+	\tilde{h}(\mathbf{k}, t) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
 \end{equation}
@@
 \begin{equation}\label{equa:hori-displacement}
-	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(t, \mathbf{k})\exp\left(i\mathbf{k}\cdot\mathbf x\right).
+	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right).
 \end{equation}
@@
 \begin{align*}
-	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
-	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
-	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
-	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
-	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right).
+	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
+	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
+	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
+	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
+	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right).
 \end{align*}
@@
 	\mathbf v(\mathbf x,y,t)=
-	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
-\end{equation}
-where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
-\begin{equation}
-	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
+	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
+\end{equation}
+where the vector $\tilde{\mathbf{v}}(\mathbf{k}, y, t)$ is given as follows:
+\begin{equation}
+	\tilde{\mathbf{v}}\left(\mathbf{k}, y, t\right)=%
 	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
@@
 \begin{equation}
-	\phi(\mathbf{x},y,t) = \tilde{\phi}(y, t,\mathbf{k})\exp(i \mathbf{k}\cdot \mathbf{x}).
+	\phi(\mathbf{x},y,t) = \tilde{\phi}(\mathbf{k}, y, t)\exp(i \mathbf{k}\cdot \mathbf{x}).
 \end{equation}
 Notice that this expression is convenient as it can be computed with an inverse Fast Fourier Transform.
-The function $\tilde{\phi}(y, t,\mathbf{k})$ is defined as follows:
-\begin{equation}
-	\tilde{\phi}(y, t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
+The function $\tilde{\phi}(\mathbf{k}, y, t)$ is defined as follows:
+\begin{equation}
+	\tilde{\phi}(\mathbf{k}, y, t)=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)
 \end{equation}
@@
 	E(\mathbf{k},y) & = \frac{\cosh(ky+kH)}{\cosh(kH)}\nonumber                            \\
-	                & = \frac{\cosh(ky)\cosh(kH) + \sinh(kY)\sinh(kH)}{\cosh(kH)}\nonumber \\
+	                & = \frac{\cosh(ky)\cosh(kH) + \sinh(ky)\sinh(kH)}{\cosh(kH)}\nonumber \\
 	                & = \cosh(ky) + \sinh(ky)\tanh(kH)
@@
 \begin{equation}
-	\tilde{\phi}(y,t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
+	\tilde{\phi}(\mathbf{k}, y, t)=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
 \end{equation}
@@
 		v_z
-	\end{pmatrix} =\tilde{\mathbf{v}}(y,\mathbf{k},t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
-\end{equation}
-where $\tilde{\mathbf{v}}(\mathbf{k},t,y)$ is the vector equal to:
-\begin{equation}
-	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
+	\end{pmatrix} =\tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
+\end{equation}
+where $\tilde{\mathbf{v}}(\mathbf{k}, y, t)$ is the vector equal to:
+\begin{equation}
+	\tilde{\mathbf{v}}\left(\mathbf{k}, y, t\right)=%
 	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
@@
 	\nabla \phi =
-	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k},t, y)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right).
+	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right).
 \end{equation}

### 2025-01-10 paper.tex
Clarify section about relation of dispersion in appendix

@@
 
-Note that while the third equation of system \ref{equa:systemTessendorf} is not used here, it can help to deduce the \emph{dispersion relation}.
+The third equation of the system \ref{equa:systemTessendorf} gives the following equality:
+\begin{equation}
+\label{equa:derivHeqDerivPhi}
+\frac{\partial h}{\partial t}(\mathbf{k},t)=\frac{\partial\phi}{\partial y}(\mathbf{k}, 0, t)
+\end{equation}
+Once again, we only considered one wave vector $\mathbf{k}$ as Equations are all linear, in consequence Equation \ref{equa:derivHeqDerivPhi} becomes:
+
+\begin{align}
+\frac{\partial \tilde{h}}{\partial t}(\mathbf{k},t) - \frac{\tilde{\partial}\phi}{\partial y}(\mathbf{k}, 0, t) &=0 \nonumber\\
+i\omega(k)\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) - \frac{igk}{\omega(k)} E(\mathbf{k},0)\nonumber\\
+ \times \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) &= 0 \nonumber\\
+ \omega(k)- \frac{gk}{\omega(k)} &= 0 \nonumber\\
+ \omega(k) = \sqrt{gk} \label{equa:relationDispersionDeduce}
+\end{align}
+
+This last Equation \ref{equa:relationDispersionDeduce} gives the expression of the dispersion relation used in the Section \ref{sec:oceanWaves}.
 

### 2025-01-10 paper.tex
We have corrected the index mistake v_z and not v_y
Remove excess term
Add "coupling" in abstract
Change the term used in scientific computing with cg to be more coherent with our article and JCGT

@@
 	calculating the real-time velocity of ocean fluids at any depth; 
-	computing the input of the fluid to solid algorithm.
+	computing the input of the fluid to solid coupling algorithm.
 \end{abstract}
@@
 \label{subsec:intro-context}
-Simulating the dynamics of an ocean and its surrounding elements remains an important objective in scientific computing.
+Simulating the dynamics of an ocean and its surrounding elements remains an important objective in computer graphics.
 As all the oceans encompass a substantial portion of the Earth's surface, understanding their behavior is of paramount importance in scientific, industrial, entertainment and military fields.
@@
 \begin{equation}
-	\mathbf{x^{\prime}} = \mathbf x + dt\left[\begin{array}{c}v_x\\ v_y\end{array}\right]
+	\mathbf{x^{\prime}} = \mathbf x + dt\left[\begin{array}{c}v_x\\ v_z\end{array}\right]
 	\label{equa:freeSurfaceFollowWater1}
@@
 	\begin{aligned}
-		\frac{h(\mathbf x, t + dt) - h(\mathbf x, t)}{dt} + v_x\frac{\partial h(\mathbf x, t)}{\partial x} + v_z\frac{\partial h(\mathbf x, t)}{\partial z} & = \frac{h(\mathbf x, t)}{dt} + v_y \\
+		\frac{h(\mathbf x, t + dt) - h(\mathbf x, t)}{dt} + v_x\frac{\partial h(\mathbf x, t)}{\partial x} + v_z\frac{\partial h(\mathbf x, t)}{\partial z} & = v_y \\
 		\frac{\partial h}{\partial t} + v_x\frac{\partial h}{\partial x} + v_z\frac{\partial h}{\partial z}                                                 & = v_y

### 2025-01-10 paper.tex
Put related work in a dedicated section
Use gravitational acceleration instead of standard gravity
Remove a doublon

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Related works}
+\section{Related works}
 \label{sec:relatedwork}
@@
 
-\subsubsection{Ocean free surface simulation}
+\subsection{Ocean free surface simulation}
 \label{subsubsec:oceanFreeSurfaceSim}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Fluid to solid coupling}
+\subsection{Fluid to solid coupling}
 \label{subsubsec:fluidToSolidCoupling}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Solid to fluid coupling}
+\subsection{Solid to fluid coupling}
 \label{subsubsec:solidToFluidCoupling}
@@
 	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
-	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
+	\item The scalar $g=9.80665~m.s^{-2}$ denotes the gravitational acceleration, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
 \end{itemize}
@@
 \url{https://www.digitaldutch.com/atmoscalc/table.htm}} in $kg.m^{-3}$;
-	\item $\rho_a$ is the density of air in $kg.m^{-3}$;
 	\item $C_d^a\in\mathbb{R}^3$ is the drag coefficient of the body in the air, and is given by the user.

### 2025-01-10 paper.tex
Fixed 2 typos

@@
 \Framework uses a simple model that tries to reproduce parts of the idea of Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
-By construction of the grid $Z$, $M$ has is bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
+By construction of the grid $Z$, $M$ has a bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
 Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.
@@
 This makes a total of $4135$ triangles. 
-Each of these ten solids uses its own grid for the fluid to solid coupling FDM. 
+Each of these ten solids uses its own grid for the solid to fluid coupling FDM. 
 Two captures of the second experiment can be seen in the bottom row of \figurename~\ref{fig:experiments}.

### 2025-01-10 paper.tex
Use yuksel 2007 in reference
Change tessendorf references

@@
 \begin{itemize}
-	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
+	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2001) \cite{tessendorfSimulating2001} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
 	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
@@
 J.
-Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} presented what could be considered a seminal work in real-time ocean simulation.
+Tessendorf (2001) \cite{tessendorfSimulating2001} presented what could be considered a seminal work in real-time ocean simulation.
 His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solution to an approximation of the Navier-Stokes equations on a 2D height field.
@@
 He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
-Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
+Yuksel (2007) \cite{yuksel2007} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
 Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
@@
 \label{subsec:waterHeight}
-Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
+Tessendorf (2001) \cite{tessendorfSimulating2001} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
 \begin{equation}
@@
 Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
-In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
+In its original paper, Tessendorf (2001) \cite{tessendorfSimulating2001} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
 \begin{equation}
@@
 \Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
-As proposed in Sections 4.5 and 4.6 of Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999}, it is calculated as follows:
+As proposed in Sections 4.5 and 4.6 of Tessendorf (2001) \cite{tessendorfSimulating2001}, it is calculated as follows:
 \begin{equation}\label{equa:hori-displacement}

### 2025-01-10 paper.tex
Rephrase a sentence

@@
 Computing the velocity for all the points of a 3D grid, or even only the points where it must be evaluated for fluid to solid coupling (Section~\ref{subsubsec:fluidToSolidCoupling}) is not possible in real-time, as it involves a large to huge number of IFFT.
-Unfortunately, the dependency of the attenuation (Equation \ref{equa:attenuationDeep}) on both the wave vectors and $y$ makes the depth dependency of the velocity fundamentally tied to the wave vector, preventing from any simplifications to reduce the needed calculations.
+Unfortunately, the dependency of the attenuation (Equation \ref{equa:attenuationDeep}) on both the wave vectors and $y$ makes the depth dependency of the velocity fundamentally tied to the wave vector, preventing any simplifications that could reduce the required calculations.
 

### 2025-01-10 paper.tex
Add a missing word
Remove the +1 to avoid confusion

@@
 This distribution is designed based on the following considerations.
-First, while the velocity decays exponentially, it relies on a logarithmic similar to $y_i\mapsto \ln(|y_i|+1)$.
+First, while the velocity decays exponentially, it relies on a logarithmic function similar to $y_i\mapsto \ln(|y_i|+1)$.
 Since this function transforms the negative $y_i$ into a positive one, the result is multiplied by the sign of $y_i$.
-Also, this function decreases slowly (actually with $\ln(250+1)=5.5$), hence the logarithmic distribution uses the square of $|y_1|$.
+Also, this function decreases slowly (actually with $\ln(250)=5.5$), hence the logarithmic distribution uses the square of $|y_1|$.
 Lastly, the parameter $\alpha$ controls the “squeezing” in the center, while the parameter $\beta$ controls the deepest value of $l_d$.

### 2025-01-15 paper.tex
Merge water drag and air drag forces

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Water drag}
-\label{subsubsec:waterDrag}
-\Framework uses a drag force depending on the relative speed, as in many previous works \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014,kernerWaterInteractionModel2016}.
-The water drag force is calculated for each triangle $T_i^s$ of the submerged body as follows:
-\begin{equation}
-	\mathbf{F_w}= -\frac{1}{2} C_d^w \rho_w A_i^{\perp} \left\Vert\mathbf{v_{i_{rel}}^w}\right\Vert\mathbf{v_{i_{rel}}^w},
+\subsubsection{Drag Force}
+\label{subsubsec:dragForce}
+\Framework calculates drag forces based on the relative velocity between the fluid medium (water or air) and the submerged or exposed portion of the body.
+This approach is inspired by previous works \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014,kernerWaterInteractionModel2016}.
+
+The drag force for each triangle $T_i$ of the body is computed as:
+\begin{equation}
+	\mathbf{F_d}= -\frac{1}{2} C_d \rho A_i^{\perp} \left\Vert\mathbf{v_{i_{rel}}}\right\Vert\mathbf{v_{i_{rel}}},
 \end{equation}
@@
 \begin{itemize}
-	\item $\mathbf{v_{i_{rel}}^w}=\mathbf{v_i}-\mathbf{v_i^w}$ is the relative velocity in $m.s^{-1}$ of the triangle $T_i^s$, equals to the difference between its velocity $\mathbf{v_i}$ and the velocity $\mathbf{v_{i}^w}$ of the water at its center;
-	\item $A_i^{\perp}$ is the projected area in $m^2$ of the triangle $T_i^s$ with respect to the relative velocity unit direction (\ie the vector $\mathbf{v_{i_{rel}}^w}/\left\Vert\mathbf{v_{i_{rel}}^w}\right\Vert$);
-	\item $\rho_w$ is still the density of water in $kg.m^{-3}$;
-	\item $C_d^w\in\mathbb{R}^3$ is the drag coefficient of the body in the water, and is given by the user.
+	\item $\mathbf{v_{i_{rel}}}=\mathbf{v_i}-\mathbf{v_{i}^{m}}$ is the relative velocity in $m.s^{-1}$ of the triangle $T_i$, equal to the difference between its velocity $\mathbf{v_i}$ and the velocity $\mathbf{v_{i}^{m}}$ of the fluid medium (water or air) at its center;
+	\item $A_i^{\perp}$ is the projected area in $m^2$ of the triangle $T_i$ with respect to the relative velocity unit direction (\ie the vector $\mathbf{v_{i_{rel}}}/\left\Vert\mathbf{v_{i_{rel}}}\right\Vert$);
+	\item $\rho$ is the density of the fluid medium in $kg.m^{-3}$ ($\rho_w$ as defined in Section~\ref{subsubsec:buoyancy} for water and $1.204~kg.m^{-3}$  for air\footnote{According to the international standard atmosphere, the air density at sea level for $20$°C is $1.204~kg.m^{-3}$ cf.
+\url{https://www.digitaldutch.com/atmoscalc/table.htm}});
+	\item $C_d$ is the drag coefficient of the body in the fluid medium, specified by the user.
 \end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Air drag}
-\label{subsubsec:airDrag}
-The force corresponding to the air drag is designed similarly to the water drag one, but using the not submerged part $\left\lbrace T_i^{\neg s}\right\rbrace$ of the body.
-For each triangle $T_i^{\neg s}$, it corresponds to the following expression:
-\begin{equation}
-	\mathbf{F_a}= -\frac{1}{2} C_d^a\rho_a A_i^{\perp}\left\Vert\mathbf{v_{i_{rel}}^a}\right\Vert \mathbf{v_{i_{rel}}^a}
-\end{equation}
-where:
-\begin{itemize}
-	\item $\mathbf{v_{i_{rel}}^a}=\mathbf{v_i}-\mathbf{v_{i}^a}$ is the relative velocity in $m.s^{-1}$ of the triangle $T_i^{\neg s}$, equals to the difference between its velocity $\mathbf{v_i}$ and the velocity $\mathbf{v_{i}^a}$ of the air at its center;
-	\item $A_i^{\perp}$ is the projected area in $m^2$ of the triangle $T_i^{\neg s}$ with respect to the relative velocity unit direction (\ie the vector $\mathbf{v_{i_{rel}}^a}/\left\Vert\mathbf{v_{i_{rel}}^a}\right\Vert$);
-	\item $\rho_a$ is the density of air\footnote{According to the international standard atmosphere, the air density at sea level for $20$°C is $1.204~kg.m^{-3}$ cf.
-\url{https://www.digitaldutch.com/atmoscalc/table.htm}} in $kg.m^{-3}$;
-	\item $C_d^a\in\mathbb{R}^3$ is the drag coefficient of the body in the air, and is given by the user.
-\end{itemize}
+For the submerged part $\left\lbrace T_i^s\right\rbrace$, the medium is water, while for the exposed part $\left\lbrace T_i^{\neg s}\right\rbrace$, the medium is air.
 
@@
 	\item Computing the buoyancy force for each body (Section \ref{subsubsec:buoyancy});
-	\item Computing the two drag forces for each triangle, and summing them for each body (Sections \ref{subsubsec:waterDrag} and \ref{subsubsec:airDrag});
+	\item Computing the two drag forces for each triangle, and summing them for each body (Section \ref{subsubsec:dragForce});
 	\item Applying the forces in the physical integration scheme.

### 2025-01-17 paper.tex
Change unit and inline equation

@@
 
-To make $P(d)$ and $A(d)$ having the same order of magnitude, the units are the \textbf{decisecond} for performance and the \textbf{meter per seconds} for accuracy (see figure \ref{fig:interpolationCriteria}).
-Moreover, the performance being the more important for real-time process, the following arbitrary choice are made for the coefficients of $J$:
-\begin{equation}
-	\begin{cases}
-		\alpha = 1 \\
-		\beta = 2
-	\end{cases}
-\end{equation}
+To make $P(d)$ and $A(d)$ having the same order of magnitude, the units are the \textbf{seconds} for performance and the \textbf{meter per seconds} for accuracy (see figure \ref{fig:interpolationCriteria}).
+Moreover, the performance being the more important for real-time process, the following arbitrary choice are made for the coefficients of $J$: $\alpha = 10$ and $\beta = 2$.
 

### 2025-01-20 paper.tex
Change notation of dx to delta to avoid confusion

@@
 It uses a square simulation zone $Z$ of dimension $L\times L$ around a given mesh $M$.
-This zone is regularly discretized with a $dx$ step.
+This zone is regularly discretized with a $\delta$ step.
 Time is also regularly discretized with a time step $dt$.
 
-In the FDM simulation, \framework denotes by $h_{i,j}^n$ the water height $h(i\times dx,j\times dx,n \times dt)$ for discrete location $(i,j)\times dx$ and time $n\times dt$.
+In the FDM simulation, \framework denotes by $h_{i,j}^n$ the water height $h(i\times \delta,j\times \delta,n \times dt)$ for discrete location $(i,j)\times \delta$ and time $n\times dt$.
 This leads to the following explicit scheme:
@@
 \begin{equation}
-	a=\frac{c^2dt^2}{dx^2}.
+	a=\frac{c^2dt^2}{\delta^2}.
 \end{equation}
@@
 	\begin{cases}
-		k = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{dx}\right\rfloor    \\
-		l = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{dx}\right\rfloor    \\
-		o = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{dx}\right\rfloor \\
-		p = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{dx}\right\rfloor
+		k = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{\delta}\right\rfloor    \\
+		l = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{\delta}\right\rfloor    \\
+		o = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{\delta}\right\rfloor \\
+		p = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{\delta}\right\rfloor
 	\end{cases}
@@
 \begin{equation}
-	\frac{c^2dt^2}{dx^2}\leq 0.5
+	\frac{c^2dt^2}{\delta^2}\leq 0.5
 \end{equation}
 The parameter $dt$ is considered fixed throughout the simulation.
-One contribution of this article is to adapt the simulation's parameters $c$ and $dx$ to the CFL, as presented below.
-
-Preliminary tests have shown that $dx$ must be small for a low velocity $\mathbf{v_M}(t)$ of $M$.
-Moreover, $dx$ must be large enough for $Z$ to encompass $M$ and the waves generated by $M$.
+One contribution of this article is to adapt the simulation's parameters $c$ and $\delta$ to the CFL, as presented below.
+
+Preliminary tests have shown that $\delta$ must be small for a low velocity $\mathbf{v_M}(t)$ of $M$.
+Moreover, $\delta$ must be large enough for $Z$ to encompass $M$ and the waves generated by $M$.
 The size of $M$ is defined from its bounding box $(b_x, b_y, b_z)$ as $S(M)=\max(b_x, b_z)$.
-\Framework lets the user define the bounds $dx_{min}$ and $dx_{max}$ so that $Z$ is at least twice as large as $M$.
-Then $dx$ is defined as follows:
-\begin{equation}
-	dx =
+\Framework lets the user define the bounds $\delta_{min}$ and $\delta_{max}$ so that $Z$ is at least twice as large as $M$.
+Then $\delta$ is defined as follows:
+\begin{equation}
+	\delta =
 	\begin{cases}
@@
 \end{equation}
-Finally $dx$ is clamped between $dx_{min}$ and $dx_{max}$.
-The value of $c$ is deduced from $dx$ and $dt$ to validate the CFL as follows:
-\begin{equation}
-	c = \sqrt{0.49} \frac{dx}{dt}
+Finally $\delta$ is clamped between $\delta_{min}$ and $\delta_{max}$.
+The value of $c$ is deduced from $\delta$ and $dt$ to validate the CFL as follows:
+\begin{equation}
+	c = \sqrt{0.49} \frac{\delta}{dt}
 \end{equation}

### 2025-01-20 paper.tex
Change notation damping formulation

@@
 \end{equation}
+
 The Dirichlet boundary condition gives $h_{i,j}^n=0$ for $i$ or $j$ being on the boundary of $Z$.
 Since a new wave is generated at each time step, a damping effect is added to make them disappear gradually.
-Hence, the right part of Equation (\ref{eq:fdm_scheme}) is multiplied by the \textbf{damping factor} $d(t)=d^n$.
-This damping factor is linearly related to the velocity of the mesh $v_M(t)$ to ensure numerical stability as follows:
-\begin{equation}
-	d(t) =
-	\begin{cases}
-		\left(1-\frac{||\mathbf{v_M}(t)||}{v_{max}}\right)d_0 + \frac{||\mathbf{v_M}(t)||}{v_{max}}d_{max} & \text{for~~}||\mathbf{v_M}(t)||\leq v_{max}, \\
-		d_{max}                                                                          & \text{else}
-	\end{cases}
-\end{equation}
-where $d_{max}$ is the maximum damping value obtained when the body's velocity is above a given threshold $v_{max}$,
-and $d_0$ is the damping for a null velocity.
+Hence, the right part of Equation (\ref{eq:fdm_scheme}) is multiplied by the \textbf{damping factor} $d(t) = d^n$.
+
+This damping factor is linearly related to the velocity of the mesh $\mathbf{v_M}(t)$ to ensure numerical stability and is defined as:
+\begin{equation*}
+u = \text{clamp}\left(\frac{\|\mathbf{v_M}(t)\|}{v_{max}}, 0, 1\right),
+\end{equation*}
+\begin{equation}
+d(t) = \text{lerp}(d_0, d_{max}, u),
+\end{equation}
+where:
+\begin{itemize}
+	\item $\text{clamp}(x, 0, 1)$ limits the value of $x$ to the range $[0, 1]$;
+	\item $\text{lerp}(a, b, t) = (1-t)a + tb$ is the linear interpolation function;
+	\item $v_{max}$ is the velocity threshold, above which the damping factor equals its maximum value $d_{max}$;
+	\item $d_0$ is the damping value for a null velocity.
+\end{itemize}
+
 For a plausible appearance, \framework uses the following values by default:

### 2025-01-20 paper.tex
Add a figure to illustrate the relationship between indices

@@
 \end{equation}
-where $\mathbf p(t)=\left[p_x(t)\ p_z(t)\right]^T$ is the position of the mesh $M$ at time $t$, $k$ and $l$ (respectively $o$ and $p$) are the shifted indices defined by the translation of $M$ at times $t+dt$ from times $t$ (respectively from times $t-dt$).
+
+% Pattern Info
+\begin{figure}[!h]
+	\centering
+	\input{Tikz/mdfIndices.tikz}
+	\caption{An example of the relationship between the indices $(i,j)$, $(k,l)$, and $(o,p)$ in the finite-difference method grid translation, showing how the grid indices are affected by the motion of the body $M$ at different time steps $t$, $t-dt$, and $t+dt$}
+	\label{fig:mdfIndices}
+\end{figure}
+
+where $\mathbf p(t)=\left[p_x(t)\ p_z(t)\right]^T$ is the position of the mesh $M$ at time $t$, $k$ and $l$ (respectively $o$ and $p$) are the shifted indices (see Figure \ref{fig:mdfIndices}) defined by the translation of $M$ at times $t+dt$ from times $t$ (respectively from times $t-dt$).
 Then, the integration scheme can be rewritten as follows:

### 2025-01-20 paper.tex
Fixed the typo
Fixed another typo

@@
 The FDM needs the height of the vertices below the body.
-This is computed from the mask, composed of the vertices of $Z$ contained into one intersection polygon at least.
+This is computed from the mask, composed of the vertices of $Z$ contained in one of the intersection polygons at least.
 This is done using the \textbf{Point in Polygon} strategy described in Haines (1994) \cite{hainesPointPolygonStrategies1994} and later in Cormen (2001) \cite{CormenIntroductionAlgorithms2001}.
@@
 To reduce the computation times, this strategy is applied to the vertices $v$ included into the bounding box of the intersections polygons.
-Since the number of vertices $v$ of $Z$ is at least twice as the size of $M$, this simple technic reduces efficiently the number of ray to polygon calculations.
+Since the number of vertices $v$ of $Z$ is at least twice as the size of $M$, this simple method reduces efficiently the number of ray to polygon calculations.
 

### 2025-01-20 paper.tex
Fixed again another typo

@@
 \Framework uses a simple model that tries to reproduce parts of the idea of Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
-By construction of the grid $Z$, $M$ has a bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
+By construction of the grid $Z$, $M$ has its bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
 Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.

### 2025-01-20 paper.tex
Add precision on starboard

--- 2025-01-20 paper.tex
@@
 \Framework uses a simple model that tries to reproduce parts of the idea of Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
-By construction of the grid $Z$, $M$ has a bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
+By construction of the grid $Z$, $M$ has its bow oriented on the positive side of the $z$ axis, and its starboard (right-hand side when facing the bow) on the positive side of the $x$ axis.
 Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.

### 2025-01-28 paper.tex
Clarify c choice
Fixed typo

@@
 \begin{equation}
+\label{equa:cfl}
 	\frac{c^2dt^2}{\delta^2}\leq 0.5
@@
 Finally $\delta$ is clamped between $\delta_{min}$ and $\delta_{max}$.
-The value of $c$ is deduced from $\delta$ and $dt$ to validate the CFL as follows:
-\begin{equation}
-	c = \sqrt{0.49} \frac{\delta}{dt}
-\end{equation}
+The value of $c$ is deduced from $\delta$ and $dt$ to ensure that the inequality in Equation \ref{equa:cfl} is always satisfied:
+\begin{equation}
+	c = \sqrt{0.49} \frac{\delta}{dt}.
+\end{equation}
+This choice of $c$ ensures numerical stability.
 
@@
 Hence, for a body $M$ the wave generation relies on two different steps:
-First, a mask is defined from this list of 3D points, corresponding to the vertices of $Z$ contained into one intersected polygons.
+First, a mask is defined from this list of 3D points, corresponding to the vertices of $Z$ contained in one intersected polygons.
 Second, the height $h$ of each point in the mask is determined using an ad hoc function.

### 2025-01-28 paper.tex
Remove CLRS (Cormen 2001) reference

@@
 This is computed from the mask, composed of the vertices of $Z$ contained in one of the intersection polygons at least.
-This is done using the \textbf{Point in Polygon} strategy described in Haines (1994) \cite{hainesPointPolygonStrategies1994} and later in Cormen (2001) \cite{CormenIntroductionAlgorithms2001}.
+This is done using the \textbf{Point in Polygon} strategy described in Haines (1994) \cite{hainesPointPolygonStrategies1994}.
 

### 2025-01-29 paper.tex
Clarify the volume computation and the assumptions that were made on the vessels

@@
 Moreover, to remain real-time \framework uses forces computed from simple geometric and physical parameters only.
+To maintain realistic physics after these discrete approximations and those described in subsequent sections, it is important to use a closed mesh, uniformly triangulated.
+In addition, for meshes representing vessels, ensuring symmetry is crucial to maintain stability.
 
@@
 The buoyancy calculation presented in Section \ref{subsubsec:buoyancy} requires the full submerged volume $v_w$ of the mesh.
-\Framework uses the prism approximation proposed by Bajo \etal (2020) \cite{bajoRealisticBuoyancyModel2020a}.
-The principles of their method are as follows: from each submerged triangle $T_i^s$ of the body, a prism is build up to the water surface; hence, each prism volume can be calculated, and so the full submerged volume $v_w$ by adding all of them.
+\Framework uses the prism approximation similarly to the approximation proposed by Bajo \etal (2020) \cite{bajoRealisticBuoyancyModel2020a}. 
+While Bajo \etal compute the submerged volume at the texel level, the same principle can be applied using triangles instead of texels using as they did the divergence theorem. 
+Specifically, $v_w$ is approximated as the sum of the prisms with bases corresponding to the triangle areas $A_i$ and heights defined by the distance between the triangle centers $p_i$ and the water surface $h(x,z,t)$.
+Each term in the sum is weighted by the vertical component of the triangle normal $n_i^y$:
+\begin{equation}
+v_w = \sum_{T_i^s} (h(x,z,t) - p_i) A_i n_i^y.
+\end{equation}
+Note that this formula is correct only under the assumption that the mesh is closed.
 

### 2025-01-29 paper.tex
Precise that we are using single precision float number

@@
 Two experiments with different solids have been conducted to demonstrate the efficiency of \framework.
-These experiments ran on a \href{https://www.intel.fr/content/www/fr/fr/products/sku/199316/intel-core-i710700-processor-16m-cache-up-to-4-80-ghz/specifications.html}{Intel Core i7-10700} and a \href{https://www.nvidia.com/fr-fr/geforce/graphics-cards/compare/?section=compare-20}{Nvidia GeForce RTX 2070 Super}.
+These experiments ran on a \href{https://www.intel.fr/content/www/fr/fr/products/sku/199316/intel-core-i710700-processor-16m-cache-up-to-4-80-ghz/specifications.html}{Intel Core i7-10700} and a \href{https://www.nvidia.com/fr-fr/geforce/graphics-cards/compare/?section=compare-20}{Nvidia GeForce RTX 2070 Super} using single-precision floating-point numbers.
 The ocean was simulated with three cascades using a resolution of $256\times265$.
 Moreover, the resolution of FDM grids is set to $512\times 512$. 
-
 \begin{figure}[htbp]

### 2025-01-29 paper.tex
Simplify water displacement expression

@@
 
-\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
+\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x(\mathbf x,t)\ D_z(\mathbf x,t)\right]^T$ of seawater to simulate more realistic choppy waves.
 As proposed in Sections 4.5 and 4.6 of Tessendorf (2001) \cite{tessendorfSimulating2001}, it is calculated as follows:
 \begin{equation}\label{equa:hori-displacement}
-	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right).
+	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \mathbf{\tilde{D}}(\mathbf k,t)\exp\left(i\mathbf{k}\cdot\mathbf x\right)
+\end{equation}
+
+where
+\begin{equation}\label{equa:hori-displacement-k}
+	\mathbf{\tilde{D}}(\mathbf k,t) = \left[\tilde{D}_x(\mathbf k,t)\ \tilde{D}_z(\mathbf k,t)\right]^T = \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t).
 \end{equation}
@@
 \begin{align*}
-	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
-	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
-	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
-	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
-	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right).
+	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -ik_x \tilde{D}_x(\mathbf k,t)\exp\left(i\mathbf{k}\cdot\mathbf x\right),  \\
+	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -ik_x \tilde{D}_z(\mathbf k,t)\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
+	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -ik_z\tilde{D}_z(\mathbf k,t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
+	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} k \tilde{D}_x(\mathbf k,t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
+	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} k \tilde{D}_z(\mathbf k,t)\exp\left( i \mathbf{k}\cdot \mathbf x\right).
 \end{align*}

### 2025-01-31 Algorithms/geometrical_parameters.tex
Add a pseudo code to details the geometrical parameters definition

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Computation of Geometrical Parameters}
\label{alg:geom_params}
\begin{algorithmic}[1]
\Require List of triangles $T_i$ with vertex positions $\mathbf{v}_i$ and normal $\mathbf{n}_i$
\Ensure Geometrical parameters computed for each triangle

\For{each triangle $T_i = (\mathbf{v}_0, \mathbf{v}_1, \mathbf{v}_2)$ with normal $\mathbf{n}_i$}
    \State Retrieve water height $d_0$, $d_1$, and $d_2$ at vertices $\mathbf{v}_0$, $\mathbf{v}_1$, and $\mathbf{v}_2$
    \State Compute area: $A_i = \frac{1}{2} \|(\mathbf{v}_1 - \mathbf{v}_0) \times (\mathbf{v}_2 - \mathbf{v}_0)\|$
    \State Compute centroid: $\mathbf{p}_i = \frac{1}{3} (\mathbf{v}_0 + \mathbf{v}_1 + \mathbf{v}_2)$
    \State Compute centroid depth: $d_i = \frac{1}{3} (d_0 + d_1 + d_2)$

    \If{$d_0 \geq 0$ and $d_1 \geq 0$ and $d_2 \geq 0$}  \Comment{Fully above water}
        \State Set submerged area $A_i^s = 0$
        \State Set volume contribution $v_{w_i} = 0$

    \ElsIf{$d_0 < 0$ and $d_1 < 0$ and $d_2 < 0$} \Comment{Fully submerged}
        \State Set submerged area $A_i^s = A_i$
        \State Compute submerged volume: $v_{w_i} = A_i d_i n_i^y$

    \ElsIf{one vertex is above the water surface (e.g., $d_0 \geq 0$)} \Comment{Partially submerged}
        \State Compute intersection points $\mathbf{p}_1, \mathbf{p}_2$ using linear interpolation:
        \State $\mathbf{p}_j = \mathbf{v}_j + \alpha (\mathbf{v}_k - \mathbf{v}_j)$ where $\alpha = \frac{d_j}{d_j - d_k}$
        \State Compute submerged area $A_i^s$ from new submerged sub-triangle(s)
        \State Compute submerged volume: $v_{w_i} = A_i^s d_i n_i^y$

    \ElsIf{two vertices are above the water surface (e.g., $d_0 \geq 0$, $d_1 \geq 0$)} \Comment{Mostly submerged}
        \State Compute intersection points $\mathbf{p}_1, \mathbf{p}_2$ using linear interpolation
        \State Compute submerged area $A_i^s$ from the remaining submerged sub-triangle
        \State Compute submerged volume: $v_{w_i} = A_i^s d_i n_i^y$
    \EndIf

    \State Store processed data ($A_i^s, v_{w_i}, \mathbf{p}_i$) for force computations
\EndFor

\end{algorithmic}
\end{algorithm}

### 2025-01-31 paper.tex
Fixed cascade length and cutoff
Add a pseudo code to details the geometrical parameters definition

@@
 \usetikzlibrary{patterns}
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\algrenewcommand{\algorithmiccomment}[1]{\hfill \textcolor{greenF}{// #1}}
 
@@
 \begin{itemize}
-	\item $L_0 = 4$ operating for waves length $k\in [0,\frac{12\pi}{16}]$
-	\item $L_1 = 16$ operating for waves length $k\in [\frac{12\pi}{16},\frac{12\pi}{256}]$
-	\item $L_2 = 256$ operating for waves length $k\in [\frac{12\pi}{256},+\infty]$.
+	\item $L_0 = 256$ operating for waves length $k\in [0,\frac{12\pi}{16}]$
+	\item $L_1 = 16$ operating for waves length $k\in [\frac{12\pi}{16},\frac{12\pi}{4}]$
+	\item $L_2 = 4$ operating for waves length $k\in [\frac{12\pi}{4},+\infty]$.
 \end{itemize}
@@
 While Bajo \etal compute the submerged volume at the texel level, the same principle can be applied using triangles instead of texels using as they did the divergence theorem. 
-Specifically, $v_w$ is approximated as the sum of the prisms with bases corresponding to the triangle areas $A_i$ and heights defined by the distance between the triangle centers $p_i$ and the water surface $h(x,z,t)$.
+Specifically, $v_w$ is approximated as the sum of the prisms with bases corresponding to the triangle areas $A_i$ and heights defined by the distance $d_i$ between the triangle centers $p_i$ and the water surface $h(p_i,t)$.
 Each term in the sum is weighted by the vertical component of the triangle normal $n_i^y$:
 \begin{equation}
-v_w = \sum_{T_i^s} (h(x,z,t) - p_i) A_i n_i^y.
+v_w = \sum_{T_i^s} A_i d_i n_i^y.
 \end{equation}
 Note that this formula is correct only under the assumption that the mesh is closed.
+
+The pseudo code for the calculations performed in this section can be read in the Algorithm \ref{alg:geom_params}.
+
+\input{Algorithms/geometrical_parameters.tex}
 

### 2025-01-31 Algorithms/water_height.tex
Add pseudo code on water height calculation

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Iterative Retrieval of Water Height}
\label{alg:compute_water_height}
\begin{algorithmic}[1]
\Require A position $\mathbf x$ to evaluate the water height and the horizontal and vertical displacements: $\mathbf{D}(\mathbf x,t)$, $h(\mathbf x,t)$
\State Initialize displacement $\mathbf{W} = (0,0,0)$
\For{$k = 1$ to $N_{\text{iter}}$} \Comment{Iterative correction to account for displacement}
    \State Compute shifted position: $\mathbf{W} = \mathbf{x} - \mathbf{W}$
    \State Compute horizontal displacement at $\mathbf{W}$ 
    \Comment{The sum of bilinear interpolation of the displacement on each cascades}
\EndFor
\State \Return The vertical component of $W$
\end{algorithmic}
\end{algorithm}

### 2025-01-31 paper.tex
Add pseudo code on water height calculation

@@
 \subsection{Water displacement and derivative}
-
+\label{subsec:water_displacement_and_derivative}
 \Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x(\mathbf x,t)\ D_z(\mathbf x,t)\right]^T$ of seawater to simulate more realistic choppy waves.
@@
 Then it does a combination with time-dependent terms at each iteration, giving an expression for the coefficients in frequency space for the eight IFFT.
+
+To compute the water height at a specific point, it is not sufficient to directly evaluate Equation \ref{equa:heightWater} after performing the IFFT.
+This is because the water surface is described parametrically, with fluid positions depending on both horizontal and vertical displacements, as detailed in Section \ref{subsec:water_displacement_and_derivative}.
+Accurately determining the water surface position at a given horizontal location requires an iterative process (see Algorithm \ref{alg:compute_water_height}): the estimated position is adjusted by subtracting the displacement vector, and the water height is recalculated until convergence.
+However, in practice, waiting for full convergence may be computationally expensive.
+Instead, the loop is limited to $N_{iter}=4$ iterations, a value chosen to balance accuracy and performance (see Appendix \ref{sec:evaluation_of_the_number_of_iterations_for_convergence} for details on this choice).
+
+\input{Algorithms/water_height.tex}
 
@@
 \end{proof}
+
+\section{Evaluation of the Number of Iterations for Convergence}
+\label{sec:evaluation_of_the_number_of_iterations_for_convergence}
+The optimal number of iterations ($N_{\text{iter}}$) required for the iterative retrieval of the water height was determined through a series of tests.
+Wind speeds were varied from $0.1 \, \text{m/s}$ to $35.0 \, \text{m/s}$ in increments of $0.5 \, \text{m/s}$, and the convergence of water height was evaluated at $1000$ random positions over a $10^4 \, \text{m} \times 10^4 \, \text{m}$ domain.
+
+The program iteratively computed the water height until the changes in height were smaller than a $1$ cm.
+
+The results, as shown in Figure~\ref{fig:convergence_steps}, demonstrate that the required number of iterations increases with wind speed, stabilizing at approximately $N_{\text{iter}} = 4$ for higher wind speeds.
+This finding suggests that using $N_{\text{iter}} = 4$ ensures reliable convergence for all tested conditions while balancing computational efficiency.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{Figures/results_cm.png}
+    \caption{Mean number of steps required for convergence as a function of wind speed. Error bars represent the variance of iterations across $1000$ test points.}
+    \label{fig:convergence_steps}
+\end{figure}
 %%%%%%%%%%%%%%%%

### 2025-01-31 paper.tex
Clarify the figure

@@
 	\caption{Summary of forces calculation. Using the same notation as Section~\ref{subsec:forces} it follows:
-		the sum of the buoyancy forces \textcolor{jauneC}{$F_b$} in \textcolor{jauneC}{yellow}, applied to the center of immersion $c_i$;
-		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to $c_i$ too;
+		the sum of the buoyancy forces \textcolor{jauneC}{$F_b$} in \textcolor{jauneC}{yellow}, applied to the center of immersion;
+		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to the center of immersion too;
 		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the unsubmerged part;

### 2025-02-03 paper.tex
Update appendix on "Evaluation of the Number of Iterations for Convergence"

@@
 \label{sec:evaluation_of_the_number_of_iterations_for_convergence}
-The optimal number of iterations ($N_{\text{iter}}$) required for the iterative retrieval of the water height was determined through a series of tests.
-Wind speeds were varied from $0.1 \, \text{m/s}$ to $35.0 \, \text{m/s}$ in increments of $0.5 \, \text{m/s}$, and the convergence of water height was evaluated at $1000$ random positions over a $10^4 \, \text{m} \times 10^4 \, \text{m}$ domain.
-
-The program iteratively computed the water height until the changes in height were smaller than a $1$ cm.
-
-The results, as shown in Figure~\ref{fig:convergence_steps}, demonstrate that the required number of iterations increases with wind speed, stabilizing at approximately $N_{\text{iter}} = 4$ for higher wind speeds.
-This finding suggests that using $N_{\text{iter}} = 4$ ensures reliable convergence for all tested conditions while balancing computational efficiency.
+
+The optimal number of iterations ($N_{\text{iter}}$) required for the iterative retrieval of water height was determined through two experiments, focusing on accuracy and computational efficiency.
+
+The first experiment aimed to evaluate the number of iterations required for the water height computation to converge within a given precision.
+This test has been conducted on different wind speeds as it makes the height of water significantly.
+Wind speeds ranged from $0.1$ m/s to $35$ m/s in $0.5$ m/s increments, with convergence tested at $10^3$ random positions uniformly distributed over a $10^4\times 10^4$ m domain.
+The iterative process was repeated until the difference between consecutive height estimates was smaller than $0.01$ m.
+The results, shown in Figure~\ref{fig:convergence_steps}, indicate that the required number of iterations increases with wind speed but stabilizes at approximately $N_{\text{iter}} = 4$ for higher wind speeds.
+
+The second experiment aimed to measure the computational cost associated with varying numbers of iterations.
+The execution time was assessed using a CUDA-based implementation, where the iterative retrieval of water height was performed for different values of $N_{\text{iter}}$.
+Each test was executed $10^3$ times on a GPU, and execution times were recorded across $10^5$ randomly sampled positions.
+The measurements included the minimum, maximum, mean, and variance of execution times for each iteration count.
+The results, presented in Figure~\ref{fig:execution_time}, show that while execution time increases with the number of iterations, the rate of increase diminishes after four iterations.
+This confirms that $N_{\text{iter}} = 4$ provides an optimal balance between computational efficiency and accuracy, ensuring reliable convergence across tested conditions.
 
@@
 \end{figure}
+
+\begin{figure}[h!] 
+	\centering 
+	\includegraphics[width=0.9\textwidth]{Figures/resultsExecutionTime.png} 
+	\caption{Average execution time per iteration count over $1000$ tests, measured across $10^5$ sample positions. Error bars indicate variance in execution time.} 
+	\label{fig:execution_time} 
+\end{figure}
+
 %%%%%%%%%%%%%%%%

### 2025-02-04 paper.tex
Add figure to clarify the desired mask

@@
 Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.
-Moreover, the boat forms a $V$ shape, with its center a bit below the extremity of the side.
+Moreover, the boat forms a $V$ shape laterally, with its center a bit below the extremity of the side.
+Figure \ref{fig:mask_maskTheoreticalView} presents the desired shape from three different perspectives.
+
+\begin{figure}[htbp]
+	\centering
+
+	\subfloat[Three quarter view]
+	{
+		\includegraphics[width=0.25\textwidth]{Figures/mask_diagonal_view.png}
+	}
+	\hspace{1cm}
+	\subfloat[Back view]
+	{
+		\includegraphics[width=0.25\textwidth]{Figures/mask_back_view.png}
+	}
+	\hspace{1cm}
+	\subfloat[Side view]
+	{
+		\includegraphics[width=0.25\textwidth]{Figures/mask_side_view.png}
+	}
+	\caption{Three different theoretical point of view of the mask for a simple rectangle.}
+	\label{fig:mask_maskTheoreticalView}
+\end{figure}
 
@@
 	\centering
-
-	\subfloat[A theoretical view of the height of the mask, with input a simple rectangle.]
-	{
-		\includegraphics[width=0.25\textwidth]{Figures/Mask3D_theoric_view.png}
-		\label{subfig:maskTheoreticalView}
-	}
-	\hspace{1cm}
 	\subfloat[The height of the mask of a simple boat encoded in a single channel texture.]
@@
 	}
-	\caption{Three different views of the height of the mask defined by Equation \ref{equa:mask_full}.}
+	\caption{Two different views of the height of the mask defined by Equation \ref{equa:mask_full} for a simple boat.}
 	\label{fig:mask}

### 2025-02-04 Algorithms/fluid_to_solid.tex
Add implementation details about the sum of forces

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Fluid-to-Solid Interaction Algorithm}
\label{alg:fluid_to_solid}
\begin{algorithmic}[1]
\For{all triangles $T_i$} \Comment{Iterate over all triangles}
\State Compute geometrical parameters \Comment{Algorithm \ref{alg:geom_params}}
\State Compute drag forces $\mathbf{F_d}$ for each triangles\Comment{Section \ref{subsubsec:dragForce}}
\EndFor
\State Apply parallel reduction to aggregate forces and geometrical parameters.
\State Transfer reduced forces and parameters to the CPU
\State Compute buoyancy force $\mathbf{F_b}$ from geometrical parameters \Comment{Section \ref{subsubsec:buoyancy}}
\State Integrate total forces and moments in the physics engine
\end{algorithmic}
\end{algorithm}

### 2025-02-04 paper.tex
Add more information about the bodies alignment
Add implementation details about the sum of forces

@@
 \label{subsec:algoForces}
-At each iteration, the fluid-to-solid algorithm proceeds as follows:
-\begin{enumerate}
-	\item Computing the vertices' position of the body relative to the water height (Section \ref{subsec:geomParam});
-	\item Computing the sets of submerged and not submerged triangles, by splitting the triangles having vertices in above and below the water surface (Section \ref{subsec:geomParam});
-	\item Computing the buoyancy force for each body (Section \ref{subsubsec:buoyancy});
-	\item Computing the two drag forces for each triangle, and summing them for each body (Section \ref{subsubsec:dragForce});
-	\item Applying the forces in the physical integration scheme.
-\end{enumerate}
-
-A summary of the forces that are applied on solid can be found on Figure \ref{fig:forcesComputation}.
+
+The fluid-to-solid algorithm integrates the forces computed on the GPU with a physics integration scheme executed on the CPU.
+To bridge this GPU-CPU workflow efficiently, a parallel reduction is applied on the GPU.
+This reduction aggregates the forces and geometrical parameters computed for individual triangles into global values, minimizing the data transferred to the CPU and reducing latency.
+
+Water and air drag forces are computed individually for each submerged or non submerged triangle and then summed using parallel reduction.
+Water drag forces are applied at the center of immersion, while air drag forces are applied at the center of the non submerged part.
+At the same time, geometric parameters such as immersion center are also aggregated for use in the integration process.
+
+Buoyancy, however, is a special case.
+Since it depends solely on the total submerged volume, it is computed once for the entire object using an analytical formula.
+The resulting buoyancy force is applied directly to the center of immersion without requiring per-triangle summation.
+
+This approach optimizes computational efficiency by reducing unnecessary calculations and data transfers, ensuring a real-time response within the physics solver.
+
+The fluid-to-solid method is presented on Algorithm \ref{alg:fluid_to_solid}.
+
+\input{Algorithms/fluid_to_solid.tex}
+
+A summary of the forces applied to the solid is shown in Figure \ref{fig:forcesComputation}.
 
@@
 	\caption{Summary of forces calculation. Using the same notation as Section~\ref{subsec:forces} it follows:
-		the sum of the buoyancy forces \textcolor{jauneC}{$F_b$} in \textcolor{jauneC}{yellow}, applied to the center of immersion;
+		the buoyancy force \textcolor{jauneC}{$F_b$} in \textcolor{jauneC}{yellow}, applied to the center of immersion;
 		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to the center of immersion too;
-		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the unsubmerged part;
+		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the non submerged part;
 		and finally the gravity force \textcolor{greenF}{$F_g$} in \textcolor{greenF}{green}, applied to the center of gravity.
@@
 $v$ is in the polygon if this number of intersections is odd; otherwise it is not.
+
+To ensure consistency regardless of the body's orientation, a rotation alignment is applied to the intersected polygons before processing.
 

### 2025-02-04 Algorithms/fdm.tex
Add more details on fdm and mask calculations

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Compute Shader for Interactive Wave Height Calculation}
\label{alg:fdm_compute}
\begin{algorithmic}[1]
\Require List of zone-specific parameters: $a$, $b$, $d(t)$, $\mathbf{p}(t-dt)$, $\mathbf{p}(t)$ and $\mathbf{p}(t-dt)$.
\For{each vertices of the FDM grid $i,j$ and zone $Z$ in the simulation domain}
    \State $k = i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{\delta}\right\rfloor$
    \State $l = j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{\delta}\right\rfloor $
    \State $o = i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{\delta}\right\rfloor$
    \State $p = j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{\delta}\right\rfloor$
    \State $h_{i,j}^{n+1} = d \cdot \big(a \cdot (h_{k-1,l} + h_{k+1,l} + h_{k,l-1} + h_{k,l+1})+ b \cdot h_{k,l} - h_{o,p}^{l-1} \big)$
\EndFor
\end{algorithmic}
\end{algorithm}

### 2025-02-04 Algorithms/mask.tex
Add more details on fdm and mask calculations

%!TEX root = ../paper.tex
\begin{algorithm}
\caption{Mask Calculation via Point-in-Polygon Test}
\label{alg:mask_calc}
\begin{algorithmic}[1]
\Require The bounding box, the intersection polygon
\For{each vertices of the FDM grid $i,j$ within the bounding box in a zone $Z$} 
    \State  $counter \gets 0$ \Comment{Initialize the counter for intersections}
    \For{each edges $(v_k,v_{k+1})$ of the intersection polygon}
        \State $p_1 = (i,j)$ 
        \State $p_2 = (i+\infty,j+\infty)$ 
        \State $p_3 = v_k$
        \State $p_4 = v_{k+1}$
        \If{Segment-Intersect($p_1$, $p_2$, $p_3$, $p_4$) do}
            \State $counter+=1$
        \EndIf
    \EndFor
    \If{$counter \bmod 2 = 1$} \Comment{Point $(i,j)$ is inside an intersection polygon.}
        \State $h_{i,j}^{n} = b_w\left(f(x)+g(z)\right)$ \Comment{See Equation \ref{equa:mask_full}}
    \EndIf
\EndFor     
\end{algorithmic}
\end{algorithm}

### 2025-02-04 paper.tex
Add more details on fdm and mask calculations

@@
 \end{figure}
+
+To ensure high performance, the FDM update step is executed entirely on the GPU,
+with a single kernel processing all points of the simulation zone in parallel.
+Each thread updates one grid point, significantly reducing computation time. The pseudo code associated to this method is shown in Algorithm \ref{alg:fdm_compute}.
+
+\input{Algorithms/fdm.tex}
 
@@
 	{
-		\includegraphics[width=0.25\textwidth]{Figures/MaskHeightMap.png}
+		\includegraphics[width=0.4\textwidth]{Figures/MaskHeightMap.png}
 		\label{subfig:maskHeightMap}
@@
 	{
-		\includegraphics[width=0.25\textwidth]{Figures/Mask3D-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/Mask3D-contrast.png}
 		\label{subfig:mask3D}
@@
 \end{figure}
+
+The mask calculation is implemented on the GPU to leverage parallel processing efficiency.
+A dedicated kernel is launched for each zone only for the vertices of the FDM grid that fall within the bounding box of the intersection polygon, reducing computational overhead.
+Each thread independently performs the point-in-polygon test using a ray-intersection method and determines whether the corresponding vertex should be included in the mask.
+Once identified as inside an intersection polygon, the height function is applied per vertex, using the predefined wave generation model.
+The full process is detailed in Algorithm \ref{alg:mask_calc}.
+
+\input{Algorithms/mask.tex}
 

### 2025-02-04 paper.tex
Add more details on fluid to solid computations

@@
 
-The fluid-to-solid algorithm integrates the forces computed on the GPU with a physics integration scheme executed on the CPU.
-To bridge this GPU-CPU workflow efficiently, a parallel reduction is applied on the GPU.
-This reduction aggregates the forces and geometrical parameters computed for individual triangles into global values, minimizing the data transferred to the CPU and reducing latency.
-
-Water and air drag forces are computed individually for each submerged or non submerged triangle and then summed using parallel reduction.
-Water drag forces are applied at the center of immersion, while air drag forces are applied at the center of the non submerged part.
-At the same time, geometric parameters such as immersion center are also aggregated for use in the integration process.
-
-Buoyancy, however, is a special case.
-Since it depends solely on the total submerged volume, it is computed once for the entire object using an analytical formula.
-The resulting buoyancy force is applied directly to the center of immersion without requiring per-triangle summation.
-
-This approach optimizes computational efficiency by reducing unnecessary calculations and data transfers, ensuring a real-time response within the physics solver.
-
-The fluid-to-solid method is presented on Algorithm \ref{alg:fluid_to_solid}.
+To ensure a structured and efficient data workflow on the GPU, all computations for a given mesh are performed separately, avoiding data interleaving between different meshes.
+Each kernel invocation processes a single mesh, iterating over its triangles before moving to the next mesh.
+
+Firstly, for each triangle in the mesh, the geometric parameters are calculated and then used to compute the drag forces.
+These include necessary attributes such as the submerged volume, normal vectors, and the drag forces acting on submerged and non submerged parts.
+
+Once per-triangle computations are completed, the forces and geometrical parameters must be integrated into the physics solver, which operates on the CPU.
+To minimize data transfers and reduce latency, a parallel reduction kernel is applied on the GPU for each mesh individually.
+This step aggregates all per-triangle forces and parameters into global values for the entire mesh.
+
+Buoyancy, however, is handled differently.
+Since it depends solely on the total submerged volume rather than individual triangles, it is computed separately using an analytical formula and applied directly at the submerged center without requiring per-triangle summation.
+
+With all global parameters computed, the final step transfers the data to the CPU, where the physics solver integrates the buoyancy and the water drag forces are applied on the submerged center, while air drag forces are applied at the center of the non submerged part.
+
+This pipeline is presented in Algorithm \ref{alg:fluid_to_solid}.
 

### 2025-02-04 paper.tex
Clarify implementation details

@@
 
-Second, as \framework simulates a tile of water of size $N\times N$ using $c$ cascades, it stores the height, the displacement and so on in an 1D array of size $c$ containing a vector of 2D arrays of size $N\times N$.
+Second, as \framework simulates a tile of water of size $N\times N$ using $c$ cascades, it stores the height, the displacement, and other relevant data in a 1D array of size $c$, where each entry contains a vector of 2D arrays of size $N\times N$.
+\Framework compute efficiently these values on the GPU, with dedicated kernel that are launched with a grid of threads structured as $N \times N \times c$.
+Specifically, $N \times N$ threads are assigned to the $x$ and $y$ dimensions to cover the spatial resolution of the water tile, while $c$ threads along the $z$ dimension correspond to the different cascades.
+This approach ensures that computations for all cascades are executed in parallel.
 

### 2025-02-04 paper.tex
Clarify why CPU isn't used too much

@@
 \tablename~\ref{fig:performances} shows a summary of GPU computation times for each method and with these two experiments. 
-CPU computation times are not considered, as they are negligible compared to GPU ones. 
+As detailed in previous sections, all heavy computations in \framework are offloaded to the GPU, while the CPU is minimally involved in physical calculations.
+This design choice ensures that the CPU remains available for other standard game or simulation tasks, optimizing overall system performance.
+Therefore, CPU computation times are not considered, as they are negligible compared to GPU ones.
 First, it can be observed that real-time performance is largely achieved, even with \emph{ten-solids}. 

### 2025-02-05 paper.tex
Made some small typos enhancement

@@
 \end{equation}
-where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
+where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the $i$-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
 \begin{equation}
@@
 \end{equation}
-where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.
+where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphs \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.
 
@@
 
-The pseudo code for the calculations performed in this section can be read in the Algorithm \ref{alg:geom_params}.
+The pseudocode for the all geometrical calculations performed in this section can be read in the Algorithm \ref{alg:geom_params}.
 
@@
 Firstly, for each triangle in the mesh, the geometric parameters are calculated and then used to compute the drag forces.
-These include necessary attributes such as the submerged volume, normal vectors, and the drag forces acting on submerged and non submerged parts.
+These include necessary attributes such as the submerged volume, normal vectors, and the drag forces acting on submerged and non-submerged parts.
 
@@
 
-With all global parameters computed, the final step transfers the data to the CPU, where the physics solver integrates the buoyancy and the water drag forces are applied on the submerged center, while air drag forces are applied at the center of the non submerged part.
+With all global parameters computed, the final step transfers the data to the CPU, where the physics solver integrates the buoyancy and the water drag forces are applied on the submerged center, while air drag forces are applied at the center of the non-submerged part.
 
@@
 		the sum of water drag force \textcolor{purpleF}{$F_w$} in \textcolor{purpleF}{purple}, applied to the center of immersion too;
-		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the non submerged part;
+		the sum of the air drag forces \textcolor{redF}{$F_a$} in \textcolor{redF}{red}, applied to the center of the non-submerged part;
 		and finally the gravity force \textcolor{greenF}{$F_g$} in \textcolor{greenF}{green}, applied to the center of gravity.
@@
 
-To ensure high performance, the FDM update step is executed entirely on the GPU,
-with a single kernel processing all points of the simulation zone in parallel.
-Each thread updates one grid point, significantly reducing computation time. The pseudo code associated to this method is shown in Algorithm \ref{alg:fdm_compute}.
+To ensure high performance, the FDM update step is executed entirely on the GPU, with a single kernel processing all vertices of the FDM grid in parallel.
+Each thread updates one grid point, significantly reducing computation time. The pseudocode associated to this method is shown in Algorithm \ref{alg:fdm_compute}.
 
@@
 The FDM needs the height of the vertices below the body.
-This is computed from the mask, composed of the vertices of $Z$ contained in one of the intersection polygons at least.
+This is computed from the mask, composed of the vertices of $Z$ contained in one of the intersections polygons at least.
 This is done using the \textbf{Point in Polygon} strategy described in Haines (1994) \cite{hainesPointPolygonStrategies1994}.
@@
 It is necessary to define a boundary condition at the interface between the ocean fluid and the air fluid.
-To simplify, it is assume that the air can be represented as a region with constant pressure\footnote{As shown by the \href{https://en.wikipedia.org/wiki/Barometric_formula}{barometric formula} the atmospheric pressure has a very low evolution in function of altitude.}.
+To simplify, it is assumed that the air can be represented as a region with constant pressure\footnote{As shown by the \href{https://en.wikipedia.org/wiki/Barometric_formula}{barometric formula} the atmospheric pressure has a very low evolution in function of altitude.}.
 Since air density is a hundred of times lighter than water, its effects on the free surface are assumed to be negligible.

### 2025-02-05 review/review.tex
Add a review in tex

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

% Define the custom color
\definecolor{answercolor}{rgb}{0.35,0.47,0.28}

% Define the macro
\newcommand{\answer}[1]{\textbf{\textcolor{answercolor}{Answer:}} \textcolor{answercolor}{#1}}

\title{Response to Reviewers' Comments}
\date{\today}

\begin{document}

\maketitle

This document provides responses to the reviewers' comments. First of all, we would like to sincerely thank the reviewers for their thoughtful and constructive feedback. We feel that their input has enhanced the clarity and quality of this manuscript a lot, and we appreciate their efforts to improve our article.

Below, we reproduce each reviewer's comments followed by our responses in green.

\section*{Review 1}

\begin{enumerate}[label=\textbf{\arabic*.}]
    \item \textbf{The normalization factor Q is missing in Equation 13.}

    \answer{The normalization factor is missing on purpose. Indeed, the normalization factor needs to be applied only to the full spectrum in Equation 15. Moreover, we applied it in Equation 8 because, as mentioned under Equation 17, "normalization problems will occur if the full-spectrum normalization factor is simply approximated before its application. For small values of $r_{\omega}$, $Q_{DB\xi}(\omega)$ will be very low and therefore numerically challenging to normalize."}

    \item \textbf{Figure 3 caption mentions the dispersion parameter, but the figure is actually about another parameter $\delta$.}

    \answer{Thanks, we have corrected and clarified the caption of Figure 3.}

    \item \textbf{I am confused about the parameters of the listed cascades in Section 2.3. First of all, the ranges listed seem wrong: $12/16$ is clearly greater than $12/256$. I presume this is a typo. Also, shouldn't $L$ be larger for smaller $k$?}

    \answer{Indeed, we confused the length of the cascades with their cutoff. We have corrected this mistake.}

    \item \textbf{Figure 9 caption says that $F_a$ is "applied to the center of the submerged part." It should be the "unsubmerged part" instead.}

    \answer{We have corrected this typo.}

    \item \textbf{It is unclear what Figure 11(c) shows.}

    \answer{We have clarified the captions of all subfigures in Figure 11 to improve clarity.}

    \item \textbf{$h^*$ in Equations 48, 50, and 59 should be negative (following Equation 46). The same goes for unlabeled equations above Equation 53.}

    \answer{We have corrected this sign error.}

    \item \textbf{In the equation above Equation 56, "sinh(kY)" should be "sinh(ky)."}

    \answer{We have corrected this case mistake.}

    \item \textbf{The parameters $y$, $t$, and $k$ in Equations 59, 60, 61, and 62 are written in an inconsistent order. They should all use the same order.}

    \answer{We have standardized the order of all parameters throughout the manuscript for consistency.}

    \item \textbf{This sentence in Appendix A is unclear and needs more explanation: "Note that while the third equation of system 1 is not used here, it can help to deduce the dispersion relation."}

    \answer{We have removed this sentence and replaced it with a full derivation of the dispersion relation for better clarity.}

    \item \textbf{In Equation 71, $v_y$ should be $v_z$.}

    \answer{We have corrected this typo.}

    \item \textbf{In Equation 73, there appears to be an extra term: $h(x,t)/dt$.}

    \answer{We have removed the excess term.}

    \item \textbf{The abstract ends with the term "fluid-to-solid algorithm." I would replace it with "fluid-to-solid coupling algorithm."}

    \answer{We have added "coupling" to make the term more coherent with the literature.}

    \item \textbf{The introduction begins by mentioning "scientific computing," but the paper is not about scientific computing at all. I would replace it with "computer graphics."}

    \answer{We replaced "scientific computing" with "computer graphics," which is more appropriate for our article and JCGT.}

    \item \textbf{I would recommend removing the subsection title 1.1 and converting Section 1.2 (related work) into Section 2.}

    \answer{Indeed, placing related work in a dedicated section is more coherent with the literature.}

    \item \textbf{On page 5, "standard gravity" should be "gravitational acceleration."}

    \answer{We changed "standard gravity" to "gravitational acceleration."}

    \item \textbf{In Section 3.2.3, the description of $\rho_a$ is repeated.}

    \answer{We have removed the duplicate description.}

    \item \textbf{In Section 4.4.2, "M has is bow" should be "M has a bow."}

    \answer{We fixed the typo.}

    \item \textbf{On page 27, "fluid-to-solid coupling FDM" should be "solid-to-fluid coupling FDM."}

    \answer{We fixed the typo.}
\end{enumerate}

\section*{Review 2}

\begin{enumerate}[label=\textbf{\arabic*.}]
    \item \textbf{The bibliography is OK, though I am a bit confused as to why Yuksel's PhD thesis is referenced instead of the related paper.}

    \answer{We used this reference instead of the thesis when discussing wave particles.}

    \item \textbf{I enjoyed reading this paper. I would have liked it a lot more if it included more detailed descriptions of the algorithms and/or source code.}

    \answer{A large part of the correction concerns adding details about implementation like pseudocode to simplify the implementation for readers.}

    \item \textbf{Give more details on the actual implementation: memory layout on the GPU, implementation of the kernels, etc. This is especially relevant since performance timing is given in Sec 5 of the paper, but the rest of the paper gives little insight into why these numbers are as they are. For example, the bottleneck in Fig 14 seems to be geometry and force computation (intersection geometry computation). Why should this be so slow?}

    \answer{TODO.}

    \item \textbf{The paper should provide more implementation details, specifically regarding GPU computation and data layout, to explain the timing numbers in Fig 14 and why some of them seem heavy (e.g., geometry processing). The paper also mentions that no serious CPU-side computation is used. It would be interesting to know why (e.g., could the geometry processing (intersection curve), which seems to be a bottleneck according to Fig 14, be offloaded to the CPU?).}

    \answer{The decision to keep all heavy computations on the GPU, including geometry processing, is motivated by minimizing data transfers between CPU and GPU, which can introduce significant latency. While offloading intersection curve computations to the CPU could theoretically balance the workload, it would require frequent memory exchanges, likely negating any performance gain. In addition, we added a comment in Section 6 to reinforce the fact that this choice of "no serious CPU-side computation" was made to ensure that the CPU remains available for other standard tasks.}

    \item \textbf{First bullet point, (TeX error) here and elsewhere Tessendorf is quoted as "Tessendorf (2004) [Tessendorf 1999]" (i.e., reference link and text do not match).}

    \answer{We changed references to make them coherent between text and citation.}

    \item \textbf{Sections 2.1 - 2.5 are very heavy on hydrodynamics.}

    \answer{Thank you for the insightful comment. We agree that the derivations are detailed; however, we believe these sections help simplify the reader's understanding of the underlying hydrodynamic principles and provide clarity for implementation. Moreover, it makes the paper a bit more self-contained.}

    \item \textbf{Isn't $k$ locally constant (wrt $x,z$)? Then, could just write:}
    \[
    \frac{\partial D_x}{\partial x} = i k_x D_x
    \]
    \textbf{for the first equation, and similar for the other 4 derivatives. This makes it easier to read and is explicit about computation needed.}

    \answer{$k$ is indeed locally constant relative to $x$ and $z$; however, the equation $\frac{\partial D_x}{\partial x} = i k_x D_x$ isn't strictly true, because $k_x$ cannot be factored out from the sum, as it is the index over which the sum is applied. Nonetheless, this suggestion holds when considering each term of the sum individually. We have rewritten this paragraph to present these equalities in a more "Fourier-fashion," aligning better with the implementation.}

    \item \textbf{Second paragraph p 12. It would be helpful to have some details on memory organization for the NxN tiles, computation efficiency on the GPU, etc - especially since quoting performance numbers in Sec 5, Figure 14.}

    \answer{Thank you for the valuable suggestion; however, we believe the current level of detail sufficiently balances technical depth and readability}

    \item \textbf{Paragraph 1, p 13, awkward phrasing "...tied to the wave vector, preventing from any simplifications to reduce the needed calculations."}

    \answer{We rephrased this sentence with your suggestions.}

    \item \textbf{Para 2, p 14 "First, while the velocity decays exponentially, it relies on a logarithmic similar to..." should probably be "..., it relies on a logarithmic function similar to..."}

    \answer{We added the missing word.}

    \item \textbf{Same paragraph, "this function decreases slowly (actually with ln(250 + 1) = 5.5),". What is the significance of "250 + 1" here?!}

    \answer{The "250 + 1" was to keep the same "form" as the function, but we removed it to avoid confusion.}

    \item \textbf{Figure 6, the vertical scale should probably be logarithmic to make these graphs more informative. As it is now, three of four graphs are scrunched on top of each other.}

    \answer{TODO.}

    \item \textbf{Not sure why the bracket "\{" in eq. (28). Also it does not look like eq (28) is quoted in the rest of the paper, so $\alpha$ and $\beta$ could just be given inline.}

    \answer{We inline the coefficients choice.}

    \item \textbf{"The Arc Blanc framework uses the prism approximation proposed by Bajo et al." How? In Bajo et al.’s paper, the geometry (position \& area vector) is stored in texels of the surface textures of the model. The prisms used there are per texel, so for example, the prism height is determined by (average) water height above the projected texel center. How is this done here?}

    \answer{Thank you for your insightful comment. We revised the paragraph to clarify and justify that we use an approximation similar to Bajo et al., but at the triangle level rather than the texel granularity. We added some formal details about the calculation.}

    \item \textbf{Sections 3.2.2 and 3.2.3 details are redundant: the only true geometric difference between the two is that in one case we are considering velocity of water while in the other velocity of air (density and drag coefficient being that of the medium under consideration can be understood from context)}

    \answer{We have merge the two sections to remove redundancy.}

    \item \textbf{Also, p.19 Paragraph 2 has a typo: the 3rd and 4th bullets are identical.}

    \answer{We have removed the doublon.}

    \item \textbf{As above, item 3, it is not clear how to do the volume computation. The paper claims it is triangle based, while the quoted source (Bajo et al) uses surface texel based computation. Does the author imply that some sort of uniform triangulation scheme is to be used? If so, then there is also the matter of closing and triangulating the polygons given by the (approximate) cutting of triangles against the waterline described in paragraph 2 of 3.1 on p 17 by referring to Kerner's paper.}

    \answer{As mentioned above, we have made the paragraph on the volume computation more clear. Moreover, we clarified the hypothesis that we have made on the meshes.}

    \item \textbf{Figure 9, p 20 mentioning }$\mathbf{c_i}$\textbf{, COG, center of submerged part etc, but these are not labeled in the diagram, which produces unnecessary cognitive load when reading the diagram. Also, can spread the vectors a bit to declutter (dont need to have them bunched up so realistically), likewise }$\mathbf{v_a}$\textbf{ and }$\mathbf{v_w}$\textbf{ are not collinear in general. Larger arrow tips would help.}

    \answer{We tried to clarify the caption while keeping information about where to apply the forces. Moreover we change the figure based on your comment to make it more clear.}

    \item \textbf{ Paragraph 1, suggesting change of notation, e.g. $\delta$ instead of $dx$ for the discretization step. This would make (35) (38) (40) (42) less confusing when first seen (e.g. it's not a differential), or $\Delta{x}, \Delta{y}, \Delta{t}$, but that might be a bit too detailed, e.g. since the grid is regular.}

    \answer{As suggested, we renamed $dx$ in $\delta$ to make equation less confuse.}

    \item \textbf{Eq (36)  probably it is easier to read something like $u = clamp_hi(|v|/v_max, 1)$ and $d(t) = lerp(d0, d_max, u)$ than the given piecewise definition.}

    \answer{Indeed this formulation is more clear. We replace our piecewise definition with this one.}

    \item \textbf{Numeric values like the ones in eq 37, which affect stability, might be meaningless if some implementation details of the framework are not been specified (e.g. floats, doubles etc)?}

    \answer{We have specified that we are using single precision float number.}

    \item \textbf{A simple diagram showing the relationship between (i,j), (k,l), and (o,p) would make (39) easier to read (e.g. could extend Fig 10 to show this).}

    \answer{We have added a figure to illustrate the relationship between indices.}

    \item \textbf{"The value of c is deduced from dx and dt to validate the CFL as follows"Deduced how? }$\mathbf{\sqrt(0.49)}$\textbf{ seems close to }$\mathbf{\sqrt(0.5)}$\textbf{ which you would get from Eq (40)?}

    \answer{We have clarified this sentence to explain that this choice has been made to make sure the equation 40 is true.}

    \item \textbf{Last paragraph on p 23, typo: "corresponding to the vertices of Z contained into one intersected polygons." should be "...in one of the intersection polygons"}

    \answer{We have fixed this typo.}

    \item \textbf{Paragraph 1, typo: "contained into one intersection polygon" should be "contained in one of the intersection polygons"}

    \answer{We have fixed this typo.}

    \item \textbf{The reference to (what I assume is) CLRS listed in the References section is incorrectly written. Also CLRS is too general so probably not needed here.}

    \answer{We removed this ref which makes doublon with the more precise reference of E. Haines (1994).}

    \item \textbf{Stabbing ray approach described might have robustness issues - are those significant enough to cause some form of instability in the simulation?}

    \answer{TODO what robustness issues and instability ?}

    \item \textbf{Last sentence of the last paragraph of 4.4.1, typo: "technic"}

    \answer{We have fixed this typo.}

    \item \textbf{Also, not explicitly mentioned, but I am guessing that the intersection polygons from sec 3.1 are projected onto xz plane in order to do z-ray stabbing in sec 4.4.1. Could some such projected intersection polygons then be generate (e.g. self intersecting when projected onto zx plane, or inverted - e.g. for turbulent water height and finely tessellated body M)}

    \answer{}

    \item \textbf{Also, doing ray stabbing on multiple components probably increases chances of bad classification of a point. Could do closed component count (provided the model M is a closed orientable surface to start with, so the intersection polygons have well-defined interior) to speed up the test.}

    \answer{We didn't face any bad classification at this point, but we might in the future considering using closed component count indeed, thanks for the suggestion.}

    \item \textbf{Sentence 2, paragraph 1, typo: "By construction of the grid Z, M has is bow oriented on the positive side of the z axis and its starboard on the positive side of the x axis." Should be "..has *its* bow.."?}

    \answer{We have fixed this typo.}

    \item \textbf{Also, for landlubbers among us, please clarify "starboard" as "right-hand side when facing the bow" :)}

    \answer{We added the precision to make the term more clear.}

    \item \textbf{End of the same paragraph:"Hence, h is made such that from the vertex at position x, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of M. Moreover, the boat forms a V shape, with its center a bit below the extremity of the side."
I can not visualize this. For example, 
- what is meant here by "front of the mask is above the free surface and the back below"?
- "..the boat forms a V shape.." - does V here mean that the boat is in the shape of the letter "V", or does V denote the shape (object) in question? For example, the sentence before eq (43) says "The function f models the side effect using a shape in V", which suggest the latter, but the next sentence says "The V-shape is obtained from the absolute value of the abscissa" (note the hyphen) which suggest the former.
Can we please have a diagram that shows and explains this geometry configuration clearly - in addition to Fig 11?}

    \answer{We have added some explanation and three figures to make clarify the desired shape.}

    \item \textbf{Figure 14, XP2 geometry, section 3.1. 4.6 msec for 4153 triangles vertex classification and splitting seems a bit excessive. The same goes for forces computation. This, for example, is why it would be nice to see more implementation details.}

    \answer{TODO.}

    \item \textbf{Also, an idea: Maybe assuming that the intersection shape is more or less the same between two consecutive frames, for low Beaufort state at least (and for some $k$ steps, where $k$ is small), then recompute the buoyancy/volume based on the previous volume of the submerged shape, correcting for the height difference of the water between the two frames, then on the $k$-th step doing the correct computation?}

    \answer{Thanks for your suggestion. The project for which this article was made requires precise hydrodynamics data at each frame. Nonetheless, your approximation could be interesting for simulations with a low Beaufort state. However, it might complicate the GPU implementation for more generalized systems.}

%     \item \textbf{Although, the solid->water action simulation is primitive and lacks explanations and obvious improvements: 
% 1. The movement of the hulls and how the corresponding translations/rotations of hulls affect the solid->water interaction is not clearly described.
% It appears that only the translation part is taken in account.}

%     \answer{We have added some precision about how to handle the rotation of the body for mask calculation.
% }
\end{enumerate}

\section*{Review 3}

\subsection*{General Remarks}

\begin{enumerate}[label=\textbf{\arabic*.}]
    \item \textbf{Although, the solid-to-water action simulation is primitive and lacks explanations and obvious improvements:}

    \begin{enumerate}[label=\textbf{\arabic*.}]
        \item \textbf{The movement of the hulls and how the corresponding translations/rotations of hulls affect the solid-to-water interaction is not clearly described. It appears that only the translation part is taken into account.}

        \answer{We have added more details about how the rotation of the body is handled in the mask calculation.}

        \item \textbf{The finite difference method used to simulate interactive waves is extremely primitive and is able to simulate vertical displacement of interactive waves only. Tessendorf's eWave, based on FFT + IFFT per simulation step, is a bit more expensive but produces full 3D displacements of the interactive waves.}

        \answer{We appreciate your insightful comment. While we agree that the finite difference method is more primitive and limited to simulating vertical displacements, its simplicity allows for straightforward implementation and efficient parallelization across multiple bodies in the sea. Implementing eWave would indeed be a valuable enhancement to Arc Blanc, and we might consider it for future work.}

        \item \textbf{Injecting displacements into water caused by movement of solid objects, called wave generation in the paper, based on functions representing side and front/back effects, looks absolutely artificial. Instead, a per-gridpoint signed difference in volumes produced by intersections of hull volume and still water volume at current vs previous simulation step could be used to generate water displacements based on incompressibility of water.}

        \answer{Thank you for the thoughtful suggestion. Using per-gridpoint signed differences based on the incompressibility of water is indeed a more realistic approach and could significantly enhance the accuracy of wave generation. However, we chose our current approximation to prioritize simplicity and straightforward implementation, which aligns better with the real-time constraints and accessibility goals of Arc Blanc. We will certainly consider your suggestion for future enhancements.}

        \item \textbf{Mask movement and orientation used in the paper, so that the front of the mask is above the water surface and the back is below the water surface, looks artificial and extremely simplified. A propulsion force acting on the hull and causing it to move, in conjunction with proper water-to-solid interaction from the first part of the paper, with its forces and moments, will naturally cause the hull to raise its bow based on a physically correct simulation instead of the oversimplified approximation used in the paper.}

        \answer{Thank you for your valuable proposition. It might be a nice refinement for the Arc Blanc framework.}
    \end{enumerate}

    \item \textbf{Overall, the paper allows for implementation of the presented technique without much trouble for people who have experience with the topic of the paper, although a better explanation of how the forces acting on a solid object in water are summed up and used to evolve solid object dynamics would improve it a lot. Is the sum of the forces applied to the center of gravity of the object? Or are there rotational moments? Are moments of inertia of the solid object taken into account? What about the moment of inertia of the water body surrounding the object? Without those explained, this part of the abstract looks like a bit of an overstatement.}

    \answer{We have added details about how forces are summed and applied in a physics solver, explicitly addressing the role of rotational moments and inertia.}

    \item \textbf{One note: deciseconds? Please, use seconds or milliseconds.}

    \answer{We used seconds, which are more standard.}

    \item \textbf{The paper lists a ton of references that are not touched upon in the paper in the sense that the paper does not derive anything from those. For instance:}

    \begin{enumerate}[label=\textbf{\arabic*.}]
        \item \textbf{SPH papers are not related.}
        \item \textbf{Wave particles are not related.}
        \item \textbf{Water wave packets are not related.}
    \end{enumerate}

    \answer{As one of the goals of the paper is to describe a complete unified framework for ocean simulation in real-time, we think that presenting alternative approaches can only benefit the reader. Additionally, it makes the article more self-contained on the subject of ocean simulation.}

    \item \textbf{In Chapter 2.6, the "interpolation degree" term is unclear and a bit misleading, and since the interpolation is based on water body displacements calculated at "slices" with fixed depth, I'd suggest using the terms "interpolation slice" and "number of interpolation slices," for instance in Figure 7.}

    \answer{Thank you for your suggestion. While we understand that the term "interpolation degree" may be misleading, we prefer to retain it as it aligns with common terminology used in interpolation methods, such as polynomial interpolation.}
\end{enumerate}

\end{document}

### 2025-02-05 review/review.tex
Add proof for intersection

@@
 \usepackage{xcolor}
-
+\newtheorem{theorem}{Theorem}
 % Define the custom color
@@
 
-    \item \textbf{Numeric values like the ones in eq 37, which affect stability, might be meaningless if some implementation details of the framework are not been specified (e.g.
-floats, doubles etc)?}
+	\item \textbf{Numeric values like the ones in eq 37, which affect stability, might be meaningless if some implementation details of the framework are not been specified (e.g. floats, doubles etc)?}
 
@@
 
-    \answer{}
+	      \answer{Thanks for your remarks that has given us plenty to think about.
+We have answered to this remarks in a dedicated Section \ref{sec:evaluation_of_the_number_of_iterations_for_convergence} at the end of this document.}
 
@@
 
+\appendix
+
+\section{Sketch of Proof Mesh-Height Map Intersection}
+\label{sec:sketch_of_proof_mesh_height_map_intersection}
+
+The question on the "stability" of the intersection between $M$ and the water height can be divided into two questions:
+
+\subsection{Dimension of the Intersection}
+
+First, does the intersection between $M$ and the water height (the height map) always result in some loop (or empty/singleton set), but not in some open loop or a stranger subset of $\mathbb{R}^3$?
+This question is non-trivial and can be summarized with the following theorem:
+
+\begin{theorem}
+	Let $M$ be a compact 2-manifold without any border (our mesh) and
+	\begin{equation}
+		H=\{(x,y,z) \in \mathbb{R}^3 \mid h(x,z)-y=0\},
+	\end{equation}
+	then $M \cap H$ can only be:
+	\begin{enumerate}
+		\item Empty,
+		\item $x \in \mathbb{R}^3$,
+		\item $\cup_i c_i$ where $c_i$ are homeomorphic sets to the unit circle $S^1$.
+	\end{enumerate}
+\end{theorem}
+
+The proof of this theorem is a bit long and out of scope for this paper, as it requires some skills in differential topology, but here is a sketch of a demonstration:
+
+The idea of the proof is to characterize the dimension of the intersection and to prove that the intersection is always compact.
+Since compact sets of dimension one in $\mathbb{R}^3$ are homeomorphic to a finite union of circles, this will establish the claim.
+
+First, note that the compactness of the intersection is quite trivial, as by definition $M$ is compact and $H$ is by definition closed, therefore their intersection must be compact.
+
+Second, determining the dimension is a bit more complex.
+However, we can take advantage of the \textit{\href{https://en.wikipedia.org/wiki/Preimage_theorem}{preimage theorem}}, which under certain conditions provides a characterization of the dimension of a set.
+Consequently, our objective is to pose the problem in a way that allows us to apply the preimage theorem.
+
+Define the function:
+\begin{equation}
+	f: M \to \mathbb{R}, \quad f(x,z,y) \mapsto h(x,z)-y.
+\end{equation}
+Note that $S \cap H$ is equal to $f^{-1}(0)$. Moreover, by the definition of $h$ and its linearity, $f$ is a smooth map.
+
+Now, we need to prove that $0$ is a regular value of $f$, that is, $\forall x \in f^{-1}(0)$, the differential
+\begin{equation}
+	df_x : T_x M \to T_0 \mathbb{R} \simeq \mathbb{R}
+\end{equation}
+is surjective. To rule out the trivial case, let us directly assume that $f^{-1}(0) \neq \emptyset$. Moreover, we have:
+\begin{equation}
+	df_x(v) = \nabla f \cdot v.
+\end{equation}
+Note that if $df_x$ is not the zero map, then $df_x$ is surjective (indeed, if it is not a zero map, then $\text{im}(df_x) = \mathbb{R}$, because it cannot be the trivial vector subspace $\{0\}$).
+Therefore, we have two cases:
+
+\begin{enumerate}
+	\item If $T_x M \subset T_x H$, then $df_x$ is not the zero map, which corresponds to the intersection being a single point.
+	\item If $T_x M$ is not a subset of $T_x H$, then $df_x$ is not the zero map and thus surjective.
+\end{enumerate}
+
+By applying the preimage theorem, we obtain:
+\begin{equation}
+	\text{codim}(f^{-1}(0)) = \dim(\mathbb{R}) = 1.
+\end{equation}
+Moreover, since we are in finite dimensions:
+\begin{equation}
+	\text{codim}(f^{-1}(0)) = \dim(M) - \dim(f^{-1}(0)).
+\end{equation}
+This concludes the demonstration by giving us:
+\begin{equation}
+	\dim(S \cap H) = 1.
+\end{equation}
+
+\subsection{Intersection Degenerate}
+
+Second, as you mentioned, does the projection on the $zx$ plane cause the intersection polygons to be self-intersected?
+To address this, let us simply remark that if the polygon is simple on the continuous height map, then the projection, which only alters the vertical component, should also yield a simple polygon.
+This would not be the case if instead of a height map we were using more complex parametric surfaces, where two points of the surface could be projected onto the same point of the $xz$ plane.
+The key point here is the bijection between the $xz$ plane and the height map.
+
 \end{document}

### 2025-02-05 review/review.tex
Fixed review ref and typos

@@
 This document provides responses to the reviewers' comments.
-First of all, we would like to sincerely thank the reviewers for their thoughtful and constructive feedback.
+First, we would like to sincerely thank the reviewers for their thoughtful and constructive feedback.
 We feel that their input has enhanced the clarity and quality of this manuscript a lot, and we appreciate their efforts to improve our article.
@@
 
-	      \answer{Thanks for your remarks that has given us plenty to think about.
-We have answered to this remarks in a dedicated Section \ref{sec:evaluation_of_the_number_of_iterations_for_convergence} at the end of this document.}
+	      \answer{Thanks for your remarks that has given us plenty to think about. We have answered to this remarks in a dedicated Appendix \ref{sec:sketch_of_proof} at the end of this document.}
 
@@
 
-	      \answer{TODO.}
+	      \answer{As mention before, we have added more details about the implementation to clarify this part.}
 
@@
 \section{Sketch of Proof Mesh-Height Map Intersection}
-\label{sec:sketch_of_proof_mesh_height_map_intersection}
+\label{sec:sketch_of_proof}
 

### 2025-02-06 paper.tex
Fixed typos

@@
 
-Second, as \framework simulates a tile of water of size $N\times N$ using $c$ cascades, it stores the height, the displacement, and other relevant data in a 1D array of size $c$, where each entry contains a vector of 2D arrays of size $N\times N$.
-\Framework compute efficiently these values on the GPU, with dedicated kernel that are launched with a grid of threads structured as $N \times N \times c$.
+Second, as \framework simulates a tile of water of size $N\times N$ using $c$ cascades, it stores the height, the displacement, and other relevant data in a 1D array of size $c$, where each entry contains a 2D array of size $N\times N$.
+\Framework computes efficiently these values on the GPU, with dedicated kernel that are launched with a grid of threads structured as $N \times N \times c$.
 Specifically, $N \times N$ threads are assigned to the $x$ and $y$ dimensions to cover the spatial resolution of the water tile, while $c$ threads along the $z$ dimension correspond to the different cascades.
@@
 
-\subsubsection{Drag Force}
+\subsubsection{Drag Forces}
 \label{subsubsec:dragForce}

### 2025-02-06 paper.tex
Update the graph with log scale

@@
 	\centering
-	\subfloat{\includegraphics[width=0.5\textwidth]{Figures/interpolationAccuracyComparison.png}\label{subfig:interpolationAccuracyComparison}}
-	\hfill
-	\subfloat{\includegraphics[width=0.5\textwidth]{Figures/interpolationAccuracyComparisonZoom.png}\label{subfig:interpolationAccuracyComparisonZoom}}
-	\caption{The left figure represents the accuracy of different type of interpolation in function of the size of the discretization $N$.
-The right one does a zoom on small velocities between $0$ and $0.1$.
-		They use 10000 positions from a continuous uniform distribution in the parallelepiped $[-1000,1000]\times[-125,4.5]\times[-1000, 1000]$ and with a wind speed of 20 $m.s^{-1}$.
+	\includegraphics[width=0.8\textwidth]{Figures/interpolationAccuracyComparisonLogScale.png}\label{subfig:interpolationAccuracyComparisonLogScale}
+	\caption{This figure represents the accuracy of different type of interpolation in function of the size of the discretization $N$. The accuracy axis used a logarithmic scale to be more readable.
+		It uses 10000 positions from a continuous uniform distribution in the parallelepiped $[-1000,1000]\times[-125,4.5]\times[-1000, 1000]$ and with a wind speed of 20 $m.s^{-1}$.
 		The interpolated velocity $\tilde{\mathbf{v}}$ and the analytical velocity $\mathbf{v}$ are evaluated.
@@
 		Logarithmic discretization is much better than uniform discretization, whatever the used interpolation.
-		At least, exponential interpolation gives slightly better result than linear interpolation, even for low $N$.}
+		At least, for low $N$, exponential interpolation gives slightly better result than linear interpolation.}
 	\label{fig:interpolationAccuracyComparison}

### 2025-02-06 paper.tex
Apply latex indent for a cleaner doc
Add some information about geometrical computation

@@
 First, it can be observed that real-time performance is largely achieved, even with \emph{ten-solids}.
+One of the most computationally intensive steps is geometry computation, which presents an opportunity for further optimization.
+A closer inspection suggests that refining memory access patterns and minimizing warp branch divergence in the GPU kernel could enhance performance and efficiency.
 Besides, as most of the computations are performed on the GPU, it can be observed that computation times do not grow linearly with the number of solids.
 Indeed, \framework processes each solid and each FDM grid simulation in parallel.
-
 \begin{figure}[!h]
@@
 		\hline
-		Tessendorf's method & Height (\S~\ref{subsec:waterHeight}) & 0.376 & 0.376\\
+		Tessendorf's method               & Height (\S~\ref{subsec:waterHeight})         & 0.376              & 0.376               \\
 		\cline{2-4}
-		(\S~\ref{sec:oceanWaves}) & Velocity (\S~\ref{subsec:velWater}) & 1.175 & 1.175\\
+		(\S~\ref{sec:oceanWaves})         & Velocity (\S~\ref{subsec:velWater})          & 1.175              & 1.175               \\
 		\cline{2-4}
-		  & Total & 1.551 & 1.551\\
+		                                  & Total                                        & 1.551              & 1.551               \\
 		\hline
-		Fluid-to-solid coupling & Geometry (\S~\ref{subsec:geomParam}) & 1.125 & 4.588\\
+		Fluid-to-solid coupling           & Geometry (\S~\ref{subsec:geomParam})         & 1.125              & 4.588               \\
 		\cline{2-4}
-        (\S~\ref{sec:fluidToSolidAction}) & Forces (\S~\ref{subsec:forces}) & 0.406 & 3.677\\
+		(\S~\ref{sec:fluidToSolidAction}) & Forces (\S~\ref{subsec:forces})              & 0.406              & 3.677               \\
 		\cline{2-4}
-        & Total & 1.531 & 8.265\\
+		                                  & Total                                        & 1.531              & 8.265               \\
 		\hline
@@
 		\cline{2-4}
-        (\S~\ref{sec:solidToFluidAction}) & Mask (\S~\ref{subsec:mask}) & 0.238 & 1.656\\
+		(\S~\ref{sec:solidToFluidAction}) & Mask (\S~\ref{subsec:mask})                  & 0.238              & 1.656               \\
 		\cline{2-4}
-        & Total & 0.338 & 2.487\\
+		                                  & Total                                        & 0.338              & 2.487               \\
 		\hline
-		\rowcolor{bleuC}\textbf{Total} & & \textbf{3.414} & \textbf{12.297}\\
+		\rowcolor{bleuC}\textbf{Total}    &                                              & \textbf{3.414}     & \textbf{12.297}     \\
 		\hline
@@
 \begin{align}
-\frac{\partial \tilde{h}}{\partial t}(\mathbf{k},t) - \frac{\tilde{\partial}\phi}{\partial y}(\mathbf{k}, 0, t) &=0 \nonumber\\
-i\omega(k)\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) - \frac{igk}{\omega(k)} E(\mathbf{k},0)\nonumber\\
- \times \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) &= 0 \nonumber\\
- \omega(k)- \frac{gk}{\omega(k)} &= 0 \nonumber\\
+	\frac{\partial \tilde{h}}{\partial t}(\mathbf{k},t) - \frac{\tilde{\partial}\phi}{\partial y}(\mathbf{k}, 0, t) & =0 \nonumber                                \\
+	i\omega(k)\left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right) - \frac{igk}{\omega(k)} E(\mathbf{k},0)\nonumber \\
+	\times \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)       & = 0 \nonumber                               \\
+	\omega(k)- \frac{gk}{\omega(k)}                                                                                 & = 0 \nonumber                               \\
 	\omega(k) = \sqrt{gk} \label{equa:relationDispersionDeduce}

### 2025-02-06 review/review.tex
Update the graph with log scale
Add some information about geometrical computation

@@
 
-	      \answer{TODO.}
+	      \answer{We have added details in the article to clarify the implementation.
+Geometry computation is one of the most intensive steps, we think that it might comes from uncoalesced memory accesses and warp branch divergence which may contribute to the observed bottleneck.}
 
@@
 
-	      \answer{TODO.}
+	      \answer{We appreciate your suggestion.
+We have updated Figure 6 to use a logarithmic vertical scale, which significantly improves readability by better separating the data and making the variations clearer.}
 

### 2025-02-17 review/review.tex
Add an answer to stabbing ray robustness

@@
 
-	      \answer{TODO what robustness issues and instability ?}
+	      \answer{The stabbing ray approach can encounter numerical errors when a ray hits a corner, potentially leading to misclassification of intersections, but this is mitigated by using a bounded interval and stable parameter numbering to ensure robustness.}
 

### 2025-02-26 paper.tex
Fixed bibliography to have url every where

@@
 \begin{itemize}
-	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2001) \cite{tessendorfSimulating2001} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
+	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2001) \cite{tessendorfSimulatingOceanWater2001} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
 	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
@@
 J.
-Tessendorf (2001) \cite{tessendorfSimulating2001} presented what could be considered a seminal work in real-time ocean simulation.
+Tessendorf (2001) \cite{tessendorfSimulatingOceanWater2001} presented what could be considered a seminal work in real-time ocean simulation.
 His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solution to an approximation of the Navier-Stokes equations on a 2D height field.
@@
 \label{subsec:waterHeight}
-Tessendorf (2001) \cite{tessendorfSimulating2001} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
+Tessendorf (2001) \cite{tessendorfSimulatingOceanWater2001} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
 \begin{equation}
@@
 Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
-In its original paper, Tessendorf (2001) \cite{tessendorfSimulating2001} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
+In its original paper, Tessendorf (2001) \cite{tessendorfSimulatingOceanWater2001} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
 \begin{equation}
@@
 \Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x(\mathbf x,t)\ D_z(\mathbf x,t)\right]^T$ of seawater to simulate more realistic choppy waves.
-As proposed in Sections 4.5 and 4.6 of Tessendorf (2001) \cite{tessendorfSimulating2001}, it is calculated as follows:
+As proposed in Sections 4.5 and 4.6 of Tessendorf (2001) \cite{tessendorfSimulatingOceanWater2001}, it is calculated as follows:
 \begin{equation}\label{equa:hori-displacement}
@@
 %-------------------------------------------------------------------------
+% The bibliography as been exported from Zotero in BibLateX (not in Better BiblateX)
 \small

### 2025-02-28 paper.tex
Add orcid logo
Fixed bibliography

@@
 \author{%
-    David Algis\\Université de Poitiers%\\Univ. Limoges\\CNRS, XLIM, France\\Studio Nyx
+    David Algis~\href{https://orcid.org/0009-0004-6033-2389}{\includegraphics[width=8pt]{Figures/ORCIDlogo}}\\Université de Poitiers%\\Univ. Limoges\\CNRS, XLIM, France\\Studio Nyx
     \and Berenger Bramas\\INRIA Nancy
@@
 He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
-Yuksel (2007) \cite{yuksel2007} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
+Yuksel (2007) \cite{yukselWaveParticles2007} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
 Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
@@
 The buoyancy calculation presented in Section \ref{subsubsec:buoyancy} requires the full submerged volume $v_w$ of the mesh.
-\Framework uses the prism approximation similarly to the approximation proposed by Bajo \etal (2020) \cite{bajoRealisticBuoyancyModel2020a}.
+\Framework uses the prism approximation similarly to the approximation proposed by Bajo \etal (2020) \cite{bajoRealisticBuoyancyModel2020}.
 While Bajo \etal compute the submerged volume at the texel level, the same principle can be applied using triangles instead of texels using as they did the divergence theorem.

### 2025-03-04 paper.tex
Add an accent to bérenger everywhere
Add a test for author
Remove jcgt dependency

@@
-% !TEX options=-synctex=-1 %for inverse forward search 
-\documentclass[final]{jcgt}
+% !TEX options=-synctex=-1
+\documentclass[11pt]{article}
 
@@
 \usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \newtheorem{theorem}{Theorem}
@@
 
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=bleuF,
+    citecolor=bleuF,
+    filecolor=magenta,      
+    urlcolor=bleuF,
+    pdftitle={Arc Blanc: a real time ocean simulation framework},
+    pdfpagemode=FullScreen,
+}
+\usepackage{authblk}
 \usepackage{algorithm}
@@
 \usepackage{multirow}
-
-\setciteauthor{David Algis, Bérenger Bramas, Emmanuelle Darles, Lilian Aveneau}
-\setcitetitle{Arc Blanc: a real time ocean simulation framework}
-
-% Mark submissions with the date of submission using the following line:
-\submitted{\today}
-
-% Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
-\accepted{submitted}{accepted}{published}{Editor Name}{vol}{issue}{1}{1}{year}
-\seturl{http://jcgt.org/published/vol/issue/num/}
 
@@
 
-\author{%
-    David Algis~\href{https://orcid.org/0009-0004-6033-2389}{\includegraphics[width=8pt]{Figures/ORCIDlogo}}\\Université de Poitiers%\\Univ. Limoges\\CNRS, XLIM, France\\Studio Nyx
-    \and Berenger Bramas\\INRIA Nancy
-    \and Emmanuelle Darles\\Université de Poitiers%\\ Univ. Limoges\\ CNRS, XLIM, France
-    \and Lilian Aveneau\\Université de Poitiers%\\ Univ. Limoges\\ CNRS, XLIM, France
+\author[1,2,*]{David Algis}
+\author[3]{Berenger Bramas}
+\author[1]{Emmanuelle Darles}
+\author[1]{Lilian Aveneau}
+
+\affil[1]{Université de Poitiers, Univ. Limoges, CNRS, XLIM, France}
+\affil[2]{Studio Nyx}
+\affil[3]{Inria Nancy Grand Est, ICube, France}
+
+\affil[*]{%
+  Corresponding Author : \href{mailto:david.algis@univ-poitiers.fr}{david.algis@univ-poitiers.fr}
 }
-
-\teaser{
+% To avoid using \today as suggested by arxiv : https://info.arxiv.org/help/faq/today.html
+\date{2025-02-13}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+\maketitle
+
+\begin{figure}[h]
+	\centering
     \includegraphics[width=4in]{Figures/experiment1-contrast.png}
@@
     \label{fig:teaser}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-\thispagestyle{firstpagestyle}
-
-\maketitle
+\end{figure}
 
@@
 Thanks to Naval Group for initiating this project.
-Finally, we would also like to thank Yannick Privat at the \href{https://irma.math.unistra.fr/}{Institut de recherche mathématique avancée (IRMA)} which help us on the velocity calculation.
+We would also like to thank Yannick Privat at the \href{https://irma.math.unistra.fr/}{Institut de recherche mathématique avancée (IRMA)} which help us on the velocity calculation.
+Finally, we would like to thanks the reviewers and editors of the Journal of Computer Graphics Techniques (JCGT) for their valuable comments and proposals to this paper.
+The original publication of this paper is available on their website \url{https://jcgt.org/}.
 
@@
 \small
-\bibliographystyle{jcgt}
+\bibliographystyle{plain}
 \bibliography{bibliography}
@@
 
-%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%
-\afterdoc
 %%%%%%%%%%%%%%%

### 2025-03-04 paper.tex
Use a compress version of the images

@@
 	\centering
-    \includegraphics[width=4in]{Figures/experiment1-contrast.png}
+    \includegraphics[width=4in]{Figures/experiment1-contrast.jpg}
     \caption{Example of rendering of \framework.}
@@
 	{
-		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion0SeaState5-contrast.png}
+		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion0SeaState5-contrast.jpg}
 		\label{fig:dispersion0}
@@
 	{
-		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion1SeaState5-contrast.png}
+		\includegraphics[width=0.45\textwidth]{Figures/seaDispersion1SeaState5-contrast.jpg}
 		\label{fig:dispersion1}
@@
 	\centering
-	\subfloat{\includegraphics[width=0.7\textwidth]{Figures/gridInterpolationDegree8.png}\label{fig:gridInterpolationDegee8}}
+	\subfloat{\includegraphics[width=0.7\textwidth]{Figures/gridInterpolationDegree8.jpg}\label{fig:gridInterpolationDegee8}}
 	\hfill
-	\subfloat{\includegraphics[width=0.3\textwidth]{Figures/gridInterpolationDegree8Legend.png}\label{fig:gridInterpolationDegree8Legend}}
+	\subfloat{\includegraphics[width=0.3\textwidth]{Figures/gridInterpolationDegree8Legend.jpg}\label{fig:gridInterpolationDegree8Legend}}
 	\caption{This figure shows a comparison between a \textcolor{bleuF}{uniform discretization in blue} and a \textcolor{redF}{logarithmic discretization in red}.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/interpolationAccuracyComparisonLogScale.png}\label{subfig:interpolationAccuracyComparisonLogScale}
+	\includegraphics[width=0.8\textwidth]{Figures/interpolationAccuracyComparisonLogScale.jpg}\label{subfig:interpolationAccuracyComparisonLogScale}
 	\caption{This figure represents the accuracy of different type of interpolation in function of the size of the discretization $N$. The accuracy axis used a logarithmic scale to be more readable.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/interpolationMeanPerformancesAndAccuracy.png}
+	\includegraphics[width=0.8\textwidth]{Figures/interpolationMeanPerformancesAndAccuracy.jpg}
 	\caption{These figures represent the different objective for choosing the interpolation degree.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/multiObjectiveAccuracy1Duration2.png}
+	\includegraphics[width=0.8\textwidth]{Figures/multiObjectiveAccuracy1Duration2.jpg}
 	\caption{Multi-objective function $P$ to optimize, with two objectives: performance and accuracy.
@@
 	{
-		\includegraphics[width=0.25\textwidth]{Figures/mask_diagonal_view.png}
+		\includegraphics[width=0.25\textwidth]{Figures/mask_diagonal_view.jpg}
 	}
@@
 	{
-		\includegraphics[width=0.25\textwidth]{Figures/mask_back_view.png}
+		\includegraphics[width=0.25\textwidth]{Figures/mask_back_view.jpg}
 	}
@@
 	{
-		\includegraphics[width=0.25\textwidth]{Figures/mask_side_view.png}
+		\includegraphics[width=0.25\textwidth]{Figures/mask_side_view.jpg}
 	}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/MaskHeightMap.png}
+		\includegraphics[width=0.4\textwidth]{Figures/MaskHeightMap.jpg}
 		\label{subfig:maskHeightMap}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/Mask3D-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/Mask3D-contrast.jpg}
 		\label{subfig:mask3D}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/experiment1-1-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/experiment1-1-contrast.jpg}
 		\label{fig:experiment1-1}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/experiment1-2-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/experiment1-2-contrast.jpg}
 		\label{fig:experiment1-2}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/experiment2-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/experiment2-contrast.jpg}
 		\label{fig:experiment2-1}
@@
 	{
-		\includegraphics[width=0.4\textwidth]{Figures/experiment2-3-contrast.png}
+		\includegraphics[width=0.4\textwidth]{Figures/experiment2-3-contrast.jpg}
 		\label{fig:experiment2-2}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{Figures/results_cm.png}
+	\includegraphics[width=0.9\textwidth]{Figures/results_cm.jpg}
 	\caption{Mean number of steps required for convergence as a function of wind speed. Error bars represent the variance of iterations across $1000$ test points.}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{Figures/resultsExecutionTime.png}
+	\includegraphics[width=0.9\textwidth]{Figures/resultsExecutionTime.jpg}
 	\caption{Average execution time per iteration count over $1000$ tests, measured across $10^5$ sample positions. Error bars indicate variance in execution time.}
